#inputting .csv file into R
```{r}
reading <- "/home/CAMPUS/mwl04747/github/SOPs/Soil_Texture_Analysis/Data/reading.csv"
sample <- "/home/CAMPUS/mwl04747/github/SOPs/Soil_Texture_Analysis/Data/sample.csv"
reading.df <- read.csv(reading)
sample.df <- read.csv(sample)
```

## Check Data Importing
```{r}
View(reading.df)
head(reading.df)
head(sample.df)
```

## Preprocess Data
```{r preprocess}
library(mosaic)

```
## Join Data with Key.

First, I created two files that have only a few columns each so we can test the join function.

```{r joinkeyID}
reading.tmp <- select(reading.df, sampleKEY, timeElapsed)
sample.tmp <- select(sample.df, sampleKEY, researcher)

left_join(sample.tmp, reading.tmp, by = "sampleKEY")
```


readingtest <-
+ reading %>%
+ select(sampleKEY, time, timeElapsed, actual, correct, temp, K, depth, diameter, PFpartial, PFtotal) %>%
+ leftjoin(sample %>% select(sampleKEY, project, sampleID, date, researcher, computed, checked, hydrometer, pretreat, disperse, gsp, pass10, retain10, tinNum, tinAir, tinOven, passPercent, hygroPercent, ovenDesired, ovenEquiv), by = c("SampleKEY" = "SampleKEY"))
Error in UseMethod("select_") : 
  no applicable method for 'select_' applied to an object of class "character"


#merge function used to combine two tables with common variable(s)
test <- merge(reading, sample)
