---
title: "Soil Texture Processing File"
author: "Marc Los Huertos"
date: "7/26/2016"
output:
  html_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Uploading and Downloading the Data

[Database Access](http://thinkertools.org/grainsize/) 

"18Pomona87"

## Reading .csv Data into R
```{r}
# ChooseRead()

reading.source <- "/home/CAMPUS/mwl04747/github/SOPs/32_Soil_Texture_Analysis/Data/reading.csv"
sample.source <- "/home/CAMPUS/mwl04747/github/SOPs/32_Soil_Texture_Analysis/Data/sample.csv"
reading.import <- read.csv(reading.source)
sample.import <- read.csv(sample.source)
```

## Check Data Importing
```{r}

head(reading.import)
head(sample.import)

library(mosaic)
library(dplyr)

# Check sample file
BadsampleKEYs <- filter(sample.import, is.na(actualSoil) == T | is.na(totalSoil)==T) %>% select(sampleKEY); BadsampleKEYs
BadreadingKEYs <- filter(reading.import, is.na(timeUnit)==T) %>% select(sampleKEY) 
 
# filter(sample, is.na(totalSoil)==T)
badKEYs <- full_join(BadsampleKEYs, BadreadingKEYs); badKEYs
reading <- reading.import %>% anti_join(badKEYs, by="sampleKEY"); 
sample <- sample.import %>% anti_join(badKEYs, by="sampleKEY");

# Write out Unprocessed samples
unprocessed <- sample.import %>% inner_join(badKEYs) %>% select(sampleKEY, sampleID, researcher); 

write.csv(unprocessed, file = paste("/home/CAMPUS/mwl04747/github/SOPs/32_Soil_Texture_Analysis/Data/unprocessed_",Sys.Date(),".csv", sep=""))
```

## Preprocess Sample File
```{r preprocess}
library(mosaic)
library(dplyr)

source("Soil_functions/PassNo10_fun.R")

sample$passNo10Percent =PassNo10(sample$pass10, sample$totalSoil)

#sample %>% select(passNo10Percent, passPercent)

source("Soil_functions/SoilEquiv_fun.R")

sample$hygroPercent2 = hygro(sample$tinAir, sample$tinOven, sample$tinTare); #head(sample)
sample$WS_e = WS_e(sample$tinAir, sample$tinOven, sample$tinTare, sample$actualSoil);

sample$GSp = 2.65

#head(sample)

SampleParams <- sample %>% select(sampleKEY, WS_e, passNo10Percent, hygroPercent2, GSp); 

#head(SampleParams)
  
```

## Preprocess Reading File

```{r preprocess_reading}

# Convert Times to Minutes
reading$Et <- round(reading$timeElapsed * ElapsedTimeConversion(reading$timeUnit), 2)
# reading$Et

```
## Join SampleParams and Reading

First, I created two files that have only a few columns each so we can test the join function.

```{r joinkeyID}
reading.select <- select(reading, sampleKEY, Et, actual, blank, temp); #head(reading.select)

joined2 <- left_join(SampleParams, reading.select, by = "sampleKEY")
```

## Loop sampleKEY to Impute Missing Temp and Blanks

### Table for Alignment for Missing Hydrometer and Temperature readings

We will align missing values to closest readings: 

| Reading | Missing Range |
| --------| ------------- | 
| 0       | 0-30          |
| 60      | 31-90         |
| 120     | 91-240        |
| 1440    | >240          |

I would like to create a more autmotic function for this, but I couldn't figure this out in an elegant way!

## Using Non-Missing to Impute Missing Temp and RBlank

I still believe there is a more elegant to do this, but this is better than my initial idea!

### Ways to Improve the Function:

First, we could make the function that address both missing values simultaneously. For example, impute(sampleID, parameter)...  I tried this, but ran into the problem that we use different columns that are identified in differeing ways. Perhaps, next version!

```{r impute_loop}

joined <- joined2[!is.na(joined2$Et),]; joined
KEY <- unique(joined$sampleKEY); KEY

#KEY <- KEY[-c(1, 2, 3)]; KEY
joined$temp2 = NA
joined$blank2 = NA

#for(i in 1:length(KEY)){
for(i in 1:length(KEY)){
  # for tesing purposes... no 22
  #i = 77
tmp = joined[joined$sampleKEY==KEY[i],]; tmp

# Temperature Imputing
tmp$temp[is.na(tmp$temp)] <- imputeT(tmp)
joined[joined$sampleKEY==KEY[i],]$temp2 <- tmp$temp

# Reading Blank Imputing
tmp$blank[is.na(tmp$blank)] <- imputeR(tmp)
joined[joined$sampleKEY==KEY[i],]$blank2 <- tmp$blank
}

```

## RcTest

Not sure what we are doing here... Isaac, can you comment?

Corrected Rc is based on the formula...



```{r RcTest}
joined$Rc <- joined$actual - (joined$blank2 - 1)
```

## Calculating K

```{r calculatingK}
joined$K <- Kfun(round(joined$temp2,0)); head(joined)
```


## Calculate Effective Depth

Effective Depth...

```{r effectivedepth}
# For High Density Hydrometers
# x = c(1.00, 1.07)
# y = c(15.5, 0.6)

# For H151
x = c(1.00, 1.031)
y = c(10.5, 2.3)
L2coef = coef(lm(y~x))
L1 = joined$Rc * L2coef[2] + L2coef[1]
L2 = 14.0
Vb = 67
A = 27.8

joined$EffectiveDepth = L1 + 0.5*(L2-(Vb/A)); joined$EffectiveDepth
rm(x, y, L2coef, L1, L2, Vb, A)

```
## Calculate D
According to ASTM D422-63 the equation to calculate D (diameter in mm) can be  simplified as D = K*sqrt(L/T) where: L=effective depth (Leff), T = measurment time in min and K is the constant given in the tables that varies with temperature (already calculated).

```{r De}
joined$De<-joined$K * sqrt(joined$EffectiveDepth/joined$Et)
joined$De
```

## Calculate PF

P=[(100,000/w)*G/(G-G1)]*(R-G1) (pretty sure G1=1, pretty sure R = Rc (ask marc about this one)) and G=2.65 g/cm^3 (ask marc about this one too)

The bracketed portion will be the same for all variables thus we can calculate it separately and call it p1 (1st part of equation for )

```{r perfecentfiner}
joined$W = joined$WS_e/joined$passNo10Percent * 100
joined$PF= ((100000/joined$W)*(joined$GSp/(joined$GSp-1)))*(joined$Rc - 1)
```

## Creating Figures

You can also embed plots, for example:

```{r figure, echo=T}
KEY <- unique(joined$sampleKEY)
KEY <- KEY[-c(1:25)]
           
for(i in 1:length(KEY)){
#with(joined[joined$sampleKEY==KEY[i],], plot(x=log(De), y=PF, xlab="Log of Diameter (mm)", ylab="Percent Finer", las=1))
}

```

## Estimating the Texture Size Classes

```{r interpolation}
clay <- 0.002; silt <- 0.05; sand <- 2.0

#joined$sampleKEY[is.nan(joined$De)]
#joined$status="good"
#badKEYs <- data.frame(sampleKEY =unique(joined$sampleKEY[is.na(joined$De)]), status = "bad"); badKEYs
#badKEYs2 <- data.frame(sampleKEY =unique(joined$sampleKEY[is.na(joined$PF)]), status = "bad"); badKEYs2 
#badKEYs <- full_join(badKEYs, badKEYs2)
#joined3 <- joined %>% anti_join(badKEYs, by="sampleKEY"); 
#joined3 <- goodKEYs %>% filter(


KEYs <- unique(joined$sampleKEY); KEYs
#KEY <- KEY[-c(1:25, 47)]; KEY

variables = 4 
iterations = length(KEYs)
textureclasses <- matrix(ncol=variables, nrow=iterations)
 
for(i in 1:iterations){
tmp3 = joined[joined$sampleKEY==KEYs[i],]; head(tmp3)
tmp3 = select(joined, sampleKEY, De, PF) %>% filter(joined$sampleKEY==KEYs[i]); tmp3 

Perclay = round(Interpolation(tmp3, clay),1)
Persilt = round(Interpolation(tmp3, silt), 1)
Persand = round(100-(Persilt + Perclay), 1)
textureclasses[i,] <- c(KEYs[i], Persand, Persilt, Perclay)
}

textureclasses
 
textureclasses <- data.frame(textureclasses)
names(textureclasses) <- c("sampleKEY", "Sand", "Silt", "Clay")

class(textureclasses)

sample2 <- sample %>% select(sampleKEY, sampleID, sampleDate, researcher)

sample3 <- sample2 %>% left_join(textureclasses, by="sampleKEY")

write.csv(sample3, file = paste("/home/CAMPUS/mwl04747/github/SOPs/32_Soil_Texture_Analysis/Data/results_",Sys.Date(),".csv", sep=""))

```

