---
title: "Soil Texture Processing File"
author: "Marc Los Huertos"
date: "7/26/2016"
output:
  html_document: default
  word_document: default
---

# Prepare and and Start Functions for Processing
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

source("Soil_functions/PassNo10_fun.R")
source("Soil_functions/RemoveMissingKEYs_fun.R")
source("Soil_functions/SoilEquiv_fun.R")
source("Soil_functions/TimeConversion_fun.R")
source("Soil_functions/Kfun.R")
source("Soil_functions/Hygro_fun.R")
source("Soil_functions/EffectiveSoilWeight_fun.R")
```

## Uploading and Downloading the Data

[Database Access](http://thinkertools.org/grainsize/) 

"18Pomona87"

## Reading .csv Data into R
```{r}
# ChooseRead()

reading.source <- "/home/CAMPUS/mwl04747/github/SOPs/32_Soil_Texture_Analysis/Data/reading.csv"
sample.source <- "/home/CAMPUS/mwl04747/github/SOPs/32_Soil_Texture_Analysis/Data/sample.csv"
reading.import <- read.csv(reading.source)
sample.import <- read.csv(sample.source)
```

## Check Data Importing
```{r}

head(reading.import)
head(sample.import)

library(mosaic)
library(dplyr)


sample <- RemoveMissingKEYs(sample.import, reading.import)

```

# Preprocess Sample File

## Calculate Percent Soil Passing Seive No. 10
```{r preprocess}
library(mosaic)
library(dplyr)
sample$passNo10Percent <- PassNo10(sample$pass10, sample$totalSoil)
```

## Calculate Percent Hygroscopic Water

```{r hygroscopic}
sample$hygroPercent2 = hygro(sample$tinAir, sample$tinOven, sample$tinTare); #head(sample)
sample$WS_e = WS_e(sample$tinAir, sample$tinOven, sample$tinTare, sample$actualSoil);
```

## Assign Soil Particle Density if Missing
```{r soildensity}
sample$GSp = 2.65
```

## Select Soil Parameters for Processing Reading File
```{r soilparams}
SampleParams <- sample %>% select(sampleKEY, WS_e, passNo10Percent, hygroPercent2, GSp); 
#head(SampleParams)
  
```

# Process Hydrometer Readings

## Preprocess Reading File


### Remove badKEYs from reading
```{r preprocess_reading}
joined <- inner_join(reading.import, SampleParams, by="sampleKEY")
```

### Convert Et to Minutes
```{r elapsedtime}
joined$Et <- round(joined$timeElapsed * ElapsedTimeConversion(joined$timeUnit), 2)
```

### Subset joined file
```{r subsetjoin}
joined2 <- joined %>% select(sampleKEY, passNo10Percent, GSp, Et, WS_e, actual, blank, temp) 
```

## Join SampleParams and Reading SECTION IS MOOT

First, I created two files that have only a few columns each so we can test the join function.

```{r joinkeyID}
#reading.select <- select(reading, sampleKEY, Et, actual, blank, temp); #head(reading.select)

#joined2 <- left_join(SampleParams, reading.select, by = "sampleKEY")
```

## Loop sampleKEY to Impute Missing Temp and Blanks

## Using Non-Missing to Impute Missing Temp and RBlank

I still believe there is a more elegant to do this, but this is better than my initial idea!

### Ways to Improve the Function:

First, we could make the function that address both missing values simultaneously. For example, impute(sampleID, parameter)...  I tried this, but ran into the problem that we use different columns that are identified in differeing ways. Perhaps, next version!

```{r imputefunction}
joined <- impute(joined2)
```

## Calculate the Corrected Hydrometer Reading

Not sure what we are doing here... Isaac, can you comment?

Corrected Rc is based on the formula...



```{r CorrectedReading}
joined$Rc <- joined$actual - (joined$blank2 - 1)
```

## Calculating K

```{r calculatingK}

joined$K <- Kfun(round(joined$temp2,0)); head(joined)
```


## Calculate Effective Depth

Effective Depth...

```{r effectivedepth}
# For High Density Hydrometers
# x = c(1.00, 1.07)
# y = c(15.5, 0.6)

# For H151
x = c(1.00, 1.031)
y = c(10.5, 2.3)
L2coef = coef(lm(y~x))
L1 = joined$Rc * L2coef[2] + L2coef[1]
L2 = 14.0
Vb = 67
A = 27.8

joined$EffectiveDepth = L1 + 0.5*(L2-(Vb/A)); joined$EffectiveDepth
rm(x, y, L2coef, L1, L2, Vb, A)

```
## Calculate D
According to ASTM D422-63 the equation to calculate D (diameter in mm) can be  simplified as D = K*sqrt(L/T) where: L=effective depth (Leff), T = measurment time in min and K is the constant given in the tables that varies with temperature (already calculated).

```{r De}
joined$De<-joined$K * sqrt(joined$EffectiveDepth/joined$Et)
joined$De
```

## Calculate PF

P=[(100,000/w)*G/(G-G1)]*(R-G1) (pretty sure G1=1, pretty sure R = Rc (ask marc about this one)) and G=2.65 g/cm^3 (ask marc about this one too)

The bracketed portion will be the same for all variables thus we can calculate it separately and call it p1 (1st part of equation for )

```{r perfecentfiner}
joined$W = joined$WS_e/joined$passNo10Percent * 100
joined$PF= ((100000/joined$W)*(joined$GSp/(joined$GSp-1)))*(joined$Rc - 1)
```

## Creating Figures

You can also embed plots, for example:

```{r figure, echo=T}
KEY <- unique(joined$sampleKEY)
KEY <- KEY[-c(1:25)]
           
for(i in 1:length(KEY)){
#with(joined[joined$sampleKEY==KEY[i],], plot(x=log(De), y=PF, xlab="Log of Diameter (mm)", ylab="Percent Finer", las=1))
}

```

## Estimating the Texture Size Classes

```{r interpolation}
clay <- 0.002; silt <- 0.05; sand <- 2.0

#joined$sampleKEY[is.nan(joined$De)]
#joined$status="good"
#badKEYs <- data.frame(sampleKEY =unique(joined$sampleKEY[is.na(joined$De)]), status = "bad"); badKEYs
#badKEYs2 <- data.frame(sampleKEY =unique(joined$sampleKEY[is.na(joined$PF)]), status = "bad"); badKEYs2 
#badKEYs <- full_join(badKEYs, badKEYs2)
#joined3 <- joined %>% anti_join(badKEYs, by="sampleKEY"); 
#joined3 <- goodKEYs %>% filter(


KEYs <- unique(joined$sampleKEY); KEYs
#KEY <- KEY[-c(1:25, 47)]; KEY

variables = 4 
iterations = length(KEYs)
textureclasses <- matrix(ncol=variables, nrow=iterations)
 
for(i in 1:iterations){
tmp3 = joined[joined$sampleKEY==KEYs[i],]; head(tmp3)
tmp3 = select(joined, sampleKEY, De, PF) %>% filter(joined$sampleKEY==KEYs[i]); tmp3 

Perclay = round(Interpolation(tmp3, clay),1)
Persilt = round(Interpolation(tmp3, silt), 1)
Persand = round(100-(Persilt + Perclay), 1)
textureclasses[i,] <- c(KEYs[i], Persand, Persilt, Perclay)
}

textureclasses
 
textureclasses <- data.frame(textureclasses)
names(textureclasses) <- c("sampleKEY", "Sand", "Silt", "Clay")

class(textureclasses)

sample2 <- sample %>% select(sampleKEY, sampleID, sampleDate, researcher)

sample3 <- sample2 %>% left_join(textureclasses, by="sampleKEY")

write.csv(sample3, file = paste("/home/CAMPUS/mwl04747/github/SOPs/32_Soil_Texture_Analysis/Data/results_",Sys.Date(),".csv", sep=""))

```

