---
title: "R, Rstudio, and Github"
author: "Marc Los Huertos"
date: "6/19/2017"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## R

### Programming Environment

### Statistical 

## Rstudio

### Rstudio -- Desktop and Server

### Reproducibility w/ R Script


## Reproducibility w/ R Markdown

This is an R Markdown presentation. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document.

## Slide with Bullets

- Bullet 1
- Bullet 2
- Bullet 3

## Slide with R Output

```{r cars, echo = TRUE}
summary(cars)
```

## Slide with Plot

```{r pressure}
plot(pressure)
```


## LaTeX and Manuscript Preparation

### What is LaTeX?

LaTeX is a high-quality typesetting system; it includes features designed for the production of technical and scientific documentation. LaTeX [some like to think] is  the de facto standard for the communication and publication of scientific documents. 

NOTE: LaTeX is a typesetting system, not a word processor. LaTeX uses source code to generate a document: you edit one file and typeset it to a generate the output (usually PDF nowadays). This is in contrast to a word processor, where you edit the text as it appears

### Why use LaTeX to create documents?

## How to integrete R and LaTeX?



## Github and Developing Repositories

### The Daily Process:

Pull, edit files, save files, Commit, and Push


