---
title: "Post_Analysis_Processing"
author: "Marc Los Huertos"
date: "5/2/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

The IRMS records values for CN(S) that reflect the performance of the instrument that require corrections. First, there are biases that fractionate the sample by internal processes of the instrument -- 

# Data Processing
## Importing Data

```{r}
# read in the first worksheet from the workbook myexcel.xlsx
# first row contains variable names
library(xlsx)
# file = file.choose()
file = "/home/CAMPUS/mwl04747/github/SOPs/75_IRMS/Data/c20180425_EA199_Seq1_MMC.xls"

IRMS <- read.xlsx(file, 1)

# read in the worksheet named mysheet
# mydata <- read.xlsx(file, sheetName = "mysheet")

```
## Black Correction 



## Signal Intensity Correction

The dependence of signal intensity is one of the most important sources of error in continuous-flow IRMS, where the $\delta$ become biased --- especially at low values.

We can correct the fractionation by using the peak amplitude which depends on mass of the sample using standards with various masses. Based on some papers, peak area is more commonly used.

Below is a ``one-point emperical equation'' (Wang, 2017). 

1. Extract Peak values for standards
 
```{r}
(LinearCorr <- IRMS[IRMS$Identifier.1 == "Sulphanilamide",c(5:6, 11:13)])
LinearCorr$mass=as.numeric(LinearCorr$Identifier.2)
```

2. Create Linear Model
```{r}
LinearCorr.lm <- lm(d.15N.14N ~ log(Ampl..28), data=LinearCorr)
summary(LinearCorr.lm)

```

3. Use m and b to correct $\delta_m$
```{r}

IRMS$correction = coef(LinearCorr.lm)[2]* log(IRMS$Ampl..28) + coef(LinearCorr.lm)[1]

```

Adjusting to assigned standard based on supplier's specs:

```{r}
Standard = -1.222

IRMS$d15_cor = IRMS$d.15N.14N + (Standard-IRMS$correction)
```

## Instrument Drift Correction

Time Correction

## Correction with (Certified) Standards


