---
title: "Air Temperature Monitoring with Portable Sensors"
sop-number: "SOP-F501"
author:
  - name: "Marc Los Huertos"
    affiliation: "Environmental Analysis, Pomona College"
date: last-modified
date-format: "YYYY-MM-DD"
revision: "0.2"

revision-history:
  - version: "1.0"
    date: "2026-05-18"
    author: "Marc Los Huertos"
    changes: "Initial release"
  - version: "0.2"
    date: "2026-05-29"
    author: "Marc Los Huertos"
    changes: "Moved definitions to Section 5; added citations for urban climatology, LCZ classification, and crowdsourced QC; expanded UHI and shielding discussion"
---

## Purpose

This SOP describes the deployment, calibration, data retrieval, and
post-processing of low-cost portable sensor networks for ambient air
temperature and relative humidity monitoring. Three sensor platforms
are covered: Maxim/Analog Devices Thermochron iButton loggers, Onset
HOBO data loggers, and Raspberry Pi single-board computers with BME280
environmental sensor modules.

Dense networks of inexpensive air temperature sensors enable studies of
urban heat island (UHI) effects, microclimatic gradients beneath
photovoltaic arrays, canopy-vs.-open temperature contrasts, and
diurnal and seasonal temperature variability at spatial resolutions
not achievable with conventional weather stations [@muller2013;
@chapman2017]. The UHI is one of the most thoroughly documented
examples of inadvertent human modification of local climate, driven
by impervious surfaces, waste heat, reduced vegetation, and altered
radiative and convective energy exchange [@oke2017; @voogt2003].

When designing a sensor network, characterizing the land cover at each
site using the Local Climate Zone (LCZ) framework provides a
standardized basis for comparing UHI intensities across cities and
studies [@stewart2012].


## Scope

- **Applies to:** Outdoor ambient air temperature and relative
  humidity monitoring at 1.5--2.0 m above ground level using portable
  sensor networks.
- **Does not apply to:** Soil temperature profiles (see
  [SOP-F404](SOP-F404-soil-temperature-profiles.qmd)), indoor
  temperature monitoring, or air quality (PM, gases; see
  [SOP-F502](SOP-F502-air-quality-portable-sensors.qmd)).
- **Location:** EA Program field sites including the Pomona College
  campus, Bernard Field Station, photovoltaic study sites, and
  surrounding Inland Empire communities.


## Responsibilities

| Role              | Responsibility                                          |
|:------------------|:--------------------------------------------------------|
| Lab PI            | Approve monitoring plan, oversee data quality           |
| Field Lead        | Coordinate site access and logistics                    |
| Trained Personnel | Calibrate, deploy, retrieve data, process results       |
| New Students      | Complete training before independent deployment         |


## Safety and Hazards

::: {.callout-warning}
### Hazards

- **Working at height:** Mounting sensors at 1.5--2.0 m on poles or
  structures may require a stepladder. Follow ladder safety protocols.
- **Traffic:** Some monitoring locations are near roadways. Wear
  high-visibility vests.
- **Heat exposure:** Deployment and retrieval in summer in the Inland
  Empire involves significant heat risk. Carry water, work in pairs,
  and schedule field work for early morning when possible.
- **Electrical:** Raspberry Pi units use low-voltage USB power. Do not
  expose connections to rain or standing water without weatherproof
  enclosures.
:::

### Required PPE

- [ ] Closed-toe shoes
- [ ] High-visibility vest (if near roadways)
- [ ] Sun protection (hat, sunscreen)
- [ ] Safety glasses (during mounting/drilling)


## Definitions

- **Urban Heat Island (UHI):** The phenomenon whereby urban areas
  experience higher air temperatures than surrounding rural areas due
  to impervious surfaces, waste heat, reduced vegetation, and altered
  energy balance. UHI intensity ($\Delta T_{\text{UHI}}$) is
  quantified as the difference between an urban and a rural reference
  site at a given time [@oke2017].
- **Local Climate Zone (LCZ):** A classification system that describes
  the physical properties of the local environment (building density,
  height, surface cover, land use) at scales relevant to urban
  temperature variation. LCZ types range from compact high-rise
  (LCZ 1) to scattered trees (LCZ B), and provide a standardized
  framework for describing monitoring sites in UHI studies
  [@stewart2012].
- **Radiation shield:** A louvered or ventilated housing that protects
  a temperature sensor from direct and reflected solar radiation while
  allowing ambient air to flow past the sensor. Without shielding,
  sensors can read 5--15 C above true air temperature in direct sun
  [@hubbart2005; @terando2017].
- **BME280:** A Bosch environmental sensor measuring temperature,
  relative humidity, and barometric pressure. Used with Raspberry Pi
  platforms in the EA Program sensor network.
- **iButton:** A coin-sized temperature (and optionally RH) data
  logger manufactured by Maxim/Analog Devices (Thermochron series).
- **HOBO:** A family of portable data loggers manufactured by Onset
  Computer Corporation.
- **Collocation:** The practice of placing multiple sensors at the
  same location for a period to assess inter-sensor agreement and
  identify systematic offsets. Collocation is the primary calibration
  strategy for dense low-cost sensor networks [@meier2017].
- **Calibration offset:** A systematic difference between a sensor's
  reading and a known reference, determined during collocation and
  applied as a correction to field data.
- **Data completeness:** The percentage of expected readings actually
  recorded during a deployment period. A minimum of 75% completeness
  (18 of 24 hours per day) is required for daily summaries.


## Prerequisites

- [SOP-F001 Field Safety](SOP-F001-field-safety.qmd)
- For Raspberry Pi deployments: basic familiarity with Raspberry Pi
  OS, SSH, and file transfer (see also
  [SOP-F502](SOP-F502-air-quality-portable-sensors.qmd))


## Sensor Platforms

### Comparison

| Feature              | iButton DS1921/DS1923 | HOBO U23 Pro v2       | Raspberry Pi + BME280 |
|:---------------------|:---------------------:|:---------------------:|:---------------------:|
| Approximate cost     | $20--60               | $130--200             | $40--80               |
| Temperature accuracy | $\pm$0.5--1.0 C       | $\pm$0.21 C           | $\pm$0.5 C (BME280)  |
| RH measurement       | DS1923 only           | Yes                   | Yes                   |
| Logging interval     | 1 min -- 255 min      | 1 s -- 18 h           | User-defined (script) |
| Memory               | 2048--8192 readings   | ~42,000 readings      | SD card (unlimited)   |
| Communication        | 1-Wire (download)     | USB (download)        | Wi-Fi / SD card       |
| Battery life         | ~1 year               | ~1 year               | Requires continuous power |
| Size                 | 17 mm coin            | 6 x 5 x 2 cm         | Variable (enclosure)  |

### Choosing a Platform

- **iButton:** Best for dense spatial networks where low cost and
  small size are priorities. Limited memory requires periodic
  retrieval. Ideal for short-term (weeks to months) UHI transects.
- **HOBO:** Best general-purpose option with highest accuracy, long
  battery life, and large memory. Good for season-long deployments.
- **Raspberry Pi + BME280:** Best for real-time or near-real-time
  monitoring with Wi-Fi. Requires continuous power. Already available
  in the EA Program inventory (16 Pi Zero+ units).


## Radiation Shielding {#sec-shielding}

::: {.callout-important}
### Radiation Shields Are Mandatory
Unshielded sensors exposed to direct or reflected solar radiation can
read 5--15 C higher than true air temperature. Every deployed sensor
**must** be housed in a radiation shield [@hubbart2005; @terando2017].
:::

### Shield Options

| Shield Type                | Cost  | Notes                          |
|:---------------------------|:------|:-------------------------------|
| 6-plate gill shield (e.g., Davis 7714) | $20--30 | Standard; good ventilation |
| 10-plate aspirated shield  | $30--50 | Better in low-wind conditions |
| DIY stacked-plate shield (white plastic plates) | $5--10 | Acceptable for student projects; see Appendix B for construction |
| Naturally aspirated PVC tube (white, louvered) | $10--15 | Compact; fits iButtons |

All shields must be white (high albedo) and mounted with the sensor
opening facing downward to prevent rain intrusion.


## Calibration {#sec-calibration}

### Ice-Bath Calibration (Single-Point Check)

1. Prepare an ice-water bath: fill an insulated container with crushed
   ice and deionized water, stir for 2 minutes
2. Submerge the sensor (in a waterproof sleeve for iButtons, or using
   a sealed bag for non-waterproof units)
3. Wait 5 minutes for equilibration
4. The sensor should read 0.0 $\pm$ 0.5 C. Record the reading.
5. If outside tolerance, note the offset; sensors with offsets greater
   than 1.0 C should be replaced.

### Room-Temperature Inter-Comparison (Collocation)

1. Place all sensors to be deployed together in a shaded, well-mixed
   indoor environment for at least 24 hours
2. Calculate the mean temperature across all sensors for each
   timestep. Each sensor's offset is its deviation from the group
   mean.
3. Record offsets in a calibration log file:
   `data/calibration/offsets_YYYYMMDD.csv`
4. Apply these offsets as corrections during data processing
   (Section 11). Collocation under controlled conditions is the most
   practical calibration approach for networks of 10 or more sensors
   [@meier2017].


## Deployment {#sec-deployment}

### Site Selection and Documentation

Select monitoring sites to represent the range of land cover types
and microclimates relevant to the research question. For UHI studies,
sample across multiple LCZ types (e.g., compact low-rise, open
low-rise, large low-rise, scattered trees, dense trees)
[@stewart2012].

At each site, record:

- GPS coordinates per
  [SOP-F006](SOP-F006-gps-site-navigation.qmd)
- Mounting height (standard: 1.5--2.0 m above ground level)
- Shield type and orientation
- Land cover within 50 m radius (impervious %, tree canopy %, grass %)
- LCZ type (see @stewart2012 for classification guide)
- Nearest road (distance and type)
- Sky view factor (estimated or photographed)
- Sensor serial number and platform type

### Mounting

1. Mount the radiation shield on a pole, post, fence, or structure at
   1.5--2.0 m height. Ensure the shield is not directly above a
   paved surface, air conditioning exhaust, or heat-radiating wall.
2. Place the sensor inside the shield, ensuring the sensing element
   is centered within the shield stack.
3. For Raspberry Pi units, route the power cable through the base of
   the enclosure. Secure with cable ties.
4. Photograph the installation and record all metadata.
5. Record metadata as above.


## Data Retrieval

| Platform    | Retrieval Method                          | Frequency           |
|:------------|:------------------------------------------|:--------------------|
| iButton     | Remove, connect to 1-Wire reader, download in OneWireViewer | Every 2--4 weeks |
| HOBO U23    | Connect USB shuttle or cable, download in HOBOware | Every 1--3 months |
| Raspberry Pi| SSH + `scp`, or remove SD card             | Weekly or as needed |

After retrieval, archive the raw file immediately. Never modify the
original.


## Data Processing

### Import and Standardize

```{r}
#| label: import-standardize
#| echo: true
#| eval: false

library(tidyverse)
library(lubridate)

# Example: import iButton CSV
raw <- read_csv("data/raw/site01_ibutton_ABC123.csv",
                skip = 14) %>%
  rename(timestamp = 1, temp_c = 2) %>%
  mutate(
    timestamp = mdy_hms(timestamp, tz = "America/Los_Angeles"),
    sensor_id = "ABC123",
    site_id   = "site01",
    platform  = "iButton"
  )

# Apply calibration offset (from Section 9)
offsets <- read_csv("data/calibration/offsets_YYYYMMDD.csv")

raw <- raw %>%
  left_join(offsets %>% select(sensor_id, offset), by = "sensor_id") %>%
  mutate(temp_corrected = temp_c - offset)
```

### Quality Control

```{r}
#| label: qc-flags
#| echo: true
#| eval: false

raw <- raw %>%
  mutate(
    qc_flag = case_when(
      temp_corrected < -10 | temp_corrected > 55 ~ "out_of_range",
      temp_corrected == lag(temp_corrected) &
        temp_corrected == lag(temp_corrected, 2) &
        temp_corrected == lag(temp_corrected, 3)   ~ "stuck_sensor",
      abs(temp_corrected - lag(temp_corrected)) > 10 ~ "spike",
      TRUE ~ "valid"
    )
  )

count(raw, qc_flag)
clean <- raw %>% filter(qc_flag == "valid")
```

### Hourly and Daily Summaries

```{r}
#| label: summaries
#| echo: true
#| eval: false

hourly <- clean %>%
  mutate(hour = floor_date(timestamp, "hour")) %>%
  group_by(site_id, sensor_id, hour) %>%
  summarize(
    temp_mean = mean(temp_corrected, na.rm = TRUE),
    temp_min  = min(temp_corrected, na.rm = TRUE),
    temp_max  = max(temp_corrected, na.rm = TRUE),
    n_obs     = n(),
    .groups   = "drop"
  ) %>%
  filter(n_obs >= 3)

daily <- hourly %>%
  mutate(date = as_date(hour)) %>%
  group_by(site_id, sensor_id, date) %>%
  summarize(
    t_mean = mean(temp_mean, na.rm = TRUE),
    t_min  = min(temp_min, na.rm = TRUE),
    t_max  = max(temp_max, na.rm = TRUE),
    n_hours = n(),
    .groups = "drop"
  ) %>%
  filter(n_hours >= 18)
```

### Urban Heat Island Intensity

UHI intensity is typically calculated as the difference between an
urban site and a rural (or least-developed) reference site [@oke2017]:

$$
\Delta T_{\text{UHI}} = T_{\text{urban}} - T_{\text{rural}}
$$

```{r}
#| label: uhi-intensity
#| echo: true
#| eval: false

ref_site <- "site_rural01"

uhi <- hourly %>%
  select(site_id, hour, temp_mean) %>%
  pivot_wider(names_from = site_id, values_from = temp_mean) %>%
  mutate(across(-c(hour, all_of(ref_site)),
                ~ . - .data[[ref_site]],
                .names = "uhi_{.col}"))
```


## Quality Control Summary

| Check                                | Criterion                        |
|:--------------------------------------|:---------------------------------|
| Ice-bath calibration                  | Reading within $\pm$0.5 C of 0.0 C |
| Inter-sensor agreement (collocation) | Within $\pm$0.5 C at night      |
| Hourly data completeness             | $\geq$75% of expected readings   |
| Daily data completeness              | $\geq$18 of 24 hours             |
| Stuck-sensor detection               | $\geq$4 identical consecutive readings flagged |
| Spike detection                      | $\geq$10 C change between consecutive readings flagged |


## Data Management

- Raw files archived unmodified: `data/raw/siteID_platform_sensorID.csv`
- Calibration records: `data/calibration/`
- Processed files: `data/processed/`
- Site metadata (coordinates, height, shield type, land cover, LCZ
  type, sensor serial numbers, deployment/retrieval dates):
  `data/metadata/site_metadata.csv`
- All scripts version-controlled in the project Git repository


## References

::: {#refs}
:::


## Related SOPs

| SOP   | Title                                  | Relationship                      |
|:------|:---------------------------------------|:----------------------------------|
| [F001](SOP-F001-field-safety.qmd)       | Field Safety        | Required prerequisite             |
| [F101](SOP-F101-electrical-power-in-the-field.qmd) | Electrical Power in the Field | Power for Pi deployments |
| [F404](SOP-F404-soil-temperature-profiles.qmd) | Soil Temperature Profiles | Companion: soil vs. air temp |
| [F502](SOP-F502-air-quality-portable-sensors.qmd) | Air Quality Monitoring | Companion: PM sensors on same Pi units |
| [F510](SOP-F510-weather-monitoring.qmd) | Weather Monitoring   | Companion: full meteorological context |


## Training Record

Before deploying air temperature sensors independently, personnel
must:

1. Read this SOP and [SOP-F001 Field Safety](SOP-F001-field-safety.qmd)
   in full
2. Perform a calibration (ice-water bath and room-temperature
   inter-comparison) under supervision
3. Assemble and mount a radiation shield
4. Deploy and retrieve a sensor at a test site
5. Import, clean, and summarize data using the R workflow
6. Be signed off by the PI or field lead
