---
title: "Random Sampling Design"
sop-number: "SOP-F005"
author:
  - name: "Marc Los Huertos"
    affiliation: "EA Program, Pomona College"
date: last-modified
date-format: "YYYY-MM-DD"
revision: "1.1"

revision-history:
  - version: "0.1"
    date: "2017-08-18"
    author: "Marc Los Huertos"
    changes: "Initial release (LaTeX/knitr)"
  - version: "1.0"
    date: "2025-05-17"
    author: "Marc Los Huertos"
    changes: "Converted to Quarto; expanded procedure sections for simple
              random, grid-based, and stratified random sampling"
  - version: "1.1"
    date: "2026-06-07"
    author: "Marc Los Huertos"
    changes: "Converted plain-text references to BibTeX; added {#refs} div;
              added inline citations; expanded Definitions"
---

## Purpose

This SOP provides standardized procedures for designing and executing
random sampling in field and laboratory settings. Random sampling
ensures that every unit in the target population has an equal and
independent probability of selection, eliminating researcher bias and
supporting valid statistical inference [@krebs1999; @elzinga1998].


## Scope

- **Applies to:** Any EA Program study where population estimates are
  the goal, including vegetation surveys, soil sampling, water quality
  monitoring, and laboratory subsampling
- **Does not apply to:** Targeted or judgmental sampling designs (e.g.,
  hot-spot investigations); mark-release-recapture studies for mobile
  organisms (see separate SOP)
- **Sampling types covered:** Simple random sampling, grid-based random
  sampling, and stratified random sampling


## Responsibilities

| Role              | Responsibility                              |
|:------------------|:--------------------------------------------|
| PI / Field Lead   | Approve sampling design, review sampling plan |
| Trained Personnel | Execute random selection, collect and record data |
| New Students      | Complete training before independent fieldwork |


## Safety and Hazards

::: {.callout-warning}
### Hazards

Field sampling involves site-specific hazards including uneven terrain,
heat exposure, wildlife encounters, and sharp equipment (stakes, pins).
Review site conditions before each sampling event.
:::

### Required PPE

- [ ] Closed-toe shoes or boots
- [ ] Sun protection (hat, sunscreen)
- [ ] Gloves (when handling equipment or samples)

### Emergency Procedures

Refer to [SOP-F001 Field Safety](SOP-F001-field-safety.qmd). Carry a
first aid kit and charged cell phone to all field sites.


## Definitions

- **Border rule:** A convention for counting organisms that overlap a
  quadrat edge; the standard rule is to count organisms touching the
  top and right edges and exclude those touching the bottom and left
  edges; must be applied consistently within a study to avoid
  double-counting [@elzinga1998].
- **Grid-based random sampling:** A method in which a coordinate grid
  is laid over the study area and quadrat locations are selected using
  randomly generated X and Y coordinates; the standard approach for
  quadrat-based ecological surveys [@elzinga1998].
- **Power analysis:** A calculation to determine the minimum sample
  size ($n$) needed to detect a biologically meaningful effect with
  acceptable Type I error ($\alpha$) and Type II error ($\beta$)
  probability; should be completed before fieldwork begins [@krebs1999].
- **Quadrat:** A square frame of known area (typically 0.25, 0.5, or
  1.0 m^2^) used to delineate sampling units in the field
  [@elzinga1998].
- **Sampling frame:** A numbered list or spatial grid that enumerates
  every unit in the population; units missing from the frame have zero
  probability of selection and introduce bias [@krebs1999].
- **Sampling universe (population):** The complete set of units from
  which a sample will be drawn; must be explicitly defined before
  generating a sampling frame [@krebs1999].
- **Simple random sampling:** A design in which $n$ units are selected
  at random from a numbered list (sampling frame) such that every
  possible combination of $n$ units has an equal probability of
  selection; appropriate when the population is homogeneous
  [@krebs1999].
- **Stratified random sampling:** A design in which the population is
  divided into internally homogeneous subgroups (strata) and random
  samples are drawn independently from each stratum; reduces variance
  compared to simple random sampling when strata differ from each
  other [@krebs1999; @elzinga1998].
- **Stratum:** A subdivision of the sampling universe into zones that
  are internally homogeneous (e.g., by elevation band, vegetation
  type, or soil moisture class); the basis for stratified random
  sampling.


## Required Materials

| Item                                      | Notes                              |
|:------------------------------------------|:-----------------------------------|
| Two survey tape measures (20--50 m)       | For establishing coordinate grids  |
| Ground pins or stakes                     | To anchor tape measures            |
| Quadrat frame(s)                          | Size appropriate to target organisms |
| Random number generator app or R          | See §8.1                           |
| GPS unit                                  | To record sampling locations       |
| Field data sheets and waterproof markers  | Pre-printed with site and date fields |
| Compass                                   | For establishing grid orientation  |
| Field identification guides               | If sampling vegetation or organisms |


## Biases and Interferences

Representative sampling requires strict adherence to randomized
procedures. Common sources of bias include:

- **Incomplete sampling frames.** Any unit missing from the frame has
  zero probability of selection [@krebs1999].
- **Haphazard placement.** Never "throw" a quadrat to choose a
  location; human throws cluster toward the center of the area,
  violating equal probability assumptions [@elzinga1998].
- **Environmental heterogeneity.** If the study area contains distinct
  gradients, simple random sampling may under-represent one end. Use
  stratified random sampling instead (§8.4).
- **Edge effects.** Establish and apply a consistent border rule when
  counting organisms in quadrats (§8.3).


## Procedure

### Generating Random Numbers

Use a software-based random number generator to produce unbiased
selections. Never rely on manual or mental number selection
[@krebs1999].

In R, generate a set of random integers within a defined range:

```{r}
#| label: random-numbers
#| echo: true

# Generate 15 random integers between 1 and 100
set.seed(NULL)  # ensure a fresh seed each session
random_vals <- sample(1:100, size = 15, replace = FALSE)
sort(random_vals)
```

For field use without a computer, a printed random number table is
acceptable. Begin at a pre-selected (arbitrary) starting position and
read values sequentially. @tbl-randomnumbers provides a reference table
of 200 random numbers between 0 and 1000.

```{r}
#| label: tbl-randomnumbers
#| tbl-cap: "Reference table of 200 random numbers (0--1000)"
#| echo: false

set.seed(2017)
rand_mat <- matrix(round(runif(200, min = 0, max = 1000), 1),
                   nrow = 20, ncol = 10)
colnames(rand_mat) <- 1:10
knitr::kable(rand_mat)
```

### Simple Random Sampling

Use this approach when the population can be enumerated as a list
(e.g., lab samples, batch numbers, individual organisms in a database)
[@krebs1999].

1. Define the target population and compile a complete, numbered
   sampling frame.
2. Determine the required sample size ($n$) using a power analysis.
3. Use `sample()` in R to draw $n$ unit numbers without replacement.
4. Document the random seed (if reproducibility is desired), the
   selected unit IDs, and the date of selection.
5. Collect the data or specimens from the selected units.

### Grid-Based Random Sampling (Field)

Use this approach when the sampling universe is a spatially defined
area [@elzinga1998].

#### Establishing the Baseline Grid

1. Locate the boundaries of the study site.
2. Drive a stake at the origin point and lay two tape measures at 90
   degrees to create X and Y axes.
3. Secure the tapes with ground pins to define a coordinate grid
   (e.g., 20 m × 20 m).

#### Generating Coordinates and Placing Quadrats

4. Generate pairs of random integers within the range of each axis:

```{r}
#| label: xy-random
#| echo: true

x_max <- 20; y_max <- 20; n_quadrats <- 15
coords <- data.frame(
  X = sample(0:x_max, size = n_quadrats, replace = TRUE),
  Y = sample(0:y_max, size = n_quadrats, replace = TRUE)
)
coords <- coords[order(coords$X), ]
coords$QuadratID <- seq_len(nrow(coords))
knitr::kable(coords, row.names = FALSE)
```

5. Walk along the axes to each coordinate intersection and place the
   bottom-left corner of the quadrat at that point.
6. Never reposition a quadrat after placement; if a coordinate falls
   on an obstruction, record the obstruction and sample as-is
   [@elzinga1998].

#### Recording Data Within Quadrats

7. Quantify the target variable using one of three standardized
   metrics:
   - **Density:** Count every individual within the frame
   - **Percent cover:** Estimate the fraction of ground area covered
   - **Frequency:** Score presence (1) or absence (0)
8. Apply the border rule consistently (§4).
9. Aim to sample at least 5--10% of total grid area, or a minimum of
   30 quadrat placements, for adequate statistical power [@krebs1999].
10. Calculate the mean count (or cover) per quadrat and extrapolate:

$$
\text{Estimated population} =
\bar{x}_{\text{quadrat}} \times
\frac{\text{Total site area}}{\text{Area of one quadrat}}
$$

### Stratified Random Sampling

Use this approach when the study area contains distinct environmental
gradients that would increase variance under simple random sampling
[@krebs1999; @elzinga1998].

1. Divide the study area into homogeneous strata based on a relevant
   environmental variable.
2. Within each stratum, establish an independent coordinate grid
   following §8.3.
3. Generate random coordinates separately for each stratum.
4. Sample each stratum independently; do not pool random selections
   across strata.
5. Weight stratum means by the proportion of total area each stratum
   represents when calculating site-level estimates.

### Map-Based Sampling with Exclusion Zones

1. Obtain or create a georeferenced map of the study area.
2. Delineate exclusion zones and calculate the remaining accessible
   area.
3. Generate random coordinates across the full extent, then discard
   any that fall within exclusion zones.
4. Continue generating coordinates until the target sample size is
   met within the accessible area.
5. Record discarded coordinates in the sampling log.

::: {.callout-tip}
### R Implementation
An R script for map-based random sampling with exclusion polygons can
be built using the `sf` package. Generate random points with
`sf::st_sample()`, then filter against exclusion polygons with
`sf::st_difference()`.
:::


## Data Management

1. Record all sampling metadata (date, site, grid dimensions, random
   seed or table starting position, personnel) on the field data sheet.
2. Enter field data into the project database within 48 hours.
3. Archive original data sheets (scan if paper).
4. Version-control all analysis scripts per
   [SOP-L801 RStudio and GitHub](../lab_sops/SOP-L801-rstudio-github.qmd).


## References

- @krebs1999
- @elzinga1998
- @apha2012

::: {#refs}
:::


## Related SOPs

| SOP   | Title                                  | Relationship                      |
|:------|:---------------------------------------|:----------------------------------|
| [F001](SOP-F001-field-safety.qmd)      | Field Safety              | Required prerequisite             |
| [F006](SOP-F006-gps-site-navigation.qmd) | GPS and Site Navigation | Navigate to random points         |
| [F603](SOP-F603-vegetation-transect-surveys.qmd) | Vegetation Transect Surveys | Applies this sampling design |


## Training Record

Before designing or executing random sampling independently, personnel
must:

1. Read this SOP in full
2. Complete [SOP-F001 Field Safety](SOP-F001-field-safety.qmd) training
3. Demonstrate proficiency with random number generation in R
4. Participate in a supervised field sampling session using the
   grid-based procedure
5. Be signed off by the PI or field lead
