---
# =============================================================================
# SOP-F502: Air Quality Monitoring (Portable Sensors)
# =============================================================================
# Deployment, operation, data retrieval, and correction of Raspberry Pi
# based low-cost particulate matter sensor networks using Plantower
# PMS5003/PMS5004 sensors.
# =============================================================================

title: "Air Quality Monitoring with Portable PM Sensors"
sop-number: "SOP-F502"
author:
  - name: "Marc Los Huertos"
    affiliation: "Environmental Analysis, Pomona College"
date: last-modified
date-format: "YYYY-MM-DD"
revision: "1.0"

revision-history:
  - version: "1.0"
    date: "2026-05-18"
    author: "Marc Los Huertos"
    changes: "Initial release"
---

## 1. Purpose

This SOP describes the assembly, deployment, data retrieval, and
post-processing of low-cost particulate matter (PM) sensor networks
built around Raspberry Pi single-board computers and Plantower
PMS5003/PMS5004 laser-scattering sensors. These sensors estimate
PM~1~, PM~2.5~, and PM~10~ mass concentrations and report particle
counts in six size bins (>0.3, >0.5, >1.0, >2.5, >5.0, >10.0 $\mu$m).

Low-cost PM sensors (LCS) provide spatial and temporal resolution that
is not achievable with Federal Reference Method (FRM) or Federal
Equivalent Method (FEM) monitors alone. However, LCS data require
quality screening, relative humidity correction, and calibration
against reference instruments to produce scientifically defensible
results [@barkjohn2021; @ardondryer2023].


## 2. Scope

- **Applies to:** Outdoor ambient air quality monitoring using EA
  Program Raspberry Pi + PMS5003/PMS5004 sensor units for PM~1~,
  PM~2.5~, and PM~10~.
- **Does not apply to:** Indoor air quality monitoring (different
  correction factors apply), gas-phase pollutant monitoring (O~3~,
  NO~2~, CO), or regulatory compliance monitoring.
- **Location:** Field deployment sites in the Pomona College area and
  surrounding Inland Empire communities.


## 3. Responsibilities

| Role              | Responsibility                                          |
|:------------------|:--------------------------------------------------------|
| Lab PI            | Approve deployment plan, oversee data quality           |
| Field Lead        | Coordinate site access, logistics, and safety           |
| Trained Personnel | Assemble, deploy, retrieve, and process data            |
| New Students      | Complete training before independent deployment         |


## 4. Safety and Hazards

::: {.callout-warning}
### Hazards

- **Electrical:** Raspberry Pi units and USB power supplies. Do not
  expose connections to rain or standing water. Use weatherproof
  enclosures for outdoor deployment.
- **Heat:** Enclosures can reach high temperatures in direct sunlight,
  especially in summer in the Inland Empire. Use ventilated or shaded
  enclosures.
- **Traffic and site access:** Many monitoring locations are near
  roadways. Follow all traffic safety procedures in
  [SOP-F001](SOP-F001-field-safety.qmd).
- **Working at height:** If mounting sensors on poles or structures,
  follow ladder safety protocols.
:::

### Required PPE

- [ ] Safety glasses (during assembly)
- [ ] Closed-toe shoes
- [ ] High-visibility vest (if near roadways)
- [ ] Sun protection (hat, sunscreen)


## 5. Prerequisites

- [SOP-F01 Field Safety](SOP-F01-field-safety.qmd)
- Basic familiarity with Raspberry Pi OS (command line, SSH, file
  transfer)
- Basic familiarity with R for data processing


## 6. Background

### 6.1 How the Sensors Work

The Plantower PMS5003 and PMS5004 are optical particle counters that
use a laser (680 $\pm$ 10 nm) and a photodiode to detect light
scattered by individual particles drawn through the sensing chamber by
a small fan. The sensor reports particle counts in six size bins and
uses a proprietary algorithm to estimate PM~1~, PM~2.5~, and PM~10~
mass concentrations in $\mu$g/m$^3$ [@barkjohn2021].

The sensor outputs data in two formats: CF=1 ("standard particles,"
factory-calibrated against Chinese urban aerosol) and CF\_atm
("atmospheric environment," with an additional adjustment at higher
concentrations). For ambient monitoring in the US, the CF=1 output is
used as the starting point for correction [@barkjohn2021].

### 6.2 Known Limitations

- **Relative humidity sensitivity:** At RH >70%, hygroscopic particle
  growth inflates readings. An RH correction term is essential
  [@barkjohn2021; @jayaratne2018].
- **Particle size dependence:** The sensors underestimate PM~2.5~ when
  coarse particles (dust) dominate the size distribution. They perform
  best for fine-mode aerosol (PM~1~) [@jaffe2023; @ouimette2026].
- **Batch variability:** Plantower changed the PMS5003 calibration in
  mid-2021, producing sensors that read lower at low concentrations.
  These can be identified by their particle count ratios and corrected
  [@ardondryer2023].
- **Sensor aging:** Fan and laser degradation over months to years of
  continuous operation can reduce sensitivity. Periodic collocation
  with a reference instrument is needed.


## 7. Equipment and Materials

### 7.1 Sensor Unit Components

| Item                       | Qty per Unit | Notes                          |
|:---------------------------|:------------:|:-------------------------------|
| Raspberry Pi Zero+ (or 3+/4+) | 1        | With microSD card (32 GB min)  |
| Plantower PMS5004 (or PMS5003) | 1       | With ribbon cable              |
| BME280 sensor module       | 1            | Temperature, RH, pressure      |
| USB power supply (5V, 2.5A)| 1            | Or solar panel + battery       |
| Weatherproof enclosure     | 1            | Ventilated; inlet tube for PM  |
| microSD card (32 GB)       | 1            | Pre-loaded with sensor software|
| Ethernet or Wi-Fi dongle   | 1            | If Pi Zero (built-in on Pi 3+/4+) |

### 7.2 Field Deployment Supplies

| Item                       | Notes                                   |
|:---------------------------|:----------------------------------------|
| Zip ties, mounting brackets| For securing enclosures                 |
| Extension cord or solar panel | Power supply for remote sites        |
| GPS unit or smartphone     | For recording coordinates               |
| Field notebook             | Site metadata                           |
| Laptop with SSH client     | For on-site diagnostics                 |


## 8. Procedure

### 8.1 Sensor Assembly and Software Setup

1. Flash the microSD card with Raspberry Pi OS Lite.
2. Install the sensor-reading Python script (maintained in the project
   repository). The script reads the PMS5003/5004 via serial (UART)
   and the BME280 via I2C, logging timestamped readings to a CSV file
   on the SD card.
3. Configure the Pi for headless operation (enable SSH, set Wi-Fi
   credentials if applicable, set timezone to America/Los_Angeles).
4. Connect the PMS sensor ribbon cable and the BME280 module.
5. Power on the unit and verify that the sensor LED illuminates and
   the fan spins. Check the log file for valid readings.
6. Allow at least 30 seconds of warm-up before trusting data
   [@plantower2016].

### 8.2 Collocation (Pre-Deployment Calibration)

Before field deployment, collocate all sensor units at a single
location for a minimum of 48 hours:

1. Place all units within 1 m of each other, at the same height, with
   inlets facing the same direction.
2. Compare readings across units to identify outliers or malfunctioning
   sensors.
3. If a reference instrument (FRM, FEM, or BAM) is available, collocate
   sensors with it to develop a site-specific correction factor
   [@barkjohn2021].
4. Calculate inter-sensor coefficients of variation (CV). Units with
   CV >15% relative to the group median should be inspected and
   recalibrated or replaced.

### 8.3 Site Selection

- Mount sensors 2--4 m above ground level, away from direct emission
  sources (exhaust vents, barbecues, construction).
- Orient the inlet away from prevailing wind to reduce rain
  intrusion.
- Ensure access to power (AC outlet, solar panel, or battery).
- Record GPS coordinates, mounting height, distance to nearest road,
  and land use context.

### 8.4 Deployment

1. Secure the enclosure at the selected location using zip ties,
   brackets, or straps.
2. Connect power.
3. Verify operation by SSH into the Pi and checking the live log.
4. Record the deployment date/time, site ID, sensor serial number,
   and any notes in the field notebook and project metadata file.

### 8.5 Data Retrieval

Retrieve data at regular intervals (weekly to monthly, depending on
SD card capacity and power reliability):

1. SSH into the Pi and copy the CSV log file to your laptop using
   `scp` or `rsync`.
2. Alternatively, remove the microSD card and copy files directly.
3. Archive the raw file immediately; never edit the original.
4. Check the log for gaps (power outages, sensor freezes).

### 8.6 Sensor Maintenance

- Inspect the inlet for debris, spider webs, or insect nests at each
  data retrieval visit.
- Replace sensors after approximately 2 years of continuous outdoor
  operation, or sooner if collocation checks show degraded
  performance.
- Clean the exterior of the enclosure. Do not open the PMS sensor
  housing.

### 8.7 Decommissioning

1. Power off the unit.
2. Retrieve the final data file.
3. Remove the enclosure and all mounting hardware.
4. Leave no trace at the site.


## 9. Data Processing and Correction

### 9.1 Overview

Raw PM~2.5~ readings from PMS5003/5004 sensors overestimate FRM/FEM
concentrations by approximately 40% under typical US ambient conditions
[@barkjohn2021]. The EPA US-wide correction reduces root mean square
error from approximately 8 $\mu$g/m$^3$ to 3 $\mu$g/m$^3$
[@barkjohn2021].

### 9.2 Data Cleaning

```{r}
#| label: data-cleaning
#| echo: true
#| eval: false

library(tidyverse)

raw <- read_csv("data/raw/site01_20260101_20260201.csv")

# Expected columns: timestamp, pm1_cf1, pm25_cf1, pm10_cf1,
#                   pm1_atm, pm25_atm, pm10_atm,
#                   count_03, count_05, count_10, count_25, count_50, count_100,
#                   temp_c, rh_pct, pressure_hpa

# Remove negative or implausibly high values
raw <- raw %>%
  filter(
    pm25_cf1 >= 0,
    pm25_cf1 < 1000,
    rh_pct >= 0,
    rh_pct <= 100,
    temp_c > -40,
    temp_c < 60
  )

# Flag and remove sensor freeze periods (identical consecutive readings
# for > 10 minutes)
raw <- raw %>%
  mutate(frozen = pm25_cf1 == lag(pm25_cf1) &
                  pm25_cf1 == lag(pm25_cf1, 2) &
                  pm25_cf1 == lag(pm25_cf1, 3)) %>%
  filter(!frozen | is.na(frozen)) %>%
  select(-frozen)
```

### 9.3 EPA US-Wide Correction (Barkjohn et al., 2021)

The US-wide correction for PM~2.5~ under typical ambient and smoke
conditions is [@barkjohn2021]:

$$
\text{PM}_{2.5,\text{corrected}} = 0.524 \times \text{PM}_{2.5,\text{cf1}}
  - 0.0862 \times \text{RH} + 5.75
$$

where PM~2.5,cf1~ is the CF=1 output in $\mu$g/m$^3$ and RH is
relative humidity in percent.

```{r}
#| label: epa-correction
#| echo: true
#| eval: false

# Apply the Barkjohn et al. (2021) US-wide correction
raw <- raw %>%
  mutate(
    pm25_corrected = 0.524 * pm25_cf1 - 0.0862 * rh_pct + 5.75
  )

# Truncate negative corrected values to zero
raw <- raw %>%
  mutate(pm25_corrected = pmax(pm25_corrected, 0))
```

::: {.callout-important}
This correction was developed for typical ambient and smoke-impacted
conditions. It underestimates PM~2.5~ during dust-dominated events.
For dust correction, see @jaffe2023 and @ouimette2026.
:::

### 9.4 Identifying Updated (Post-2021) PMS5003 Sensors

Plantower changed the PMS5003 internal calibration in mid-2021. The
updated sensors (PMS-U) read lower than the original sensors (PMS-O)
at concentrations below approximately 16 $\mu$g/m$^3$. PMS-U sensors
can be identified by a ratio of count~0.3~/count~0.5~ that is
approximately 3 times lower than PMS-O sensors [@ardondryer2023]. If
your sensors are PMS-U, apply the additional linear correction from
@ardondryer2023 before the EPA correction.

### 9.5 Hourly and Daily Averaging

```{r}
#| label: averaging
#| echo: true
#| eval: false

# Hourly averages
hourly <- raw %>%
  mutate(hour = floor_date(timestamp, "hour")) %>%
  group_by(hour) %>%
  summarize(
    pm25_mean   = mean(pm25_corrected, na.rm = TRUE),
    pm25_sd     = sd(pm25_corrected, na.rm = TRUE),
    rh_mean     = mean(rh_pct, na.rm = TRUE),
    temp_mean   = mean(temp_c, na.rm = TRUE),
    n_obs       = n(),
    .groups     = "drop"
  ) %>%
  # Require at least 75% data completeness per hour
  filter(n_obs >= 45)  # assuming 1-min logging

# Daily averages (24-hour, midnight to midnight)
daily <- hourly %>%
  mutate(date = as_date(hour)) %>%
  group_by(date) %>%
  summarize(
    pm25_daily = mean(pm25_mean, na.rm = TRUE),
    n_hours    = n(),
    .groups    = "drop"
  ) %>%
  # Require at least 75% of hours (18 of 24)
  filter(n_hours >= 18)
```

### 9.6 AQI Calculation

```{r}
#| label: aqi-calculation
#| echo: true
#| eval: false

# Calculate US EPA AQI for PM2.5 (24-hour average)
calc_aqi <- function(pm25) {
  breakpoints <- tibble(
    lo_pm  = c(0.0, 9.1, 35.5, 55.5, 125.5, 225.5),
    hi_pm  = c(9.0, 35.4, 55.4, 125.4, 225.4, 325.4),
    lo_aqi = c(0, 51, 101, 151, 201, 301),
    hi_aqi = c(50, 100, 150, 200, 300, 500)
  )
  bp <- breakpoints %>% filter(pm25 >= lo_pm, pm25 <= hi_pm)
  if (nrow(bp) == 0) return(NA_real_)
  with(bp, lo_aqi + (hi_aqi - lo_aqi) / (hi_pm - lo_pm) * (pm25 - lo_pm))
}

daily <- daily %>%
  mutate(aqi = map_dbl(pm25_daily, calc_aqi))
```


## 10. Quality Control

### 10.1 Pre-Deployment

- All sensors pass collocation check (inter-sensor CV <15%).
- Software logging verified (valid timestamps, all columns populated).

### 10.2 During Deployment

- Weekly data check (via SSH or at retrieval) for gaps and anomalies.
- Inspect enclosure and inlet monthly.

### 10.3 Post-Processing QC Flags

| Flag | Condition                                  | Action              |
|:-----|:-------------------------------------------|:--------------------|
| F1   | RH >95% for >4 consecutive hours           | Flag; consider excluding |
| F2   | PM~2.5~ >500 $\mu$g/m$^3$ for >1 hour     | Verify against regional monitors |
| F3   | Sensor freeze (identical readings >10 min) | Exclude             |
| F4   | Hourly data completeness <75%              | Exclude hour        |
| F5   | Daily data completeness <75% of hours      | Exclude day         |

### 10.4 Periodic Collocation

Re-collocate sensors with a reference instrument (or with each other)
at least once per year and after any sensor replacement. Document
drift and update correction factors if needed.


## 11. Data Management

- Archive raw CSV files unmodified in `data/raw/` with the naming
  convention: `siteID_YYYYMMDD_YYYYMMDD_raw.csv`
- Processed/corrected files in `data/processed/`
- Metadata (site coordinates, sensor serial numbers, deployment dates,
  correction equations applied) in `data/metadata/`
- All processing scripts version-controlled in the project Git
  repository


## 12. Reporting

When reporting LCS data in publications or reports, include:

- Sensor make and model (Plantower PMS5003 or PMS5004)
- Platform (Raspberry Pi model)
- Correction equation used (cite the reference)
- Whether sensors were collocated with a reference instrument
- Data completeness and QC criteria applied
- Known limitations (RH sensitivity, dust underestimation)


## 13. Definitions

AQI
: Air Quality Index. A standardized index for reporting daily air
  quality developed by the US EPA.

BAM
: Beta Attenuation Monitor. A FEM instrument for measuring PM~2.5~.

CF=1
: "Standard particles" output from the Plantower sensor, factory
  calibrated against Chinese ambient aerosol.

FEM
: Federal Equivalent Method. An air quality measurement method
  approved by the US EPA as equivalent to the Federal Reference Method.

FRM
: Federal Reference Method. The EPA-designated reference method for
  measuring criteria pollutants.

LCS
: Low-Cost Sensor. A PM sensor costing less than approximately $250
  per unit.

PM~2.5~
: Particulate matter with aerodynamic diameter less than 2.5 $\mu$m.


## 14. References

::: {#refs}
:::


## 15. Related SOPs

| SOP   | Title                                  | Relationship                      |
|:------|:---------------------------------------|:----------------------------------|
| [F001](SOP-F001-field-safety.qmd)       | Field Safety        | Required prerequisite             |
| [F101](SOP-F101-electrical-power-in-the-field.qmd) | Electrical Power in the Field | Power supply guidance  |
| [F504](SOP-F504-picarro-field-deployment.qmd) | Picarro GHG Analyzer | Companion field instrument SOP |


## 16. Training Record

Before deploying air quality sensors independently, personnel must:

1. Read this SOP and [SOP-F001 Field Safety](SOP-F001-field-safety.qmd)
   in full
2. Demonstrate Raspberry Pi setup (OS flash, SSH, sensor script
   configuration) under supervision
3. Participate in a supervised collocation and deployment
4. Perform a data retrieval and run the correction/QC pipeline
5. Be signed off by the PI or field lead
