---
# =============================================================================
# SOP-F501: Air Temperature Monitoring (Portable Sensors)
# =============================================================================
# Deployment, calibration, and data processing for low-cost air
# temperature and relative humidity sensor networks using iButton,
# HOBO, and Raspberry Pi + BME280 platforms.
# =============================================================================

title: "Air Temperature Monitoring with Portable Sensors"
sop-number: "SOP-F501"
author:
  - name: "Marc Los Huertos"
    affiliation: "Environmental Analysis, Pomona College"
date: last-modified
date-format: "YYYY-MM-DD"
revision: "1.0"

revision-history:
  - version: "1.0"
    date: "2026-05-18"
    author: "Marc Los Huertos"
    changes: "Initial release"
---

## 1. Purpose

This SOP describes the deployment, calibration, data retrieval, and
post-processing of low-cost portable sensor networks for ambient air
temperature and relative humidity monitoring. Three sensor platforms are
covered: Maxim/Analog Devices Thermochron iButton loggers, Onset HOBO
data loggers, and Raspberry Pi single-board computers with BME280
environmental sensor modules.

Dense networks of inexpensive air temperature sensors enable studies of
urban heat island (UHI) effects, microclimatic gradients beneath
photovoltaic arrays, canopy-vs.-open temperature contrasts, and diurnal
and seasonal temperature variability at spatial resolutions not
achievable with conventional weather stations [@muller2013;
@chapman2017].


## 2. Scope

- **Applies to:** Outdoor ambient air temperature and relative humidity
  monitoring at 1.5--2.0 m above ground level using portable sensor
  networks.
- **Does not apply to:** Soil temperature profiles (see
  [SOP-F404](SOP-F404-soil-temperature-profiles.qmd)), indoor
  temperature monitoring, or air quality (PM, gases; see
  [SOP-F502](SOP-F502-air-quality-portable-sensors.qmd)).
- **Location:** EA Program field sites including the Pomona College
  campus, Bernard Field Station, photovoltaic study sites, and
  surrounding Inland Empire communities.


## 3. Responsibilities

| Role              | Responsibility                                          |
|:------------------|:--------------------------------------------------------|
| Lab PI            | Approve monitoring plan, oversee data quality           |
| Field Lead        | Coordinate site access and logistics                    |
| Trained Personnel | Calibrate, deploy, retrieve data, process results       |
| New Students      | Complete training before independent deployment         |


## 4. Safety and Hazards

::: {.callout-warning}
### Hazards

- **Working at height:** Mounting sensors at 1.5--2.0 m on poles or
  structures may require a stepladder. Follow ladder safety protocols.
- **Traffic:** Some monitoring locations are near roadways. Wear
  high-visibility vests.
- **Heat exposure:** Deployment and retrieval in summer in the Inland
  Empire involves significant heat risk. Carry water, work in pairs,
  and schedule field work for early morning when possible.
- **Electrical:** Raspberry Pi units use low-voltage USB power. Do not
  expose connections to rain or standing water without weatherproof
  enclosures.
:::

### Required PPE

- [ ] Closed-toe shoes
- [ ] High-visibility vest (if near roadways)
- [ ] Sun protection (hat, sunscreen)
- [ ] Safety glasses (during mounting/drilling)


## 5. Prerequisites

- [SOP-F001 Field Safety](SOP-F001-field-safety.qmd)
- For Raspberry Pi deployments: basic familiarity with Raspberry Pi OS,
  SSH, and file transfer (see also
  [SOP-F502](SOP-F502-air-quality-portable-sensors.qmd))


## 6. Sensor Platforms

### 6.1 Comparison

| Feature              | iButton DS1921/DS1923 | HOBO U23 Pro v2       | Raspberry Pi + BME280 |
|:---------------------|:---------------------:|:---------------------:|:---------------------:|
| Approximate cost     | $20--60               | $130--200             | $40--80               |
| Temperature accuracy | $\pm$0.5--1.0 °C      | $\pm$0.21 °C          | $\pm$0.5 °C (BME280) |
| RH measurement       | DS1923 only           | Yes                   | Yes                   |
| Logging interval     | 1 min -- 255 min      | 1 s -- 18 h           | User-defined (script) |
| Memory               | 2048--8192 readings   | ~42,000 readings      | SD card (unlimited)   |
| Communication        | 1-Wire (download)     | USB (download)        | Wi-Fi / SD card       |
| Battery life         | ~1 year               | ~1 year               | Requires continuous power |
| Size                 | 17 mm coin            | 6 x 5 x 2 cm         | Variable (enclosure)  |

### 6.2 Choosing a Platform

- **iButton:** Best for dense spatial networks where low cost and small
  size are priorities. Limited memory requires periodic retrieval.
  Ideal for short-term (weeks to months) UHI transects.
- **HOBO:** Best general-purpose option with highest accuracy, long
  battery life, and large memory. Good for season-long deployments.
- **Raspberry Pi + BME280:** Best for real-time or near-real-time
  monitoring with Wi-Fi. Requires continuous power. Already available
  in the EA Program inventory (16 Pi Zero+ units).


## 7. Radiation Shielding {#sec-shielding}

::: {.callout-important}
### Radiation Shields Are Mandatory
Unshielded sensors exposed to direct or reflected solar radiation can
read 5--15 °C higher than true air temperature. Every deployed sensor
**must** be housed in a radiation shield [@hubbart2005; @terando2017].
:::

### 7.1 Commercial Shields

- **Onset RS3:** Louvered, naturally ventilated, white plastic. Fits
  HOBO loggers. Proven performance in the literature [@hubbart2005].
- **Onset M-RSA:** Aspirated shield (fan-ventilated). Highest accuracy
  but requires power.
- **Davis 7714:** Louvered shield compatible with various loggers.

### 7.2 Custom Shields

For iButton deployments where cost must be minimized, fabricate shields
from white plastic funnels, stacked white plastic plates, or PVC
fittings. Key design principles [@terando2017; @hubbart2005]:

- Use white, UV-resistant material to reflect solar radiation.
- Allow natural ventilation from below while blocking direct and
  reflected shortwave radiation from above.
- Orient the sensor opening downward to prevent rain entry.
- Test custom shields against a commercial shield before deployment.

### 7.3 Mounting

Mount radiation shields at 1.5--2.0 m above ground on a pole, post, or
structure. Avoid mounting directly against walls, fences, or dark
surfaces that radiate heat. Secure with zip ties, hose clamps, or
U-bolts.


## 8. Calibration {#sec-calibration}

Calibrate all sensors before each deployment season and after any
suspected damage.

### 8.1 Ice-Water Bath Check

1. Prepare a well-stirred ice-water bath (crushed ice + deionized water
   in an insulated container). This produces 0.0 °C.
2. Submerge all sensors simultaneously. Wait at least 5 minutes for
   equilibration.
3. Record the reading from each sensor.
4. Sensors should read within $\pm$0.5 °C of 0.0 °C. Document offsets.

### 8.2 Room-Temperature Inter-Comparison

1. Place all sensors in a well-mixed, stable indoor environment
   (e.g., a lab bench away from HVAC vents) alongside a NIST-traceable
   reference thermometer.
2. Log for at least 2 hours.
3. Calculate the mean offset of each sensor relative to the reference.
4. Sensors with offsets >1.0 °C should be replaced or have a correction
   applied in post-processing.

### 8.3 Collocation in the Field

For highest accuracy, collocate 2--3 sensors at one site for the first
48 hours of deployment to confirm agreement under field conditions
(including radiation shield effects). All sensors should agree within
0.5 °C at night (when radiation bias is absent).

```{r}
#| label: calibration-offsets
#| echo: true
#| eval: false

library(tidyverse)

# Read calibration bath data
cal <- read_csv("data/calibration/ice_bath_YYYYMMDD.csv")

# Calculate offsets from 0.0 C
cal_summary <- cal %>%
  group_by(sensor_id) %>%
  summarize(
    mean_reading = mean(temp_c, na.rm = TRUE),
    offset       = mean_reading - 0.0,
    sd           = sd(temp_c, na.rm = TRUE),
    .groups      = "drop"
  )

# Flag sensors outside tolerance
cal_summary <- cal_summary %>%
  mutate(status = if_else(abs(offset) <= 0.5, "PASS", "REVIEW"))

print(cal_summary)
write_csv(cal_summary, "data/calibration/offsets_YYYYMMDD.csv")
```


## 9. Deployment Procedure

### 9.1 iButton

1. Program the iButton using OneWireViewer or equivalent software:
   set logging interval (typically 10--15 minutes), delayed start
   time, and enable rollover if the study will exceed memory capacity.
2. Place the iButton in its radiation shield (see @sec-shielding).
3. Mount the shield at the site at 1.5--2.0 m height.
4. Record: site ID, sensor serial number, GPS coordinates, mounting
   height, shield type, start time, land use/land cover description.

### 9.2 HOBO U23

1. Launch the logger in HOBOware: set logging interval, delayed start,
   and file name.
2. Attach the logger to the RS3 radiation shield.
3. Mount the shield at the site.
4. Record metadata as above.

### 9.3 Raspberry Pi + BME280

1. Verify the BME280 is reading correctly (SSH in, check the log).
2. Mount the Pi enclosure with the BME280 sensor exposed inside the
   radiation shield (sensor element facing downward for ventilation).
3. Connect power (AC, solar + battery, or long USB cable).
4. Verify data logging remotely.
5. Record metadata as above.


## 10. Data Retrieval

| Platform    | Retrieval Method                          | Frequency           |
|:------------|:------------------------------------------|:--------------------|
| iButton     | Remove, connect to 1-Wire reader, download in OneWireViewer | Every 2--4 weeks |
| HOBO U23    | Connect USB shuttle or cable, download in HOBOware | Every 1--3 months |
| Raspberry Pi| SSH + `scp`, or remove SD card             | Weekly or as needed |

After retrieval, archive the raw file immediately. Never modify the
original.


## 11. Data Processing

### 11.1 Import and Standardize

```{r}
#| label: import-standardize
#| echo: true
#| eval: false

library(tidyverse)
library(lubridate)

# Example: import iButton CSV
raw <- read_csv("data/raw/site01_ibutton_ABC123.csv",
                skip = 14) %>%     # skip header lines (varies by software)
  rename(timestamp = 1, temp_c = 2) %>%
  mutate(
    timestamp = mdy_hms(timestamp, tz = "America/Los_Angeles"),
    sensor_id = "ABC123",
    site_id   = "site01",
    platform  = "iButton"
  )

# Apply calibration offset (from Section 8)
offsets <- read_csv("data/calibration/offsets_YYYYMMDD.csv")

raw <- raw %>%
  left_join(offsets %>% select(sensor_id, offset), by = "sensor_id") %>%
  mutate(temp_corrected = temp_c - offset)
```

### 11.2 Quality Control

```{r}
#| label: qc-flags
#| echo: true
#| eval: false

raw <- raw %>%
  mutate(
    qc_flag = case_when(
      temp_corrected < -10 | temp_corrected > 55 ~ "out_of_range",
      temp_corrected == lag(temp_corrected) &
        temp_corrected == lag(temp_corrected, 2) &
        temp_corrected == lag(temp_corrected, 3)   ~ "stuck_sensor",
      abs(temp_corrected - lag(temp_corrected)) > 10 ~ "spike",
      TRUE ~ "valid"
    )
  )

# Summary
count(raw, qc_flag)

# Keep only valid data for analysis
clean <- raw %>% filter(qc_flag == "valid")
```

### 11.3 Hourly and Daily Summaries

```{r}
#| label: summaries
#| echo: true
#| eval: false

hourly <- clean %>%
  mutate(hour = floor_date(timestamp, "hour")) %>%
  group_by(site_id, sensor_id, hour) %>%
  summarize(
    temp_mean = mean(temp_corrected, na.rm = TRUE),
    temp_min  = min(temp_corrected, na.rm = TRUE),
    temp_max  = max(temp_corrected, na.rm = TRUE),
    n_obs     = n(),
    .groups   = "drop"
  ) %>%
  filter(n_obs >= 3)  # at least 3 of 4 readings per hour (15-min interval)

daily <- hourly %>%
  mutate(date = as_date(hour)) %>%
  group_by(site_id, sensor_id, date) %>%
  summarize(
    t_mean = mean(temp_mean, na.rm = TRUE),
    t_min  = min(temp_min, na.rm = TRUE),
    t_max  = max(temp_max, na.rm = TRUE),
    n_hours = n(),
    .groups = "drop"
  ) %>%
  filter(n_hours >= 18)  # 75% completeness
```

### 11.4 Urban Heat Island Intensity

UHI intensity is typically calculated as the difference between an
urban site and a rural (or least-developed) reference site:

$$
\Delta T_{\text{UHI}} = T_{\text{urban}} - T_{\text{rural}}
$$

```{r}
#| label: uhi-intensity
#| echo: true
#| eval: false

# Define the reference (rural) site
ref_site <- "site_rural01"

uhi <- hourly %>%
  select(site_id, hour, temp_mean) %>%
  pivot_wider(names_from = site_id, values_from = temp_mean) %>%
  mutate(across(-c(hour, all_of(ref_site)),
                ~ . - .data[[ref_site]],
                .names = "uhi_{.col}"))
```


## 12. Quality Control Summary

| Check                                | Criterion                        |
|:--------------------------------------|:---------------------------------|
| Ice-bath calibration                  | Reading within $\pm$0.5 °C of 0.0 °C |
| Inter-sensor agreement (collocation) | Within $\pm$0.5 °C at night      |
| Hourly data completeness             | $\geq$75% of expected readings   |
| Daily data completeness              | $\geq$18 of 24 hours             |
| Stuck-sensor detection               | $\geq$4 identical consecutive readings flagged |
| Spike detection                      | $\geq$10 °C change between consecutive readings flagged |


## 13. Data Management

- Raw files archived unmodified: `data/raw/siteID_platform_sensorID.csv`
- Calibration records: `data/calibration/`
- Processed files: `data/processed/`
- Site metadata (coordinates, height, shield type, land cover, sensor
  serial numbers, deployment/retrieval dates):
  `data/metadata/site_metadata.csv`
- All scripts version-controlled in the project Git repository


## 14. Definitions

BME280
: Bosch environmental sensor measuring temperature, relative humidity,
  and barometric pressure. Used with Raspberry Pi platforms.

iButton
: A coin-sized temperature (and optionally RH) data logger
  manufactured by Maxim/Analog Devices (Thermochron series).

HOBO
: A family of portable data loggers manufactured by Onset Computer
  Corporation.

Radiation shield
: A louvered or ventilated housing that protects a temperature sensor
  from direct and reflected solar radiation while allowing ambient air
  to flow past the sensor.

Urban Heat Island (UHI)
: The phenomenon whereby urban areas experience higher air temperatures
  than surrounding rural areas due to impervious surfaces, waste heat,
  reduced vegetation, and altered energy balance.


## 15. References

::: {#refs}
:::


## 16. Related SOPs

| SOP   | Title                                  | Relationship                      |
|:------|:---------------------------------------|:----------------------------------|
| [F001](SOP-F001-field-safety.qmd)       | Field Safety        | Required prerequisite             |
| [F101](SOP-F101-electrical-power-in-the-field.qmd) | Electrical Power in the Field | Power for Pi deployments |
| [F404](SOP-F404-soil-temperature-profiles.qmd) | Soil Temperature Profiles | Companion: soil vs. air temp |
| [F502](SOP-F==502-air-quality-portable-sensors.qmd) | Air Quality Monitoring | Companion: PM sensors on same Pi units |


## 17. Training Record

Before deploying air temperature sensors independently, personnel must:

1. Read this SOP and [SOP-F001 Field Safety](SOP-F001-field-safety.qmd)
   in full
2. Perform a calibration (ice-water bath and room-temperature
   inter-comparison) under supervision
3. Assemble and mount a radiation shield
4. Deploy and retrieve a sensor at a test site
5. Import, clean, and summarize data using the R workflow
6. Be signed off by the PI or field lead
