---
title: "iNaturalist for Biodiversity Documentation"
sop-number: "SOP-F602"
author:
  - name: "Marc Los Huertos"
    affiliation: "EA Program, Pomona College"
date: last-modified
date-format: "YYYY-MM-DD"
revision: "0.3"

revision-history:
  - version: "0.1"
    date: "2025-05-17"
    author: "Marc Los Huertos"
    changes: "Initial release"
  - version: "0.2"
    date: "2026-05-28"
    author: "Marc Los Huertos"
    changes: "Added bibtex citations for community science data quality, GBIF, and observation best practices"
  - version: "0.3"
    date: "2026-05-28"
    author: "Marc Los Huertos"
    changes: "Added definitions section; expanded data quality discussion; added citations for large-scale monitoring and CV model limitations"
---

## Purpose

This SOP describes how to use iNaturalist
(<https://www.inaturalist.org>) as a standardized tool for documenting
biodiversity at EA Program field sites. iNaturalist is a community
science platform that combines field observation, AI-assisted species
identification, and expert community verification into a single
workflow [@ueda2024]. Observations that receive community agreement on
identification reach "Research Grade" status and are automatically
shared with the Global Biodiversity Information Facility (GBIF),
making them part of the global scientific record [@gbif2024].

Community science platforms like iNaturalist generate large volumes of
occurrence data, but the scientific value of those data depends
critically on the quality of individual observations, including
photographic evidence, spatial accuracy, and taxonomic verification
[@dickinson2010]. When data quality is maintained, iNaturalist records
can support large-scale biodiversity monitoring and contribute
meaningfully to national and international biodiversity assessments
[@callaghan2021; @chandler2017]. This SOP covers making high-quality
observations, creating and managing collaborative projects for EA
Program research, and exporting data for analysis in R.


## Scope

- **Applies to:** All EA Program biodiversity surveys, species
  inventories, and community science activities where iNaturalist is
  used to document plant, animal, fungal, or other organism
  occurrences
- **Does not apply to:** Quantitative vegetation transect surveys
  ([SOP-F603](SOP-F603-vegetation-transect-surveys.qmd)); camera trap
  image processing ([SOP-F620](SOP-F620-camera-trap.qmd)); formal
  specimen collection requiring voucher preparation; eDNA sampling
- **Location:** All EA Program field sites and campus


## Responsibilities

| Role              | Responsibility                                   |
|:------------------|:-------------------------------------------------|
| PI / Field Lead   | Create and administer iNaturalist projects; review data quality |
| Project Admin     | Manage project settings, curate observations, invite members |
| Trained Personnel | Make observations per this SOP; contribute identifications |
| New Students      | Create an account and complete training before using iNaturalist for research data |


## Definitions

- **Observation:** A record of an encounter with an individual organism
  at a specific place and time, documented with at least one photo or
  sound recording.
- **Research Grade:** An observation that has a date, location, photo
  or sound, and community agreement on the identification (at least
  two-thirds of identifiers agree at the species level). Research
  Grade observations are shared with GBIF [@gbif2024].
- **Needs ID:** An observation that has not yet received enough
  community identifications to reach Research Grade. These records are
  scientifically useful but carry higher taxonomic uncertainty.
- **Casual:** An observation that is missing date, location, or
  evidence (photo/sound), or that the community has voted as not
  meeting quality standards. Casual observations are excluded from
  most research exports.
- **Collection Project:** A project that automatically includes all
  observations matching defined criteria (taxa, place, users, dates,
  quality grade). No manual curation needed.
- **Traditional Project:** A project where observations must be added
  manually by members. Can require custom observation fields. Requires
  the creator to have at least 50 verifiable observations.
- **Umbrella Project:** A project that aggregates observations from
  multiple other projects. Useful for coordinating across sites or
  taxa.
- **Observation Field:** A custom data field that can be attached to
  observations (e.g., "habitat type," "number of individuals,"
  "flowering status"). Only Traditional Projects can require
  observation fields.
- **Computer Vision (CV) Model:** The machine learning model trained
  on iNaturalist's community-verified image database that suggests
  identifications for new observations. The model is retrained
  periodically and performs best for taxa and regions with large
  training datasets; accuracy declines for rare, under-photographed,
  or morphologically cryptic taxa [@loarie2023].
- **Taxon Geoprivacy:** An automatic system that obscures the
  coordinates of observations for threatened, endangered, or
  sensitive species to protect them from collection or disturbance.
- **Coordinate Uncertainty:** The estimated spatial accuracy of the
  GPS location, expressed as a radius in meters. High uncertainty
  (greater than 1000 m) reduces the value of occurrence records for
  fine-scale spatial analyses.


## Safety and Hazards

Standard field safety applies when making observations in the field.
Refer to [SOP-F001 Field Safety](SOP-F001-field-safety.qmd). Additional
considerations specific to iNaturalist use:

- **Location privacy:** iNaturalist observations include GPS
  coordinates by default. Be aware that your location is visible to
  other users unless you obscure or hide it. Do not post observations
  from private residences without obscuring the location.
- **Sensitive species:** iNaturalist automatically obscures the
  locations of threatened and endangered species ("taxon geoprivacy").
  Do not override this obscuring for sensitive taxa.
- **Disturbance:** Do not disturb, handle, or collect organisms solely
  for an iNaturalist photo. Photograph organisms in situ where
  possible.


## Required Materials

| Item                                     | Notes                              |
|:-----------------------------------------|:-----------------------------------|
| Smartphone with iNaturalist app          | iOS or Android; free download      |
| iNaturalist account                      | Free at <https://www.inaturalist.org> |
| Camera (optional)                        | For higher-quality photos; phone camera is sufficient for most uses |
| Field notebook                           | For recording additional data not captured in iNaturalist |
| GPS unit (optional)                      | Phone GPS is usually adequate      |


## Procedure

### Account Setup

1. Go to <https://www.inaturalist.org> or download the iNaturalist app
2. Create an account using your Claremont email address
3. In your profile settings:
   - Set your default photo license to **CC-BY** or **CC-BY-NC**
     (this allows your photos to be used in research with attribution)
   - Set your default observation license to **CC0** or **CC-BY**
     (this allows your observation data to be shared with GBIF)
   - Set your default location to "open" (you can obscure individual
     observations as needed)
4. Send your iNaturalist username to the project administrator so you
   can be added to the EA Program project

::: {.callout-important}
### License Settings Matter for Research
Observations with "All Rights Reserved" licenses are excluded from
GBIF exports and many research datasets. For EA Program work, always
use a Creative Commons license. CC-BY is recommended.
:::

### Making High-Quality Observations

A useful iNaturalist observation needs four things: a photo (or sound),
a date, a location, and enough visual detail for identification.
Follow these guidelines to maximize the chances of reaching Research
Grade.

#### Photography Tips

1. Take multiple photos of each organism from different angles:
   - **Plants:** Whole plant habit, close-up of leaves (both sides),
     flowers or fruit, stem, bark (for trees)
   - **Insects:** Dorsal view (top), lateral view (side), close-up of
     head and wing venation if possible
   - **Birds:** Best angle available; include habitat context if
     possible
   - **Fungi:** Top of cap, underside (gills/pores), stem, habitat
     context
2. Include something for scale (a finger, a coin, a ruler) when
   practical
3. Photograph the organism in its natural position; do not relocate it
4. For plants, photograph vegetative and reproductive structures
   together if both are present
5. Ensure the photo is in focus and adequately lit

#### Recording the Observation

6. Open the iNaturalist app and tap the camera icon to start a new
   observation
7. Take photos directly in the app, or select photos from your camera
   roll
8. The app will suggest an identification using its computer vision
   (CV) model [@loarie2023]. Accept the suggestion if it looks
   correct, or refine the identification manually. When in doubt,
   identify to the broadest level you are confident in (e.g., "Plants"
   or "Oaks" rather than guessing a specific species). The CV model
   performs best for common, well-photographed taxa in North America
   and Europe; for under-represented groups (e.g., lichens, mosses,
   many invertebrate families), manual identification by the observer
   or community identifiers is essential.
9. Verify that the date and location are correct. The app captures
   these automatically from the photo metadata, but check that the GPS
   pin is accurate on the map. If coordinate uncertainty exceeds 500 m,
   manually reposition the pin.
10. Add any relevant notes in the "Notes" field (habitat, behavior,
    abundance, associated species)
11. If the observation is part of a Traditional Project, add it to the
    project before submitting (see Section 7.4)
12. Tap **Submit**

::: {.callout-tip}
### Sound Observations
For birds, frogs, and insects, a sound recording can be as useful as
(or more useful than) a photo. The iNaturalist app can record audio
directly. Hold the phone steady and minimize background noise.
:::

### Bulk Uploading Observations via CSV

For observations recorded outside the app (for example, transcribed
from a field notebook, exported from another platform, or compiled in
a spreadsheet during data entry), iNaturalist accepts a bulk CSV
import. This is useful when you have many records to create at once and
do not want to enter them one at a time. The importer creates the
observation records only; it does not upload photos or sound files, so
media must be attached afterward (see below).

#### Preparing the CSV

1. Build the spreadsheet with up to eight columns, in this order. Only
   the taxon column is mandatory.

| Column        | Format / example          | Notes                                      |
|:--------------|:--------------------------|:-------------------------------------------|
| Taxon name    | text (e.g., `Quercus agrifolia`) | Mandatory. Use the scientific name for reliable matching. |
| Date observed | `YYYY-MM-DD HH:MM`        | Use day-level precision or finer for research use. |
| Description   | text                      | Free-text notes (habitat, behavior, abundance). |
| Place name    | text                      | Optional named locality.                   |
| Latitude      | decimal degrees (`34.1064`) | Use a decimal point, not a comma.        |
| Longitude     | decimal degrees (`-117.7120`) | Use a decimal point, not a comma.      |
| Tags          | `tag,tag`                 | Comma-separated tags.                      |
| Geoprivacy    | `open`, `obscured`, or `private` | Set to `obscured` for sensitive taxa. |

2. Keep the file under 10,000 rows. Split larger datasets into
   multiple files.
3. Wrap any field that contains a comma in double quotes so it is not
   misread as a column break.
4. Confirm that coordinates use decimal points. Spreadsheets configured
   for some locales export decimal commas, which the importer rejects.

#### Importing

1. Go to the import page at
   <https://www.inaturalist.org/observations/import#csv_import> (or
   from the upload page, click **More import options** and choose
   **CSV**).
2. Choose your file and upload it.
3. Open **Edit Observations** to review the created records. Use
   **Batch Edit** to select multiple observations, attach photos or
   sound files, and apply a shared location from the map.
4. Verify dates, coordinates, and identifications, then save.

::: {.callout-warning}
### CSV Imports Do Not Include Media
The CSV importer creates observation records without photos or sounds.
An observation with no media cannot reach Research Grade and is treated
as Casual, so it will be excluded from most research exports and from
GBIF. Attach media through Batch Edit after import if the records are
intended for analysis.
:::

::: {.callout-tip}
### Large or Photo-Linked Datasets: Use the API
For large batches, or when you need to match many photos to the
correct records automatically, the spreadsheet importer becomes
impractical. In those cases, script the upload against the iNaturalist
API (<https://api.inaturalist.org>). The `rinat` package is read-only
(it retrieves data but does not create observations), so creating
observations programmatically requires direct authenticated calls,
for example with `httr2` in R:

```{r}
#| label: inat-api-upload
#| echo: true
#| eval: false

library(httr2)

# Requires an OAuth API token from your iNaturalist account.
# Create the observation, then POST the photo to /observation_photos
# using the returned observation id. See the API reference for the
# full request schema before running against your account.
```

Test any upload script on a handful of records first, and never use it
to create observations for organisms you did not personally encounter
(see Responsible Use).
:::

### Contributing Identifications

Identifying other people's observations is as valuable as making your
own. It builds community expertise, helps observations reach Research
Grade, and improves the CV model.

1. Go to the **Identify** tab on the website (or use the Explore page
   and filter by "Needs ID")
2. Filter by your area of expertise (taxa, location)
3. Add your identification. Agree with the existing suggestion if
   correct, or provide a different identification with a brief
   explanation.
4. Use the "Compare" tool to check similar species
5. Mark observations as "No, it's as good as it can be" only if the
   evidence is genuinely insufficient for identification

### Creating and Managing Projects

iNaturalist offers three project types. Choose based on your research
needs:

| Feature                          | Collection Project | Traditional Project | Umbrella Project |
|:---------------------------------|:-------------------|:--------------------|:-----------------|
| Observations added               | Automatically (by filter criteria) | Manually by members | From member projects |
| Custom observation fields         | No                 | Yes (can be required) | No              |
| Creator requirements             | Any user           | 50+ verifiable observations | Any user      |
| Best for                         | Site-based inventories, class bioblitzes | Projects needing custom data fields | Multi-site or multi-taxon coordination |
| Obscured coordinate access       | Yes (with Trust enabled) | Yes (with member permission) | No (via member projects) |

#### Creating a Collection Project

This is the recommended type for most EA Program site inventories and
class projects. Observations matching your criteria are included
automatically.

1. Go to <https://www.inaturalist.org/projects/new>
2. Click **Get Started** under Collection Projects
3. Fill in the project details:
   - **Title:** Use a descriptive name (e.g., "EA Program -- Bernard
     Field Station Biodiversity 2025")
   - **Icon:** Upload the EA Program logo or a representative photo
   - **Banner:** Upload a site photo
   - **Description:** Explain the project's purpose, study site, and
     time frame
4. Set the project requirements (filters):
   - **Place:** Select the iNaturalist Place for your study area. If
     one does not exist, create a custom Place by drawing a boundary
     on the map (Community > Places > Add a New Place).
   - **Taxa:** Leave blank to include all organisms, or restrict to
     specific groups (e.g., Plantae, Aves)
   - **Users:** Leave blank to include anyone, or list specific
     usernames for a closed team
   - **Date range:** Set start and end dates if the project covers a
     specific period (e.g., a class semester or field season)
   - **Quality grade:** Select "Research Grade" and "Needs ID" for
     most research projects
5. Under **Trust** (optional but recommended for research): enable
   Trust to allow project admins to access the true coordinates of
   observations with obscured or private locations. Members must
   individually opt in to share their coordinates.
6. Click **Save**

::: {.callout-note}
### Creating a Custom Place
If your study area (e.g., the Bernard Field Station, a specific
watershed) does not exist as an iNaturalist Place, create one:

1. Go to <https://www.inaturalist.org/places/new>
2. Name the place (e.g., "Bernard Field Station, Claremont, CA")
3. Draw the boundary on the map using the polygon tool
4. Save

The place will be available as a filter in your Collection Project.
Note that very small places (e.g., a single backyard) work better
with Traditional Projects because GPS accuracy and obscured
coordinates can cause observations to fall outside tiny boundaries.
:::

#### Creating a Traditional Project

Use when you need custom observation fields or when the study area is
too small for reliable GPS-based filtering.

1. You must have at least 50 verifiable observations on your account
2. Go to <https://www.inaturalist.org/projects/new>
3. Look for the small text link at the bottom of the page: "click here
   to create a traditional project"
4. Fill in the project details (title, icon, banner, description)
5. Set the membership model:
   - **Open:** Anyone can join
   - **Invite-only:** Only users you invite can join (recommended for
     research teams)
6. Set the submission model:
   - **Any member:** All members can add observations
   - **Project curators:** Only curators can add observations
     (recommended when the PI needs to control data quality)
7. Add observation fields if needed (e.g., "Habitat type," "Transect
   number," "Number of individuals").


## Data Quality Considerations

iNaturalist data are heterogeneous in quality because they are
generated by observers with widely varying skill levels, equipment,
and motivations [@dickinson2010]. For EA Program research, apply the
following data quality filters when exporting records for analysis:

- **Quality grade:** Use Research Grade observations only for
  quantitative analyses. Needs ID observations can supplement species
  inventories but carry higher misidentification risk.
- **Coordinate uncertainty:** Exclude records with positional accuracy
  greater than 1000 m unless the analysis does not require fine
  spatial resolution.
- **Date precision:** Exclude records without day-level date
  precision for phenological analyses.
- **Taxonomic verification:** For critical analyses, manually verify
  identifications against the photographic evidence, especially for
  taxa where the CV model is known to struggle (e.g., grasses,
  juvenile insects, sterile plants).
- **Spatial bias:** iNaturalist data are spatially biased toward
  accessible, human-frequented areas (trails, roads, urban parks).
  This bias must be acknowledged and, where possible, corrected for
  in species distribution models [@callaghan2021].

These filters should be documented in the methods section of any
publication or report using iNaturalist-derived data.


## Data Export and Analysis

### Exporting from iNaturalist

1. Go to the **Explore** page and filter to your project, place, or
   taxa of interest
2. Click **Filters** to narrow results
3. Click **Download** (bottom of the page) to download a CSV file

### Exporting from GBIF

For Research Grade observations that have been shared with GBIF:

1. Go to <https://www.gbif.org/occurrence/search>
2. Filter by dataset (search for "iNaturalist"), location, and taxa
3. Click **Download** and select the desired format
4. GBIF assigns a DOI to every download; cite the DOI in your
   publications

### Using `rinat` in R

The `rinat` package provides programmatic access to iNaturalist data
[@barve2025]:

```{r}
#| label: rinat-example
#| echo: true
#| eval: false

library(rinat)

# Get observations from a specific project
obs <- get_inat_obs_project("ea-program-bfs-biodiversity-2025",
                            type = "observations")

# Get observations by taxon and place
birds <- get_inat_obs(taxon_name = "Aves",
                      place_id = 12345,
                      quality = "research",
                      maxresults = 500)
```


## Citing iNaturalist Data

- **Platform citation:** iNaturalist. Available at:
  <https://www.inaturalist.org>. Accessed [date].
- **GBIF download citation:** Use the DOI provided with the GBIF
  download.
- **Individual observations:** Cite by observation URL (e.g.,
  <https://www.inaturalist.org/observations/123456>) when referencing
  specific records.

When using iNaturalist data in publications, describe the data
quality filters applied (e.g., Research Grade only, coordinate
uncertainty threshold) and acknowledge the community identifiers whose
expertise validated the records [@dickinson2010; @chandler2017].


## Responsible Use

- Respect copyright: each user retains ownership of their observations
  and photos. Your rights to use them are governed by the license the
  user selected. Most iNaturalist photos use CC-BY-NC.
- Do not create observations for specimens you did not personally
  encounter (no uploading photos from field guides, the internet, or
  other people's cameras)
- Do not harass, collect, or relocate organisms for the purpose of
  getting a better photo
- For sensitive species (rare plants, nesting sites), manually obscure
  the location even if iNaturalist does not do so automatically
- Follow iNaturalist's Community Guidelines:
  <https://www.inaturalist.org/pages/community+guidelines>

::: {.callout-warning}
### Student Use
Before assigning students to use iNaturalist, read the iNaturalist
Teacher's Guide
(<https://www.inaturalist.org/pages/teacher's+guide>). The majority
of low-quality observations and inappropriate content on iNaturalist
comes from poorly managed class assignments. Set clear expectations,
demonstrate proper technique, and monitor student contributions.
Structured class projects with specific data quality goals produce
better scientific outcomes and learning [@aristeidou2021].
:::


## References

::: {#refs}
:::


## Training Record

Before using iNaturalist for EA Program research data, personnel must:

1. Read this SOP in full
2. Create an iNaturalist account with appropriate license settings
   (Section 7.1)
3. Make at least 10 observations that reach Research Grade (or "Needs
   ID" with correct identifications added by the community)
4. Contribute at least 10 identifications to other users' observations
5. Demonstrate familiarity with the project type relevant to their
   research (Collection or Traditional)
6. Be signed off by the PI or project lead
