---
# =============================================================================
# SOP-F005: Random Sampling Design
# =============================================================================
title: "Random Sampling Design"
sop-number: "SOP-F005"
author:
  - name: "Marc Los Huertos"
    affiliation: "EA Program, Pomona College"
date: last-modified
date-format: "YYYY-MM-DD"
revision: "1.0"

revision-history:
  - version: "0.1"
    date: "2017-08-18"
    author: "Marc Los Huertos"
    changes: "Initial release (LaTeX/knitr)"
  - version: "1.0"
    date: "2025-05-17"
    author: "Marc Los Huertos"
    changes: "Converted to Quarto; expanded procedure sections for simple random, grid-based, and stratified random sampling"
---

## Purpose

This SOP provides standardized procedures for designing and executing
random sampling in field and laboratory settings. Random sampling
ensures that every unit in the target population has an equal and
independent probability of selection, eliminating researcher bias and
supporting valid statistical inference.


## Scope

- **Applies to:** Any EA Program study where population estimates are
  the goal, including vegetation surveys, soil sampling, water quality
  monitoring, and laboratory subsampling
- **Does not apply to:** Targeted or judgmental sampling designs (e.g.,
  hot-spot investigations); mark-release-recapture studies for mobile
  organisms (see separate SOP)
- **Sampling types covered:** Simple random sampling, grid-based random
  sampling, and stratified random sampling


## Responsibilities

| Role              | Responsibility                              |
|:------------------|:--------------------------------------------|
| PI / Field Lead   | Approve sampling design, review sampling plan |
| Trained Personnel | Execute random selection, collect and record data |
| New Students      | Complete training before independent fieldwork |


## Safety and Hazards

::: {.callout-warning}
### Hazards

Field sampling involves site-specific hazards including uneven terrain,
heat exposure, wildlife encounters, and sharp equipment (stakes, pins).
Review site conditions before each sampling event.
:::

### Required PPE

- [ ] Closed-toe shoes or boots
- [ ] Sun protection (hat, sunscreen)
- [ ] Gloves (when handling equipment or samples)

### Emergency Procedures

Refer to [SOP-F001 Field Safety](SOP-F001-field-safety.qmd). Carry a
first aid kit and charged cell phone to all field sites.


## Definitions

- **Sampling universe (population):** The complete set of units from
  which a sample will be drawn.
- **Sampling frame:** A numbered list or spatial grid that enumerates
  every unit in the population. Missing units have zero probability of
  selection and introduce bias.
- **Quadrat:** A square frame of known area (typically 0.25, 0.5, or
  1.0 m^2^) used to delineate sampling units in the field.
- **Stratum:** A subdivision of the sampling universe into internally
  homogeneous zones (e.g., by elevation, vegetation type, or soil
  moisture gradient).


## Required Materials

| Item                                          | Notes                              |
|:----------------------------------------------|:-----------------------------------|
| Two survey tape measures (20--50 m)           | For establishing coordinate grids  |
| Ground pins or stakes                         | To anchor tape measures            |
| Quadrat frame(s)                              | Size appropriate to target organisms |
| Random number table or generator app          | See Section 8.1                    |
| GPS unit                                      | To record sampling locations       |
| Field data sheets and waterproof markers      | Pre-printed with site and date fields |
| Compass                                       | For establishing grid orientation  |
| Field identification guides                   | If sampling vegetation or organisms |


## Biases and Interferences

Representative sampling requires strict adherence to randomized
procedures. Common sources of bias include:

- **Incomplete sampling frames.** Any unit missing from the frame has
  zero probability of selection.
- **Haphazard placement.** Never "throw" a quadrat to choose a
  location. Human throws cluster toward the center of the study area,
  violating equal probability assumptions.
- **Environmental heterogeneity.** If the study area contains distinct
  gradients (e.g., wet to dry, shaded to open), simple random sampling
  may under-represent one end of the gradient. Use stratified random
  sampling instead (Section 8.4).
- **Edge effects.** Establish and apply a consistent border rule when
  counting organisms in quadrats (Section 8.3).


## Procedure

### Generating Random Numbers

Use a software-based random number generator to produce unbiased
selections. Never rely on manual or mental number selection.

In R, generate a set of random integers within a defined range:

```{r}
#| label: random-numbers
#| echo: true

# Generate 15 random integers between 1 and 100
set.seed(NULL)  # ensure a fresh seed each session
random_vals <- sample(1:100, size = 15, replace = FALSE)
sort(random_vals)
```

For field use without a computer, a printed random number table is
acceptable. Begin at a pre-selected (arbitrary) starting position and
read values sequentially. @tbl-randomnumbers provides a reference table
of 200 random numbers between 0 and 1000.

```{r}
#| label: tbl-randomnumbers
#| tbl-cap: "Reference table of 200 random numbers (0--1000)"
#| echo: false

set.seed(2017)
rand_mat <- matrix(round(runif(200, min = 0, max = 1000), 1),
                   nrow = 20, ncol = 10)
colnames(rand_mat) <- 1:10
knitr::kable(rand_mat)
```

### Simple Random Sampling

Use this approach when the population can be enumerated as a list
(e.g., lab samples, batch numbers, individual organisms in a database).

1. Define the target population and compile a complete, numbered
   sampling frame
2. Determine the required sample size ($n$) using a power analysis or
   standard formula
3. Use `sample()` in R (or an equivalent tool) to draw $n$ unit numbers
   from the frame without replacement
4. Document the random seed (if reproducibility is desired), the
   selected unit IDs, and the date of selection
5. Collect the data or specimens from the selected units

### Grid-Based Random Sampling (Field)

Use this approach when the sampling universe is a spatially defined
area (e.g., a meadow, forest plot, intertidal zone). This is the
standard method for quadrat-based ecological sampling.

#### Establishing the Baseline Grid

1. Locate the boundaries of the study site and select a level,
   representative area
2. Drive a stake at the origin point and lay two tape measures at 90
   degrees to create X and Y axes
3. Secure the tapes with ground pins to define a coordinate grid (e.g.,
   20 m $\times$ 20 m)

#### Generating Coordinates and Placing Quadrats

4. Generate pairs of random integers within the range of each axis. In
   R:

```{r}
#| label: xy-random
#| echo: true

# Grid dimensions
x_max <- 20
y_max <- 20
n_quadrats <- 15

# Generate random coordinates
coords <- data.frame(
  X = sample(0:x_max, size = n_quadrats, replace = TRUE),
  Y = sample(0:y_max, size = n_quadrats, replace = TRUE)
)

# Sort by X to create an efficient walking order
coords <- coords[order(coords$X), ]
coords$QuadratID <- seq_len(nrow(coords))
knitr::kable(coords, row.names = FALSE)
```

5. Walk along the axes to each coordinate intersection and place the
   bottom-left corner of the quadrat at that point
6. Never reposition a quadrat after placement; if a coordinate falls on
   an obstruction, record the obstruction and sample as-is

#### Recording Data Within Quadrats

7. Quantify the target variable using one of three standardized
   metrics:

   - **Density:** Count every individual of the target species within
     the frame
   - **Percent cover:** Estimate the percentage of ground area covered
     by the target species (suited for colonial plants, mosses, or
     ground cover)
   - **Frequency:** Score presence (1) or absence (0) of the species

8. Apply a consistent **border rule** to avoid double-counting:
   count organisms touching the top and right edges of the quadrat;
   exclude those touching the bottom and left edges.

#### Replication and Extrapolation

9. Aim to sample at least 5--10% of the total grid area, or a minimum
   of 30 quadrat placements, to achieve adequate statistical power
10. Calculate the mean count (or cover) per quadrat and extrapolate to
    the full site:

$$
\text{Estimated population} =
\bar{x}_{\text{quadrat}} \times
\frac{\text{Total site area}}{\text{Area of one quadrat}}
$$

### Stratified Random Sampling

Use this approach when the study area contains distinct environmental
gradients or habitat types that would increase variance under simple
random sampling.

1. Divide the study area into homogeneous strata based on a relevant
   environmental variable (e.g., elevation bands, vegetation type, soil
   moisture class, distance from a water source)
2. Within each stratum, establish an independent coordinate grid
   following the procedure in Section 8.3
3. Generate random coordinates separately for each stratum
4. Sample each stratum independently; do not pool random selections
   across strata
5. Weight stratum means by the proportion of total area each stratum
   represents when calculating site-level estimates

### Map-Based Sampling with Exclusion Zones

For sites where portions of the area are inaccessible or must be
excluded (buildings, water bodies, restricted zones):

1. Obtain or create a georeferenced map of the study area
2. Delineate exclusion zones and calculate the remaining accessible
   area
3. Generate random coordinates across the full extent, then discard any
   that fall within exclusion zones
4. Continue generating coordinates until the target sample size is met
   within the accessible area
5. Record discarded coordinates in the sampling log for documentation

::: {.callout-tip}
### R Implementation
An R script for map-based random sampling with exclusion polygons can be
built using the `sf` package. Generate random points with
`sf::st_sample()`, then filter against exclusion polygons with
`sf::st_difference()`.
:::


## Data Management

1. Record all sampling metadata (date, site, grid dimensions, random
   seed or table starting position, personnel) on the field data sheet
2. Enter field data into the project database within 48 hours
3. Archive original data sheets (scan if paper)
4. Version-control all analysis scripts per
   [SOP-L801 RStudio and GitHub](../lab_sops/SOP-L801-rstudio-github.qmd)


## References

- APHA, AWWA, WEF. (2012). *Standard Methods for Examination of Water
  and Wastewater*. 22nd ed. American Public Health Association (Eds.).
  Washington. 1360 pp.
- Krebs, C.J. (1999). *Ecological Methodology*. 2nd ed. Benjamin
  Cummings, Menlo Park, CA.
- Elzinga, C.L., Salzer, D.W., & Willoughby, J.W. (1998). *Measuring
  and Monitoring Plant Populations*. U.S. Bureau of Land Management
  Technical Reference 1730-1.


## Training Record

Training is documented in EA Program files. Before designing or
executing random sampling independently, personnel must:

1. Read this SOP in full
2. Complete [SOP-F001 Field Safety](SOP-F001-field-safety.qmd) training
3. Demonstrate proficiency with random number generation in R
4. Participate in a supervised field sampling session using the
   grid-based procedure
5. Be signed off by the PI or field lead
