%SOP Template 
% Version 02 Added revision date
% Version 03 Added TOC and acknowledgements
%           New SOP3_alpha.cls


\documentclass[12pt]{../SOP3_alpha}

\usepackage[english]{babel}
\usepackage{blindtext}
\usepackage{lipsum}

\title{Refridgerators and Freezers}
\date{8/15/2016}
\author{Marc Los Huertos}
\approved{TBD}
\ReviseDate{\today}
\SOPno{16 v.01}

\usepackage{Sweave}
\begin{document}
\input{Refridgerators_and_Freezers_v01-concordance}

\maketitle

\section{Scope and Application}

\NP The scope of this SOP is for the laboratory refigerator, freezer (-30) and high XX freezer (-80).

\NP The applications of this SOP is to storage and what to put in which freezer...

\section{Summary of Method}

\NP This SOP does this...

\tableofcontents

\newpage

\section{Acknowledgements}

\section{Definitions}

\NP Term1: is...

\section{Interferences}

\NP Biases and interferences can come from...

\section{Health and Safety}

\NP The refrigerator and freezers in the laboratory are not approved to store flammable material. In fact, these instruments can generate signficant static electricity that can cause an explosion when there are flammable fumes present. 

\NP No food shall be stored in the laboratory refrigerator or freezers.


\subsection*{Safety and Personnnel Protective Equipment}


\section{Personnel \& Training Responsibilities}

\NP Researchers training is required before this the procedures in this method can be used... 

\NP Researchers using this SOP should be trained for the following SOPs:

\begin{itemize}
  \item SOP01 Laboratory Safety
  \item SOP02 Field Safety
\end{itemize}

\section{Required Materials and Apparati}

\NP Item 1 w/catalog number!

\NP Item 2

\section{Reagents and Standards}

\section{Estimated Time}

\NP This procedure requires XX minutes...

\section{Sample Collection, Preservation, and Storage}

\section{Procedure}

\NP Prepare \dots

\NP

\section{Data Analysis and Calculations}

\section{QC/QA Criteria}

\section{References}

\NP APHA, AWWA. WEF. (2012) Standard Methods for examination of water and wastewater. 22nd American Public Health Association (Eds.). Washington. 1360 pp. (2014).

\end{document}
