---
title: "Quality Control and Quality Assurance"
sop-number: "SOP-L802"
author:
  - name: "Marc Los Huertos"
    affiliation: "EA Program, Pomona College"
date: last-modified
date-format: "YYYY-MM-DD"
revision: "1.1"

revision-history:
  - version: "1.0"
    date: "2025-05-16"
    author: "Marc Los Huertos"
    changes: "Converted from LaTeX to Quarto; comprehensive revision"
  - version: "1.1"
    date: "2026-06-07"
    author: "Marc Los Huertos"
    changes: "Converted plain-text references to BibTeX; added inline
              citations throughout"
---

## Purpose

This SOP establishes the overarching quality assurance (QA) and quality
control (QC) framework for all analytical work in the EA Program
laboratory. It defines the minimum QC requirements that apply across
methods, describes how to calculate and interpret QC metrics, specifies
detection limit procedures, and establishes the data documentation and
reproducibility practices required for every analysis. Individual
analytical SOPs reference this document for general QC principles and
add method-specific requirements [@usepa2007; @apha2012].

Quality assurance is the system of management actions that ensures
the analytical process produces results of known and documented quality.
Quality control is the set of technical activities (blanks, standards,
duplicates, spikes) embedded in each analytical batch to measure and
control that quality in real time [@taylor1987].


## Scope

- **Applies to:** All quantitative chemical, physical, and biological
  analyses performed in the EA Program laboratory, including those
  described in SOPs L002 through L999 and analyses of field-collected
  samples (SOPs F001 through F999)
- **Does not apply to:** Qualitative screening tests (e.g., visual soil
  texture by feel, [SOP-F401](../field_sops/SOP-F401-soil-texture-by-feel.qmd))
  unless those tests include a quantitative component
- **Location:** EA Program laboratory, SGM 133


## Responsibilities

| Role              | Responsibility                              |
|:------------------|:--------------------------------------------|
| Lab PI (QA Manager) | Approve this SOP; review QC data quarterly; authorize corrective actions |
| Lab Manager       | Maintain instrument logs, reagent inventory, and QC records; train new analysts |
| Analyst           | Implement all QC requirements for each batch; flag QC failures; document corrective actions |
| Data Reviewer     | Verify QC compliance before data are reported or published |


## Definitions

- **Accuracy:** The closeness of a measured value to the true or
  accepted reference value; assessed using certified reference
  materials (CRMs) and laboratory-fortified blanks (LFBs)
  [@taylor1987].
- **Calibration blank:** A solution containing only the analytical
  solvent or reagent matrix (no analyte); used to establish the
  instrument baseline (zero) for each calibration sequence.
- **Certified reference material (CRM):** A material with certified
  analyte concentrations traceable to a national or international
  metrology institute (NIST, USGS, IRMM); used to verify analytical
  accuracy across instruments and over time [@apha2012].
- **Control chart (Shewhart chart):** A time-series plot of QC sample
  results with control limits (typically ±2σ warning limits and ±3σ
  action limits); used to distinguish random analytical variation from
  systematic drift or instrument malfunction.
- **Continuing calibration verification (CVS):** A mid-range
  calibration standard analyzed at regular intervals (typically every
  10 samples) during an analytical run; must recover within ±10% of
  the stated value; failure requires re-calibration and re-analysis of
  affected samples [@apha2012].
- **Detection limit -- Method Detection Limit (MDL):** The minimum
  concentration of an analyte that can be detected with 99% confidence
  under stated method conditions; determined from at least 7 replicate
  analyses of a low-level spike following 40 CFR Part 136 Appendix B
  [@usepa_mdl_2016].
- **Detection limit -- Limit of Blank (LOB):** The highest signal
  expected from a blank sample; LOB = mean~blank~ + 1.645 × SD~blank~
  [@armbruster2008].
- **Detection limit -- Limit of Detection (LOD):** The lowest analyte
  concentration that can be reliably distinguished from the LOB;
  LOD = LOB + 1.645 × SD~low-concentration~; represents 95% detection
  probability [@armbruster2008].
- **Detection limit -- Limit of Quantitation (LOQ):** The lowest
  concentration at which the analyte can be measured with acceptable
  precision and accuracy, typically defined as 10 × SD at the LOD
  spike concentration; also known as the practical quantitation limit
  (PQL) [@armbruster2008].
- **Field blank:** A sample of reagent-grade water processed through
  the entire field collection and preservation procedure; detects
  contamination introduced during sampling and transport.
- **Laboratory-fortified blank (LFB):** A reagent water blank spiked
  with a known concentration of target analytes and processed through
  the full analytical procedure; monitors ongoing accuracy and
  extraction efficiency.
- **Laboratory-fortified matrix (LFM):** A field sample spiked with a
  known concentration of target analytes; detects matrix-specific
  suppression or enhancement of the analytical signal.
- **Method blank:** A reagent water sample carried through the full
  analytical procedure alongside field samples; detects contamination
  introduced by reagents, equipment, or laboratory environment
  [@apha2012].
- **Precision:** The reproducibility of repeated measurements of the
  same sample; expressed as relative percent difference (RPD) for
  duplicates or relative standard deviation (RSD) for replicate
  analyses [@taylor1987].
- **Relative percent difference (RPD):** |A - B| / ((A + B) / 2)
  × 100; the standard measure of duplicate precision; typical
  acceptance criteria are RPD ≤ 10--20% depending on method and
  matrix.
- **Spike recovery (%):** (Measured spiked concentration --
  Unspiked concentration) / Added concentration × 100; typical
  acceptance range is 80--120% for most environmental matrices
  [@apha2012].


## Safety and Hazards

QC samples contain the same analytes as field samples and are subject
to the same hazard considerations as the parent analytical method.
Consult the relevant analytical SOP for safety information.


## General QC Requirements

### Every Analytical Batch

Every analytical batch must include at minimum [@apha2012; @usepa2007]:

| QC Element            | Minimum Frequency | Purpose                         |
|:----------------------|:------------------|:--------------------------------|
| Method blank          | 1 per batch       | Detect laboratory contamination |
| Calibration blank     | Start of each calibration | Instrument baseline       |
| CVS (check standard)  | Every 10 samples  | Detect calibration drift        |
| Duplicate             | 10% of samples    | Assess precision                |
| LFB                   | 1 per batch       | Assess accuracy                 |

### Additional Elements (as applicable)

| QC Element            | When Required                              |
|:----------------------|:-------------------------------------------|
| Field blank           | Each sampling event                        |
| Matrix spike (LFM)    | 5% of samples or 1 per batch               |
| CRM                   | 1 per batch (where certified material is available) |
| Equipment blank       | When non-dedicated sampling equipment is used |


## Acceptance Criteria

| QC Element  | Typical Acceptance Criteria                              | Action if Failed |
|:------------|:---------------------------------------------------------|:-----------------|
| Method blank | < MDL (analyte absent)                                  | Investigate contamination; do not report affected samples |
| CVS         | ±10% of stated value                                    | Re-calibrate; re-analyze affected samples |
| Duplicate RPD | ≤ 10--20% (method-specific)                          | Flag results; investigate and re-analyze if warranted |
| LFB recovery | 80--120%                                               | Flag results; check reagents and instrument |
| LFM recovery | 70--130%                                               | Suspect matrix interference; consult PI |
| CRM recovery | Within certified uncertainty interval                   | Recalibrate; re-analyze |

Acceptance criteria in individual analytical SOPs take precedence over
these general values when they are stricter or when regulatory
requirements apply.


## Detection Limit Determination

### Method Detection Limit (MDL)

The MDL is determined following 40 CFR Part 136 Appendix B
[@usepa_mdl_2016]:

1. Select a spike concentration at 1--5× the estimated MDL.
2. Prepare and analyze at least 7 replicates of the spiked solution
   through the full analytical method.
3. Calculate the standard deviation ($s$) of the replicates.
4. Multiply by the one-tailed $t$-value at 99% confidence
   ($t_{n-1, 0.99}$; for $n=7$: $t = 3.143$):

$$\text{MDL} = t_{(n-1,\, 0.99)} \times s$$

5. If the spike concentration is outside the range of 0.5--20× the
   computed MDL, repeat with an adjusted spike concentration
   [@usepa_mdl_2016].

### Limit of Detection and Limit of Quantitation

The LOB, LOD, and LOQ hierarchy follows @armbruster2008:

$$\text{LOB} = \mu_{\text{blank}} + 1.645\,\sigma_{\text{blank}}$$

$$\text{LOD} = \text{LOB} + 1.645\,\sigma_{\text{low}}$$

$$\text{LOQ} = 10\,\sigma_{\text{low}}$$

where $\sigma_{\text{low}}$ is the standard deviation of replicate
analyses at the low-level spike concentration. For right-censored
environmental data, apply the non-parametric methods described in
@helsel2012.


## Control Charts

Maintain control charts for each analyte/method combination to track
long-term analytical performance:

1. Plot CRM, LFB, or CVS recovery for each batch on a time-series chart.
2. Calculate and draw lines for:
   - **Center line:** Mean recovery over the last 20+ batches
   - **Warning limits:** ±2 standard deviations from the mean
   - **Action limits:** ±3 standard deviations from the mean
3. A result between warning and action limits is a warning; investigate
   but do not necessarily re-analyze.
4. A result beyond the action limits requires corrective action and
   re-analysis.
5. Systematic trends (7+ consecutive results on one side of the center
   line, or a steady drift) require investigation even if individual
   points are within limits [@taylor1987].


## Corrective Actions

When a QC failure occurs:

1. Stop the analytical run; do not report affected samples.
2. Identify the probable cause (contaminated blank, degraded standard,
   instrument malfunction, matrix interference).
3. Correct the problem (replace reagents, recalibrate, perform
   maintenance).
4. Re-run the affected samples with fresh QC elements.
5. Verify that the repeated run passes all QC criteria.
6. Document the failure, root cause, corrective action, and re-analysis
   results in the QC log and lab notebook.


## Metadata Documentation and Reproducibility

### Laboratory Notebook Requirements

Every analytical session must be documented with at minimum:

- Date, analyst name, SOP number and revision
- Instrument ID and calibration status
- Reagent lot numbers and preparation dates
- Sample IDs, matrix type, and collection dates
- Calibration data (standards, equation, $r^2$)
- QC results (blanks, duplicates, LFBs, spikes, CRMs) with pass/fail
- Raw data file names and storage location
- Any deviations from the SOP, with justification
- Corrective actions taken for QC failures

### Electronic Data Management

1. Store all raw instrument data files on the lab shared drive in a
   folder structure organized by project, year, and instrument.
2. Never modify raw data files. All processing must be done
   programmatically in R with version-controlled scripts.
3. Commit all analysis scripts to the lab GitHub repository per
   [SOP-L801](SOP-L801-rstudio-github.qmd) [@gandrud2015; @xie2015].

### R-Based QC Tracking System

The following R framework is the EA Program standard for QC
documentation and reproducibility [@gandrud2015; @xie2015]:

```{r}
#| label: qc-tracking-setup
#| echo: true
#| eval: false

library(tidyverse)
library(lubridate)
library(knitr)

session_meta <- list(
  analyst    = "Your Name",
  date       = Sys.Date(),
  sop        = "SOP-L802 v1.1",
  method_sop = "SOP-LXX vX.X",
  instrument = "Instrument ID",
  project    = "Project Name"
)

cat(sprintf("QC session: %s | Analyst: %s | Instrument: %s\n",
            session_meta$date, session_meta$analyst,
            session_meta$instrument))
```

```{r}
#| label: qc-functions
#| echo: true
#| eval: false

calc_mdl <- function(reps, alpha = 0.01) {
  n <- length(reps); s <- sd(reps)
  t_val <- qt(1 - alpha, df = n - 1)
  list(n = n, sd = s, t_value = round(t_val, 3),
       MDL = round(t_val * s, 4), LOQ = round(10 * s, 4))
}

calc_rpd <- function(a, b) abs(a - b) / ((a + b) / 2) * 100

calc_recovery <- function(measured, expected) (measured / expected) * 100
```

Each analytical R script should:

1. Read the raw data file
2. Run all QC checks programmatically
3. Produce a formatted QC summary table
4. Generate calibration plots and control charts
5. Export a clean, QC-verified dataset as CSV
6. Render to HTML or PDF for archiving

Store templates in the project GitHub repository under
`analysis/templates/` [@gandrud2015; @xie2015].


## QC Requirements by Method Category

| QC Element         | Water Chemistry | Soil Extraction | Digestion (Metals) | Biological (Colilert) | Instrumental (TOC, ICP) |
|:-------------------|:---------------|:---------------|:-------------------|:---------------------|:----------------------|
| Calibration blank  | Each cal       | Each cal       | Each cal           | N/A                  | Each cal              |
| Method blank       | 1/batch        | 1/batch        | 1/batch            | 1/batch              | 1/batch               |
| Field blank        | 1/event        | N/A            | N/A                | N/A                  | N/A                   |
| Duplicate          | 10%            | 10%            | 10%                | 10%                  | 10%                   |
| LFB                | 1/batch        | 1/batch        | 1/batch            | N/A                  | 1/batch               |
| Matrix spike       | 5% (or 1/batch)| 5%             | 5%                 | N/A                  | 5%                    |
| CRM                | 1/batch (if avail.) | 1/batch   | 1/batch            | Positive control     | 1/batch               |
| CVS                | Every 10 samples | Every 10     | Every 10           | N/A                  | Every 10              |


## Waste Disposal

QC samples generate the same waste streams as routine samples. Dispose
of QC waste (blanks, standards, spikes) according to the parent
method's waste disposal procedure.


## References

- @apha2012
- @taylor1987
- @usepa2007
- @usepa_mdl_2016
- @armbruster2008
- @helsel2012
- @gandrud2015
- @xie2015

::: {#refs}
:::


## Related SOPs

| SOP   | Title                                  | Relationship                      |
|:------|:---------------------------------------|:----------------------------------|
| [L801](SOP-L801-rstudio-github.qmd)   | RStudio and GitHub        | Version control for analysis scripts |
| [L803](SOP-L803-data-entry-database.qmd) | Data Entry and Database  | Data recording standards          |
| [F300](../field_sops/SOP-F300-surface-water-sampling.qmd) | Surface Water Sampling | Field blank requirements  |
| [F008](../field_sops/SOP-F008-sample-labeling-coc-transport.qmd) | Sample Labeling and COC | Chain of custody          |


## Training Record

Before applying QC frameworks independently, personnel must:

1. Read this SOP in full
2. Demonstrate correct MDL calculation using the calc_mdl() function
   on a test dataset
3. Prepare a complete QC summary table for one analytical batch under
   supervision
4. Set up and interpret a control chart for one analyte/method
   combination
5. Be signed off by the PI or lab manager
