---
# =============================================================================
# SOP-L415: Carbonate Isotope Analysis by GasBench II-IRMS
# =============================================================================
# Application SOP for d13C and d18O determination in carbonate samples
# using phosphoric acid digestion on the GasBench II coupled to the
# Delta V IRMS.  Includes sample preparation, vial flushing, PAL
# autosampler operation, and post-analysis data correction with R code.
# =============================================================================

title: "Carbonate Isotope Analysis by GasBench II-IRMS"
sop-number: "SOP-L415"
author:
  - name: "Marc Los Huertos"
    affiliation: "Environmental Analysis, Pomona College"
  - name: "Kyle McCarty"
    affiliation: "Oxtoby Isotope Lab, Pomona College"
date: last-modified
date-format: "YYYY-MM-DD"
revision: "1.0"

revision-history:
  - version: "1.0"
    date: "2026-05-18"
    author: "Marc Los Huertos"
    changes: "Initial release"
---

## 1. Purpose

This SOP describes the determination of $\delta^{13}$C and
$\delta^{18}$O in carbonate minerals and sediments using phosphoric
acid digestion and headspace analysis on the GasBench II coupled to
the Delta V IRMS. It covers sample preparation, vial flushing and
acidification, the analytical run, and post-analysis data correction.

For general instrument operation, startup/shutdown, gas checks, tuning,
and pre-analysis tests, see
[SOP-L208](SOP-L208-irms-general-instrument.qmd).


## 2. Scope

- **Applies to:** Carbonate-bearing samples (calcite, dolomite,
  aragonite, shell, foraminifera, soil carbonates, cement, and similar
  materials) analyzed for $\delta^{13}$C (VPDB) and $\delta^{18}$O
  (VPDB or VSMOW) by phosphoric acid digestion and continuous-flow
  IRMS.
- **Does not apply to:** DIC in water, breath gas, CO~2~ at atmospheric
  concentrations, or water equilibration methods (these use the
  GasBench but with different sample preparation and Isodat methods).
- **Location:** Sample preparation in SGM Rm 133; analysis in the
  Oxtoby Isotope Lab, SGM Rm 135.


## 3. Responsibilities

| Role              | Responsibility                                           |
|:------------------|:---------------------------------------------------------|
| Lab PI            | Approve methods, oversee data quality                    |
| Lab Manager       | Prepare/verify Isodat methods; maintain GasBench and PAL |
| Certified User    | All steps from sample prep through data reduction        |
| Student Researcher| May perform sample prep and data reduction under supervision |


## 4. Safety and Hazards

See [SOP-L208, Section 4](SOP-L208-irms-general-instrument.qmd) for the
full hazard list. Hazards specific to this application:

::: {.callout-warning}
### Application-Specific Hazards

- **100% phosphoric acid (H~3~PO~4~):** Corrosive. Handle in the fume
  hood. Wear nitrile gloves and safety glasses at all times. In case of
  skin contact, flush with water for at least 15 minutes.
- **Syringes and needles:** Puncture hazard. Dispose of used needles in
  a sharps container.
- **Pressurized vials:** Sealed Exetainer vials may be pressurized with
  He after flushing. Vent carefully.
:::

### Required PPE

- [ ] Safety glasses
- [ ] Lab coat
- [ ] Nitrile gloves


## 5. Prerequisites

- Certification as an IRMS User per
  [SOP-L208](SOP-L208-irms-general-instrument.qmd)
- [SOP-L002 Laboratory Safety](SOP-L002-laboratory-safety.qmd)
- [SOP-L003 Hazardous Materials Handling](SOP-L003-hazardous-materials-handling.qmd)
  (for phosphoric acid)
- Familiarity with the Isodat software and GasBench/PAL interface


## 6. Required Materials

### 6.1 Equipment

| Item                               | Notes                                    |
|:-----------------------------------|:-----------------------------------------|
| GasBench II with PAL autosampler   | Oxtoby Lab, SGM 135                      |
| Heating block (or water bath)      | For maintaining reaction temperature     |
| Analytical balance                 | For weighing carbonate samples           |
| Fume hood                          | For acid handling                        |

### 6.2 Consumables

| Item                               | Notes                                    |
|:-----------------------------------|:-----------------------------------------|
| Exetainer vials (12 mL)           | Borosilicate glass with screw-cap septum |
| Septa (butyl rubber or equivalent)| Replace after each use                   |
| Syringes (various volumes)         | For acid injection                       |
| Needles (side-port or blunt-tip)   | For acid injection and vial flushing     |
| 100% phosphoric acid (H~3~PO~4~)  | Reagent grade; store in desiccator       |
| Helium (He)                        | For vial flushing                        |

### 6.3 Standards

Carbonate reference materials with certified $\delta^{13}$C and
$\delta^{18}$O values. Common choices:

| Standard  | $\delta^{13}$C (VPDB) | $\delta^{18}$O (VPDB) |
|:----------|:---------------------:|:---------------------:|
| NBS-18    | --5.01                | --23.01               |
| NBS-19    | 1.95                  | --2.20                |
| IAEA-603  | 2.46                  | --2.37                |
| LSVEC     | --46.6                | ---                   |

::: {.callout-note}
NBS-19 is exhausted at IAEA but may still be available in the lab.
IAEA-603 is its replacement. Confirm which standards are on hand with
the lab manager before planning your run.
:::


## 7. Sample Preparation

### 7.1 Weighing

1. Determine the carbonate content of your samples (if unknown, run a
   pilot batch at several masses).
2. Weigh the target mass of powdered sample into a clean, labeled
   Exetainer vial. Typical masses:

| Sample Type                | Approximate Mass |
|:---------------------------|:----------------:|
| Pure calcite / aragonite   | 100--200 $\mu$g  |
| Impure carbonate / marl    | 0.5--2.0 mg      |
| Soil carbonate (low CaCO~3~%) | 2--10 mg     |

3. Prepare standards at a mass that produces a similar signal intensity
   to your samples.
4. Include blanks (empty vials, flushed and acidified identically).

### 7.2 Vial Flushing

Each vial must be flushed with helium to remove atmospheric N~2~,
O~2~, and CO~2~ before acidification.

**Online flushing (PAL autosampler):** The PAL can flush vials
automatically, but this ties up the GasBench and IRMS. You cannot
analyze samples while flushing.

**Offline flushing (recommended for large batches):** Flush vials with
a manifold or individual He line while the IRMS analyzes a previously
prepared batch. Flush for a minimum of 5 minutes per vial at a flow
rate sufficient to displace the headspace several times over.

### 7.3 Acidification

1. Work in the fume hood.
2. Using a syringe and needle, inject the appropriate volume of 100%
   phosphoric acid through the septum into each vial:

| Reaction Temperature | Acid Volume (typical) | Reaction Time  |
|:---------------------|:---------------------:|:--------------:|
| 25 °C                | 0.1--0.3 mL           | 24 hours       |
| 50 °C                | 0.1--0.3 mL           | 2--4 hours     |
| 70 °C                | 0.1--0.3 mL           | 1--2 hours     |

3. Place acidified vials in the heating block or water bath at the
   specified reaction temperature for the required duration.
4. Gently swirl or agitate vials periodically to ensure complete
   reaction.

::: {.callout-caution}
The acid-fractionation factor for $\delta^{18}$O is temperature
dependent. Use a consistent reaction temperature for all samples and
standards in a batch. The Isodat method must match the temperature used.
:::


## 8. Sequence Design

A typical GasBench carbonate sequence:

| Position(s) | Contents                          | Purpose                        |
|:------------|:----------------------------------|:-------------------------------|
| 1--3        | Standard A (e.g., NBS-19)        | Normalization + conditioning   |
| 4--6        | Standard B (e.g., NBS-18)        | Normalization anchor 2         |
| 7           | Blank (empty flushed/acidified vial) | Blank check               |
| 8--15       | Unknown samples                   |                                |
| 16--18      | Check standard (e.g., NBS-19)    | Drift/QC monitor               |
| 19--26      | Unknown samples                   |                                |
| 27--29      | Check standard                    | Drift/QC monitor               |
| ...         | Continue pattern                  |                                |
| Last 3      | Standard A                        | End-of-run normalization check |
| Last 3      | Standard B                        | End-of-run normalization check |

The GasBench samples each vial 11 times (11 injections). The
CO~2~\_630 method takes approximately 10.5 minutes per vial plus ~1
minute for peak centering and magnet switching.


## 9. Running the Analysis

1. Complete all pre-analysis checks per
   [SOP-L208, Section 13](SOP-L208-irms-general-instrument.qmd#sec-pre-analysis).
   The GasBench is a low-flow peripheral, so backgrounds tend to be
   lower and on/off tests pass more quickly than with the Flash EA.
2. Load vials into the PAL autosampler tray. Verify positions match
   the sequence.
3. Build the sequence in Isodat. Select the appropriate carbonate
   method (e.g., CO~2~\_630). Ensure **Peak Center** is checked.
4. Click **Start**. Save as:
   `YYYYMMDD_Project_SequenceID_Username.seq`
5. Monitor the first few vials to confirm peak shapes and signal
   intensities are within the expected range.


## 10. End-of-Run Shutdown and Cleanup

1. Confirm the sequence completed successfully.
2. Remove vials from the PAL tray.
3. Dispose of acidified vials and used needles/syringes properly
   (sharps container for needles; acid waste for residual acid).
4. Follow the shutdown procedure in SOP-L208, Section 11.3.
5. Log gas usage and issues in the lab logbook.
6. Clean up. Push in your chair.


## 11. Post-Analysis Data Correction

### 11.1 Overview

The correction sequence is the same as for EA analyses
(see [SOP-L414, Section 10](SOP-L414-soil-ea-irms.qmd)), applied to
$\delta^{13}$C and $\delta^{18}$O:

1. Blank correction (subtract blank CO~2~ contribution)
2. Linearity correction
3. Scale normalization (two-point or multi-point)
4. Drift correction
5. QC evaluation

### 11.2 GasBench-Specific Considerations

The GasBench produces 11 replicate injections per vial. Standard
practice is to:

- Discard the first 2--3 injections (carry-over and equilibration
  artifacts).
- Average the remaining 8--9 injections to produce a single
  $\delta^{13}$C and $\delta^{18}$O value per vial.
- Evaluate the within-vial standard deviation as an internal precision
  check.

```{r}
#| label: gasbench-averaging
#| echo: true
#| eval: false

library(tidyverse)

raw <- read_csv("data/raw/YYYYMMDD_Project_SeqID_User_raw.csv")

# Each vial has 11 injections. Discard the first 3.
vial_means <- raw %>%
  group_by(Identifier1) %>%
  mutate(injection = row_number()) %>%
  filter(injection > 3) %>%
  summarize(
    d13C_mean    = mean(d13C_raw, na.rm = TRUE),
    d13C_sd      = sd(d13C_raw, na.rm = TRUE),
    d18O_mean    = mean(d18O_raw, na.rm = TRUE),
    d18O_sd      = sd(d18O_raw, na.rm = TRUE),
    Area44_mean  = mean(Area44, na.rm = TRUE),
    n_injections = n(),
    .groups      = "drop"
  )

# Classify
vial_means <- vial_means %>%
  mutate(
    type = case_when(
      str_detect(Identifier1, regex("blank|BLK", ignore_case = TRUE)) ~ "blank",
      str_detect(Identifier1, "NBS|IAEA|LSVEC")                      ~ "standard",
      TRUE                                                             ~ "sample"
    )
  )
```

### 11.3 Scale Normalization for Carbonates

```{r}
#| label: carbonate-normalization
#| echo: true
#| eval: false

std_cert <- tibble(
  Identifier1 = c("NBS-19", "NBS-18"),
  d13C_true   = c(1.95, -5.01),
  d18O_true   = c(-2.20, -23.01)
)

std_meas <- vial_means %>%
  filter(Identifier1 %in% std_cert$Identifier1) %>%
  left_join(std_cert, by = "Identifier1")

fit_C <- lm(d13C_true ~ d13C_mean, data = std_meas)
fit_O <- lm(d18O_true ~ d18O_mean, data = std_meas)

cat(sprintf("d13C: slope = %.4f, intercept = %.4f\n",
            coef(fit_C)[2], coef(fit_C)[1]))
cat(sprintf("d18O: slope = %.4f, intercept = %.4f\n",
            coef(fit_O)[2], coef(fit_O)[1]))

vial_means <- vial_means %>%
  mutate(
    d13C_cal = coef(fit_C)[1] + coef(fit_C)[2] * d13C_mean,
    d18O_cal = coef(fit_O)[1] + coef(fit_O)[2] * d18O_mean
  )
```

### 11.4 Drift and QC

Apply drift correction following the same approach as
[SOP-L414, Section 10.7](SOP-L414-soil-ea-irms.qmd). Use your check
standard (e.g., NBS-19) to monitor drift.

```{r}
#| label: carbonate-qc
#| echo: true
#| eval: false

qc <- vial_means %>%
  filter(type == "standard") %>%
  group_by(Identifier1) %>%
  summarize(
    n         = n(),
    d13C_mean = mean(d13C_cal, na.rm = TRUE),
    d13C_sd   = sd(d13C_cal, na.rm = TRUE),
    d18O_mean = mean(d18O_cal, na.rm = TRUE),
    d18O_sd   = sd(d18O_cal, na.rm = TRUE),
    .groups   = "drop"
  )

print(qc)

results <- vial_means %>%
  filter(type == "sample") %>%
  select(Identifier1, d13C_cal, d18O_cal)

write_csv(results,
  "data/processed/YYYYMMDD_Project_SeqID_carbonates_calibrated.csv")
write_csv(qc,
  "data/processed/YYYYMMDD_Project_SeqID_carbonates_qc.csv")
```


## 12. QC/QA Criteria

### 12.1 Within-Vial Precision

The standard deviation of the 8--9 retained injections per vial should
be:

| Parameter        | Maximum 1$\sigma$ (per mil) |
|:-----------------|:---------------------------:|
| $\delta^{13}$C   | 0.06                        |
| $\delta^{18}$O   | 0.10                        |

Vials exceeding these thresholds may indicate insufficient sample,
incomplete reaction, or septum leakage.

### 12.2 Between-Vial Standard Reproducibility

| Parameter        | Maximum 1$\sigma$ (per mil) |
|:-----------------|:---------------------------:|
| $\delta^{13}$C   | 0.10                        |
| $\delta^{18}$O   | 0.15                        |

### 12.3 Calibration Accuracy

After normalization, the mean of each reference material should fall
within 0.10 per mil ($\delta^{13}$C) or 0.15 per mil ($\delta^{18}$O)
of its certified value.

### 12.4 Blanks

CO~2~ signal in blank vials should be <1% of the mean sample signal.


## 13. Reporting

Report final values as:

- $\delta^{13}$C vs. VPDB
- $\delta^{18}$O vs. VPDB (conventional for carbonates) or VSMOW
  (specify which)

Include: reference materials and certified values, within-run
precision, corrections applied, acid digestion temperature, and the
acid-fractionation factor used (if converting from acid-liberated CO~2~
to mineral $\delta^{18}$O).


## 14. Troubleshooting

| Symptom                              | Possible Cause / Action                       |
|:-------------------------------------|:----------------------------------------------|
| No CO~2~ peak or very small signal   | Insufficient sample mass; incomplete reaction |
| High blank signal                    | Atmospheric CO~2~ leaking through septum; replace septa |
| Poor within-vial precision           | Septum leak; vial not fully flushed           |
| $\delta^{18}$O offset but $\delta^{13}$C OK | Wrong reaction temperature; acid fractionation mismatch |
| First 3 injections wildly different  | Normal carry-over; discard as planned         |
| Signal too large (off scale)         | Too much sample; reduce mass                  |


## 15. References

Coplen, T. B. (2011). Guidelines and recommended terms for expression
of stable-isotope-ratio and gas-ratio measurement results. *Rapid
Communications in Mass Spectrometry*, 25(17), 2538--2560.

Kim, S.-T., Mucci, A., & Taylor, B. E. (2007). Phosphoric acid
fractionation factors for calcite and aragonite between 25 and 75 °C.
*Chemical Geology*, 246(3--4), 135--146.

Swart, P. K., Burns, S. J., & Leder, J. J. (1991). Fractionation of
the stable isotopes of oxygen and carbon in carbon dioxide during the
reaction of calcite with phosphoric acid as a function of temperature
and technique. *Chemical Geology: Isotope Geoscience Section*, 86(2),
89--96.


## 16. Related SOPs

| SOP   | Title                                       | Relationship                      |
|:------|:--------------------------------------------|:----------------------------------|
| [L001](SOP-L002-laboratory-safety.qmd)       | Laboratory Safety       | Required prerequisite             |
| [L003](SOP-L003-hazardous-materials-handling.qmd) | Hazardous Materials Handling | Acid handling               |
| [L802](SOP-L802-quality-control.qmd)         | Quality Control and Quality Assurance | General QC framework   |
| [L208](SOP-L208-irms-general-instrument.qmd) | IRMS -- General Instrument Use | Instrument-level procedures |
| [L414](SOP-L414-soil-ea-irms.qmd)            | Soil CN by Flash EA-IRMS | Companion application SOP        |


## 17. Training Record

Before performing carbonate isotope analysis independently, personnel
must:

1. Read this SOP and [SOP-L208](SOP-L208-irms-general-instrument.qmd) in
   full
2. Be a certified IRMS User (SOP-L208, Section 6)
3. Complete
   [SOP-L003 Hazardous Materials Handling](SOP-L003-hazardous-materials-handling.qmd)
   (acid handling)
4. Observe a GasBench carbonate run from vial preparation through data
   export
5. Perform a supervised run with acceptable QC results
6. Complete a supervised data reduction session
7. Be signed off by the lab manager or PI
