---
# =============================================================================
# SOP-L804: GIS and Spatial Data Management
# =============================================================================
title: "GIS and Spatial Data Management"
sop-number: "SOP-L804"
author:
  - name: "Marc Los Huertos"
    affiliation: "EA Program, Pomona College"
date: last-modified
date-format: "YYYY-MM-DD"
revision: "0.1"

revision-history:
  - version: "0.1"
    date: "2026-05-21"
    author: "Marc Los Huertos"
    changes: "Initial release"
---

## Purpose

This SOP establishes procedures for acquiring, organizing, projecting,
and archiving geospatial data used in EA Program research. It covers
coordinate reference systems, file formats, metadata standards, and
reproducible mapping workflows in R, ensuring that spatial data are
interoperable across projects and analysts [@pebesma2023; @lovelace2019].


## Scope

- **Applies to:** All vector (point, line, polygon) and raster
  (satellite imagery, DEMs, interpolated surfaces) data used in EA
  Program projects
- **Does not apply to:** Drone image acquisition or photogrammetric
  processing (see [SOP-L805](SOP-L805-drone-photogrammetry.qmd));
  GPS receiver operation in the field (see relevant field SOPs)
- **Location:** EA Program laboratory (SGM 133) and RStudio Server


## Responsibilities

| Role              | Responsibility                              |
|:------------------|:--------------------------------------------|
| Lab PI            | Approve spatial data management plan; ensure data-sharing compliance |
| GIS Lead          | Maintain spatial data library; enforce projection and metadata standards |
| Analyst / Student | Follow naming conventions; document coordinate systems; archive final layers |
| Data Reviewer     | Verify spatial metadata completeness before publication |

<!-- Definitions snippet for SOP-L804: GIS and Spatial Data Management -->
<!-- Insert immediately before ## Safety and Hazards -->

## Definitions

- **Coordinate reference system (CRS):** A framework that associates coordinates with locations on Earth; consists of a datum (reference ellipsoid), a projection (for 2D maps), and a coordinate system (e.g., latitude/longitude or meters); data layers must share the same CRS for spatial operations.
- **Geodatabase:** A structured container for storing multiple spatial datasets, attribute tables, and topological rules in a single file; the preferred format for complex GIS projects with many related layers.
- **GeoTIFF:** A TIFF image file with embedded geospatial metadata (CRS, extent, resolution); the standard raster format for spatial data exchange in remote sensing and GIS.
- **Raster:** A spatial data model representing continuous surfaces as a grid of equally sized cells (pixels), each with one or more attribute values; used for elevation, imagery, and interpolated surfaces.
- **Shapefile (.shp):** A common vector spatial data format introduced by Esri; stores geometry (points, lines, polygons) and attribute data in multiple associated files (.shp, .dbf, .shx, .prj minimum); widely supported but limited by 2 GB file size and 10-character field name constraints.
- **Vector:** A spatial data model representing discrete features as points, lines, or polygons defined by coordinate pairs; used for sample locations, stream networks, and land parcel boundaries.
- **WGS84 (World Geodetic System 1984):** The global geodetic datum used by GPS; EPSG code 4326; the required CRS for all EA Program GPS-collected point data before projection into a local coordinate system.


## Safety and Hazards

GIS work is a computational activity with no physical hazards.
When using GPS receivers or conducting ground-truth fieldwork, follow
[SOP-F001 Field Safety](../field_sops/SOP-F001-field-safety.qmd).


## Equipment and Materials

### Software

| Item                    | Access                | Notes                        |
|:------------------------|:----------------------|:-----------------------------|
| R (version 4.x)        | RStudio Server        | Primary analysis platform    |
| `sf` package            | CRAN                  | Vector data handling         |
| `terra` package          | CRAN                  | Raster data handling         |
| `tmap` / `ggplot2`      | CRAN                  | Map production               |
| QGIS (3.x)             | Free download         | Visual inspection and digitizing |
| Google Earth Engine     | earthengine.google.com | Satellite imagery processing |

### Hardware

| Item                    | Location              | Notes                        |
|:------------------------|:----------------------|:-----------------------------|
| GPS receivers (Garmin)  | SGM 133 equipment cabinet | Field coordinate collection |
| High-resolution monitor | SGM 133               | For digitizing and map review |


## Procedure

### Coordinate Reference Systems

All EA Program spatial data must use one of the following coordinate
reference systems (CRS). Document the CRS in every file, script, and
metadata record.

| CRS Name                      | EPSG Code | Use Case                     |
|:------------------------------|:----------|:-----------------------------|
| WGS 84 (geographic)          | 4326      | GPS data, web maps, data exchange |
| NAD83 / UTM Zone 11N         | 26911     | Local analysis in Southern California |
| NAD83 / California Zone 5    | 2229      | County and municipal data    |

::: {.callout-tip}
### Tip
When in doubt, store data in EPSG:4326 (latitude/longitude, WGS 84).
Transform to a projected CRS for distance and area calculations.
:::

When loading spatial data in R, always verify and set the CRS
explicitly:

```{r}
#| label: crs-check
#| echo: true
#| eval: false

library(sf)

# Read a shapefile
sites <- st_read("data/spatial/sampling_sites.shp")

# Check CRS
st_crs(sites)

# Transform to UTM Zone 11N for area calculations
sites_utm <- st_transform(sites, crs = 26911)
```

### File Formats

| Format        | Extension    | Use Case                          | Notes                      |
|:--------------|:-------------|:----------------------------------|:---------------------------|
| GeoPackage    | `.gpkg`      | Preferred for all vector data     | Single file, open standard |
| Shapefile     | `.shp` + `.dbf` + `.shx` + `.prj` | Legacy compatibility | Keep all component files together |
| GeoTIFF       | `.tif`       | Raster data                       | Include CRS in header      |
| CSV with XY   | `.csv`       | Simple point data                 | Requires separate CRS documentation |
| GeoJSON       | `.geojson`   | Web applications                  | Always WGS 84              |

GeoPackage (`.gpkg`) is the preferred format for new projects because
it stores geometry, attributes, and CRS in a single file without the
limitations of shapefiles (10-character field names, 2 GB size limit)
[@pebesma2023].

### File Naming and Directory Structure

Follow the conventions in [SOP-L803](SOP-L803-data-entry-database.qmd)
with the addition of a `data/spatial/` subdirectory:

```
project_name/
├── data/
│   ├── raw/
│   ├── processed/
│   ├── spatial/
│   │   ├── raw/          # Downloaded or collected spatial data
│   │   ├── processed/    # Clipped, projected, joined layers
│   │   └── metadata/     # FGDC or ISO metadata files
│   └── metadata/
├── analysis/
│   ├── scripts/
│   └── maps/             # Exported map figures
└── README.md
```

### Spatial Data Acquisition

Common data sources for EA Program research:

| Source                        | URL                                    | Data Types             |
|:------------------------------|:---------------------------------------|:-----------------------|
| USGS National Map             | nationalmap.gov                        | DEMs, hydrography, boundaries |
| USGS EarthExplorer            | earthexplorer.usgs.gov                 | Landsat, NAIP imagery  |
| CA State Geoportal            | gis.data.ca.gov                        | State and county layers |
| NRCS Web Soil Survey          | websoilsurvey.sc.egov.usda.gov         | SSURGO soil data       |
| OpenStreetMap                 | openstreetmap.org                      | Roads, buildings, land use |
| Google Earth Engine           | earthengine.google.com                 | Multi-temporal satellite data |

When downloading external data, record the source URL, download date,
and any subsetting parameters in the metadata file.

### Spatial Data Processing in R

A standard workflow for processing spatial data:

```{r}
#| label: spatial-workflow
#| echo: true
#| eval: false

library(sf)
library(terra)
library(tidyverse)

# --- 1. Load vector data ---
sites <- st_read("data/spatial/raw/sampling_sites.gpkg")
watershed <- st_read("data/spatial/raw/claremont_watershed.gpkg")

# --- 2. Verify and harmonize CRS ---
st_crs(sites) == st_crs(watershed)
sites <- st_transform(sites, crs = st_crs(watershed))

# --- 3. Spatial operations ---
# Clip sites to watershed boundary
sites_in_watershed <- st_intersection(sites, watershed)

# Buffer sites by 100 m
site_buffers <- st_buffer(sites_in_watershed, dist = 100)

# --- 4. Load and extract raster data ---
dem <- rast("data/spatial/raw/dem_10m.tif")
sites_in_watershed$elevation_m <- extract(dem, sites_in_watershed)$value

# --- 5. Save processed outputs ---
st_write(sites_in_watershed,
         "data/spatial/processed/sites_with_elevation.gpkg",
         delete_dsn = TRUE)
```

### Map Production

All maps for reports and publications must include:

1. Title and date
2. Scale bar (use `tmap::tm_scale_bar()` or `ggspatial::annotation_scale()`)
3. North arrow
4. Legend with clear labels and units
5. Data source attribution
6. CRS/projection noted in the caption or margin

```{r}
#| label: map-template
#| echo: true
#| eval: false

library(tmap)

tm_shape(watershed) +
  tm_polygons(col = "lightblue", alpha = 0.3) +
  tm_shape(sites_in_watershed) +
  tm_dots(col = "elevation_m", size = 0.5,
          palette = "viridis",
          title = "Elevation (m)") +
  tm_scale_bar(position = c("left", "bottom")) +
  tm_compass(position = c("right", "top")) +
  tm_layout(
    title = "Sampling Sites: Claremont Watershed",
    title.size = 1.2,
    legend.outside = TRUE
  )
```

### Metadata Documentation

Each spatial dataset must have a metadata record. At minimum, include:

| Field                 | Description                               |
|:----------------------|:------------------------------------------|
| Title                 | Descriptive name of the layer             |
| Abstract              | Brief description of content and purpose  |
| CRS                   | EPSG code and full name                   |
| Extent                | Bounding box coordinates                  |
| Source                 | Origin of data (URL, agency, field GPS)   |
| Date created/acquired | When the data were generated or downloaded |
| Processing steps      | Summary of transformations applied        |
| Contact               | Analyst name and email                    |

For formal publications, use FGDC or ISO 19115 metadata standards.
For internal projects, the table above saved as a text file alongside
the data is sufficient.

### GPS Data Import

When importing GPS waypoints or tracks:

1. Download data from the GPS unit to `data/spatial/raw/` in GPX format
2. Import into R using `sf::st_read("file.gpx", layer = "waypoints")`
3. Verify coordinates visually by plotting over a basemap
4. Assign the correct CRS (typically EPSG:4326 for consumer GPS units)
5. Document the GPS model, estimated accuracy, and collection conditions


## Quality Control and Acceptance Criteria

- Verify CRS for every imported dataset before analysis
- Check that point locations fall within expected geographic bounds
  (e.g., all sites should be within the study area polygon)
- Validate topology for polygon layers (no gaps, overlaps, or
  self-intersections) using `sf::st_is_valid()`
- Compare GPS coordinates with known reference points when available
- Document the spatial accuracy and precision of each data source


## Waste Disposal

Not applicable (computational procedure).


## References

::: {#refs}
:::


## Training Record

Training is documented in the EA Program training log. Before working
independently with spatial data, personnel must:

1. Read this SOP in full
2. Complete the RStudio and GitHub orientation
   ([SOP-L801](SOP-L801-rstudio-github.qmd))
3. Demonstrate proficiency with `sf` and basic spatial operations
   in R under supervision
4. Produce a correctly formatted map with all required elements
5. Be signed off by the PI or GIS lead


## Appendix A: Recommended R Packages for Spatial Analysis

| Package          | Purpose                                     |
|:-----------------|:--------------------------------------------|
| `sf`             | Vector data (points, lines, polygons)       |
| `terra`          | Raster data analysis                        |
| `stars`          | Spatiotemporal raster/vector data cubes     |
| `tmap`           | Thematic maps (static and interactive)      |
| `leaflet`        | Interactive web maps                        |
| `ggspatial`      | Spatial layers in ggplot2                   |
| `mapview`        | Quick interactive map previews              |
| `exactextractr`  | Fast raster extraction by polygons          |
| `geodata`        | Download climate, soil, and elevation data  |
| `tidyterra`      | Tidy methods for terra objects              |
