---
# =============================================================================
# SOP-L82: ddPCR Data Export and Analysis -- Nitrogen Cycle Gene Abundance
# =============================================================================
title: "ddPCR Data Export and Analysis -- Nitrogen Cycle Gene Abundance"
sop-number: "SOP-L62"
author:
  - name: "Marc Los Huertos"
    affiliation: "EA Program, Pomona College"
date: last-modified
date-format: "YYYY-MM-DD"
revision: "0.1"

revision-history:
  - version: "0.1"
    date: "2025-05-14"
    author: "Marc Los Huertos"
    changes: "Initial draft"
---

## 1. Purpose

This procedure describes the workflow for exporting raw data from the
Bio-Rad QX200 ddPCR system, performing quality control checks,
calculating gene copy numbers normalized to soil mass or DNA
concentration, and conducting statistical analyses of nitrogen cycle gene
abundance data. The SOP links the instrument output from
[SOP-L81](SOP-L81-ddpcr-nitrogen-cycle-primers.qmd) to reproducible
analysis scripts written in R [@rcoreteam2024].


## 2. Scope

- **Applies to:** Processing and analysis of ddPCR data (QuantaSoft or
  QuantaSoft Analysis Pro output) for nitrogen cycle functional genes
  (*amoA*, *nirK*, *nirS*, *nosZ*, *nifH*, *nrfA*) from environmental
  samples
- **Does not apply to:** Amplicon sequencing data, metagenomics, or
  qPCR data processed with different software
- **Location:** Instrument computer (QuantaSoft) and RStudio Server
  (see [SOP-L05 RStudio and GitHub](SOP-L05-rstudio-github.qmd))


## 3. Responsibilities

| Role              | Responsibility                              |
|:------------------|:--------------------------------------------|
| Lab PI            | Approve analysis pipeline and interpretation|
| Trained Personnel | Export data, run QC, perform analysis        |
| New Students      | Complete training before independent analysis|


## 4. Safety and Hazards

No physical or chemical hazards are associated with this procedure. Data
integrity and reproducibility are the primary concerns.

::: {.callout-warning}
### Data Integrity

- Never overwrite raw exported data files. Work only with copies.
- All analysis scripts must be version-controlled in the lab GitHub
  repository.
- Back up raw `.qlp` files from QuantaSoft to the lab shared drive
  immediately after each run.
:::


## 5. Equipment and Materials

### Equipment

| Item                              | Location           | Notes                          |
|:----------------------------------|:-------------------|:-------------------------------|
| QX200 instrument computer         | Instrument room    | Runs QuantaSoft or QuantaSoft Analysis Pro |
| RStudio Server                    | Remote (see SOP-L05) | For R-based analysis          |
| USB drive or network share        | Lab                | For transferring `.qlp` and `.csv` files |

### Software

| Software                          | Version         | Purpose                        |
|:----------------------------------|:----------------|:-------------------------------|
| QuantaSoft or QuantaSoft Analysis Pro | 1.7+ or 1.0+ | Threshold setting, data export |
| R                                 | 4.3+            | Statistical analysis           |
| RStudio                           | Current         | IDE for R scripts              |
| `tidyverse` R package             | Current         | Data wrangling and visualization |
| `ggplot2` R package               | Current         | Plotting                       |
| *TBD -- identify additional R packages for ddPCR analysis (e.g., `ddpcr`, `twoddpcr`)* | | |


## 6. Procedure

### 6.1 Data Export from QuantaSoft

1. Open the completed experiment file (`.qlp`) in QuantaSoft
2. Review the droplet plots for each well. Confirm that the
   positive/negative threshold was set appropriately during the run
   (see SOP-L81, Section 6.6)
3. If threshold adjustment is needed, use QuantaSoft Analysis Pro to
   re-apply thresholds manually. Document any threshold changes in the
   lab notebook.
4. Export data using **Options > Export CSV** to generate a
   comma-separated file containing well-level results
5. The exported CSV includes the following key columns:

   - `Well`: plate position (e.g., A01)
   - `Sample`: sample name as entered during plate setup
   - `Target`: gene target name
   - `Concentration`: copies per uL of the final ddPCR reaction
     (Poisson-corrected) [@hindson2011]
   - `Positives`: number of positive droplets
   - `Negatives`: number of negative droplets
   - `AcceptedDroplets`: total accepted droplets
   - `MeanAmplitudeOfPositives`: mean fluorescence of positive cluster
   - `MeanAmplitudeOfNegatives`: mean fluorescence of negative cluster

6. Save the CSV with the naming convention:
   `YYYY-MM-DD_gene_project_plate.csv`
   (e.g., `2025-05-14_amoA_wetland_plate1.csv`)
7. Copy the `.qlp` file and the exported `.csv` to the lab shared drive
   under `ddPCR_raw_data/`

### 6.2 Quality Control Checks

Perform the following checks on the exported data before proceeding to
analysis. These can be scripted in R for reproducibility.

1. **Droplet count filter:** Exclude any well with fewer than 10,000
   accepted droplets. Wells below this threshold have unreliable Poisson
   statistics [@biorad_qx200_2024].

2. **No-template controls (NTC):** NTC wells must show 0 or fewer than
   3 positive droplets. If NTC wells show more than 3 positives, the
   entire plate or the affected target should be considered contaminated
   and rerun.

3. **Extraction blanks:** Extraction blanks (from SOP-L80) must show
   fewer positive droplets than the NTC acceptance threshold. Elevated
   counts in extraction blanks indicate contamination in the DNA
   extraction workflow.

4. **Positive controls:** Positive control gBlock wells should fall
   within 20% of the expected concentration based on the dilution
   prepared.

5. **Technical replicate agreement:** If technical replicates were run,
   calculate the coefficient of variation (CV) across replicates. A CV
   greater than 25% suggests pipetting error or sample heterogeneity.

::: {.callout-tip}
### Tip
Create a QC report template in R Markdown or Quarto that automatically
flags wells failing any of the criteria above. This makes batch QC
efficient and ensures nothing is missed.
:::

### 6.3 Calculating Gene Copy Numbers

The QuantaSoft `Concentration` output is in copies per uL of the ddPCR
reaction mixture. To express results per gram of dry soil (or per unit
of extracted DNA), apply the following conversions:

$$
\text{Copies per g dry soil} =
\frac{C_{\text{ddPCR}} \times V_{\text{reaction}} \times DF \times V_{\text{elution}}}{V_{\text{template}} \times m_{\text{soil}} \times (1 - \theta_g)}
$$

where:

- $C_{\text{ddPCR}}$ = concentration from QuantaSoft (copies/uL)
- $V_{\text{reaction}}$ = total ddPCR reaction volume (typically 20 uL)
- $DF$ = dilution factor of the DNA template (e.g., 10 if diluted 1:10)
- $V_{\text{elution}}$ = total DNA elution volume from extraction (uL;
  see SOP-L80)
- $V_{\text{template}}$ = volume of DNA template added to the ddPCR
  reaction (uL)
- $m_{\text{soil}}$ = mass of soil extracted (g)
- $\theta_g$ = gravimetric soil moisture (as a decimal; see
  [SOP-L41](SOP-L41-soil-moisture.qmd))

To express results per ng of extracted DNA instead:

$$
\text{Copies per ng DNA} =
\frac{C_{\text{ddPCR}} \times V_{\text{reaction}}}{V_{\text{template}} \times C_{\text{DNA}}}
$$

where $C_{\text{DNA}}$ is the Qubit DNA concentration (ng/uL) from
SOP-L80.

```{r}
#| label: copy-number-calc
#| echo: true

# Example calculation: copies per g dry soil
c_ddpcr <- 52.3         # copies/uL (from QuantaSoft)
v_reaction <- 20         # uL
dilution_factor <- 10    # 1:10 dilution of template
v_elution <- 100         # uL (from SOP-L80)
v_template <- 2          # uL added to ddPCR reaction
soil_mass_g <- 0.25      # g soil extracted
moisture_decimal <- 0.35 # gravimetric moisture as decimal

copies_per_g <- (c_ddpcr * v_reaction * dilution_factor * v_elution) /
                (v_template * soil_mass_g * (1 - moisture_decimal))

cat(sprintf("Gene copies per g dry soil: %.0f\n", copies_per_g))
cat(sprintf("Gene copies per g dry soil (log10): %.2f\n", log10(copies_per_g)))
```

### 6.4 Data Organization and Merging

1. Merge the QC-filtered ddPCR results with the sample metadata (site,
   treatment, depth, date, soil moisture, DNA concentration) using a
   shared sample ID column
2. Organize the merged data in a tidy (long) format with one row per
   sample-by-gene combination [@wickham2014]
3. Save the merged analysis-ready dataset as a `.csv` in the project
   GitHub repository under `data/processed/`
4. *TBD -- define the standard column naming convention for the lab's
   ddPCR datasets (e.g., `sample_id`, `gene`, `copies_per_g_dry`,
   `copies_per_ng_dna`, `site`, `treatment`, `depth_cm`, `date`)*

### 6.5 Statistical Analysis

::: {.callout-note}
The choice of statistical approach depends on the experimental design
(e.g., paired vs. independent sites, factorial treatments, time series).
The guidance below covers common analyses for nitrogen cycle gene
abundance. Consult the PI before selecting a final analytical approach.
:::

#### 6.5.1 Data Exploration and Visualization

1. Plot gene copy numbers (log~10~-transformed) by treatment, site, or
   depth using boxplots or dot plots
2. Assess normality of log-transformed data using Shapiro-Wilk tests and
   Q-Q plots
3. Check for outliers and investigate whether they represent true
   biological variation or analytical artifacts

#### 6.5.2 Comparing Gene Abundance Across Treatments

- For two-group comparisons (e.g., amended vs. control): use Welch's
  t-test on log~10~-transformed copies per g dry soil, or the
  non-parametric Wilcoxon rank-sum test if normality assumptions are
  not met [@frostegard1999]
- For multi-group comparisons: use one-way ANOVA on log~10~-transformed
  data with Tukey HSD post hoc tests, or Kruskal-Wallis with Dunn's
  test for non-parametric data
- For factorial designs (e.g., treatment x depth): use two-way ANOVA or
  linear mixed-effects models with site as a random effect
  [@pinheiro2000]

#### 6.5.3 Gene Ratios

Ratios of functional gene abundances can provide insight into the
balance of nitrogen cycling processes [@hallin2009]:

- *nirK*/*nirS*: relative contribution of the two nitrite reductase
  pathways to denitrification
- *nosZ*/(*nirK* + *nirS*): completeness of denitrification (higher
  values suggest greater N~2~O reduction capacity)
- *amoA* AOA/*amoA* AOB: relative dominance of archaeal vs. bacterial
  ammonia oxidation
- *nrfA*/(*nirK* + *nirS*): relative importance of DNRA vs.
  denitrification

*TBD -- determine which gene ratios are most relevant for the lab's
current research questions and establish standard visualization formats.*

#### 6.5.4 Correlations with Environmental Variables

- Correlate gene abundances with soil physicochemical data (pH, moisture,
  total C, total N, nitrate, ammonium) using Pearson or Spearman
  correlations
- *TBD -- evaluate whether multivariate approaches (e.g., PCA,
  redundancy analysis) are appropriate given the number of samples and
  variables*
- *TBD -- determine whether gene abundance data should be related to
  process rate measurements (e.g., N~2~O flux, nitrification potential)
  and define the analytical framework for doing so*

#### 6.5.5 Reporting

- Report gene abundance as copies per g dry soil (mean +/- standard
  error) and/or copies per ng DNA
- Present data on a log~10~ scale in figures; report back-transformed
  means and confidence intervals in tables where appropriate
  [@bolker2009]
- Include all QC metrics (droplet counts, NTC results, positive control
  recovery) as supplementary tables
- *TBD -- create a standardized R Markdown/Quarto reporting template
  for ddPCR results that the lab can reuse across projects*


## 7. Quality Control and Acceptance Criteria

- Raw `.qlp` files must be archived on the lab shared drive before any
  analysis proceeds
- All R scripts must be committed to the lab GitHub repository with
  meaningful commit messages
- QC filtering must be applied before any statistical analysis;
  unfiltered data should never be reported
- Positive control recovery within 20% of expected value
- *TBD -- define minimum sample sizes required for each statistical test
  based on power analysis*
- *TBD -- establish whether and how to handle values below the limit of
  detection (e.g., substitution with LOD/2, censored regression, or
  Tobit models) [@helsel2011]*


## 8. Waste Disposal

No physical waste is generated by this procedure. Data management
practices are described in Section 6.4.


## 9. References

::: {#refs}
:::


## 10. Training Record

Before performing this procedure independently, personnel must:

1. Read this SOP, [SOP-L81 Primer Design](SOP-L81-ddpcr-nitrogen-cycle-primers.qmd),
   and [SOP-L80 DNA Extraction](SOP-L80-dna-extraction.qmd)
2. Complete [SOP-L05 RStudio and GitHub](SOP-L05-rstudio-github.qmd)
   training
3. Demonstrate competence exporting data from QuantaSoft
4. Successfully run the QC and analysis pipeline on a practice dataset
   under supervision
5. Be signed off by the PI or lab manager


## Appendix A: Example QC Filtering Script {.appendix}

```{r}
#| label: qc-filter-example
#| echo: true
#| eval: false

library(tidyverse)

# Read exported QuantaSoft CSV
raw <- read_csv("data/raw/2025-05-14_amoA_wetland_plate1.csv")

# QC filtering
qc_filtered <- raw %>%
  filter(
    AcceptedDroplets >= 10000        # Minimum droplet threshold
  ) %>%
  mutate(
    qc_flag = case_when(
      Sample == "NTC" & Positives > 3 ~ "NTC_contamination",
      Sample == "Extraction_Blank" & Positives > 3 ~ "Extraction_contamination",
      TRUE ~ "Pass"
    )
  )

# Flag and report QC failures
qc_failures <- qc_filtered %>%
  filter(qc_flag != "Pass")

if (nrow(qc_failures) > 0) {
  warning("QC failures detected:")
  print(qc_failures %>% select(Well, Sample, Target, Positives, qc_flag))
}

# Retain only passing sample wells
analysis_ready <- qc_filtered %>%
  filter(qc_flag == "Pass",
         !Sample %in% c("NTC", "Extraction_Blank", "Positive_Control"))
```


## Appendix B: Visualization Template {.appendix}

```{r}
#| label: gene-abundance-plot
#| echo: true
#| eval: false

library(tidyverse)

# Assumes analysis_ready has columns: sample_id, gene, copies_per_g_dry, treatment
analysis_ready %>%
  ggplot(aes(x = treatment, y = log10(copies_per_g_dry), fill = treatment)) +
  geom_boxplot(alpha = 0.7, outlier.shape = NA) +
  geom_jitter(width = 0.2, alpha = 0.5, size = 1.5) +
  facet_wrap(~ gene, scales = "free_y") +
  labs(
    x = "Treatment",
    y = expression(log[10]~"(gene copies g"^{-1}~"dry soil)"),
    title = "Nitrogen Cycle Gene Abundance by Treatment"
  ) +
  theme_minimal() +
  theme(legend.position = "none")
```


## Appendix C: Sending Samples to External Laboratories {.appendix}

*TBD -- if ddPCR analysis is outsourced to an external service provider
(e.g., university core facility, commercial lab), document the following:*

- *Preferred service provider(s) and contact information*
- *Sample submission requirements (volume, concentration, format)*
- *Data format and file types returned by the provider*
- *Turnaround time expectations*
- *Cost per sample and funding source/account number*
- *How to reconcile externally generated data with in-house QC standards*
