---
# =============================================================================
# SOP-L414: Soil Carbon and Nitrogen Analysis by Flash EA-IRMS
# =============================================================================
# Application SOP for d13C, d15N, %C, and %N determination in soil and
# sediment samples using the Flash IRMS EA coupled to the Delta V IRMS.
# Includes sample preparation, sequence design, and complete post-analysis
# data correction workflow with R code.
# =============================================================================

title: "Soil Carbon and Nitrogen Analysis by Flash EA-IRMS"
sop-number: "SOP-L414"
author:
  - name: "Marc Los Huertos"
    affiliation: "Environmental Analysis, Pomona College"
  - name: "Kyle McCarty"
    affiliation: "Oxtoby Isotope Lab, Pomona College"
date: last-modified
date-format: "YYYY-MM-DD"
revision: "1.0"

revision-history:
  - version: "1.0"
    date: "2026-05-18"
    author: "Marc Los Huertos"
    changes: "Initial release; migrated from LaTeX SOPs 75A/75B and SOP 75 v0.1"
---

## Purpose

This SOP describes the complete workflow for determining
$\delta^{13}$C, $\delta^{15}$N, and elemental concentrations (%C, %N)
in soil and sediment samples by combustion on the Thermo Scientific
Flash IRMS Elemental Analyzer (EA) coupled to the Delta V IRMS through
the ConFlo IV interface. It covers sample preparation, sequence design,
the analytical run, and post-analysis data correction including blank
correction, linearity correction, scale normalization, and drift
correction.

For general instrument operation, startup/shutdown, gas checks, tuning,
and pre-analysis tests, see
[SOP-L208](SOP-L208-irms-general-instrument.qmd).


## Scope

- **Applies to:** Soil, sediment, and plant tissue samples analyzed for
  $\delta^{13}$C, $\delta^{15}$N, %C, and %N by combustion EA-IRMS in
  NC (dual reactor) mode.
- **Does not apply to:** Sulfur isotope analysis (NCS mode; requires
  dedicated reactor and Sulfinert column), GasBench applications
  (see [SOP-L415](SOP-L415-carbonate-gasbench-irms.qmd)), TC/EA
  applications ($\delta^{2}$H, $\delta^{18}$O), or liquid samples.
- **Location:** Sample preparation in SGM Rm 133; analysis in the
  Oxtoby Isotope Lab, SGM Rm 135.


## Responsibilities

| Role              | Responsibility                                           |
|:------------------|:---------------------------------------------------------|
| Lab PI            | Approve methods, oversee data quality                    |
| Lab Manager       | Prepare/verify Isodat methods; maintain reactors         |
| Certified User    | All steps from sample prep through data reduction        |
| Student Researcher| May perform sample prep and data reduction under supervision |


## Safety and Hazards

See [SOP-L208, Section 4](SOP-L208-irms-general-instrument.qmd) for the
full hazard list. Hazards specific to this application:

- **Hot ash fingers and reactor tubes** during reactor inspection or
  exchange.
- **Chromium oxide, silvered cobaltous/cobaltic oxide, reduced copper
  wires** used in reactor packing. Consult SDS sheets.
- **Tin capsules:** small and sharp when crimped.

### Required PPE

- [ ] Safety glasses
- [ ] Lab coat
- [ ] Nitrile gloves


## Prerequisites

- Certification as an IRMS User per
  [SOP-L208](SOP-L208-irms-general-instrument.qmd)
- [SOP-L002 Laboratory Safety](SOP-L002-laboratory-safety.qmd)
- Familiarity with the Isodat software (see SOP-L208, Section 15)


## Sample Preparation

### Drying and Grinding

1. Weigh soil or sediment into an aluminum crucible.
2. Oven-dry at 105 °C until constant mass.
3. Grind dried samples to a fine, homogeneous powder using a mortar and
   pestle or ball mill.

### Elemental Pre-screening

Submit a subset of samples to elemental analysis (EA only, no IRMS) to
determine %C and %N. Use these results to calculate the target mass for
isotopic analysis. An online calculator is available at
<http://stableisotopefacility.ucdavis.edu/sample-weight-calculator.html>.

### Capsule Weighing

1. Tare the microbalance with an empty tin capsule.
2. Weigh the target mass of sample (typically 0.5--20 mg, depending on
   %C and %N) into the capsule.
3. If required by your method, add approximately 2 parts WO~3~ as a
   combustion aid.
4. Crimp the capsule tightly into a ball, eliminating all air pockets.
   A poorly sealed capsule introduces atmospheric N~2~.
5. Place each capsule into a labeled cell well.
6. Prepare standards and blanks identically. Plan for at least 2--3
   replicates of each bracketing standard per run.

::: {.callout-tip}
Maintain a detailed sample log with IDs, target masses, and actual
masses. This log is essential for data reduction.
:::


## Sequence Design

A well-designed sequence is critical for data quality. A typical layout:

| Position(s) | Contents                          | Purpose                      |
|:------------|:----------------------------------|:-----------------------------|
| 1--3        | Conditioning standard (e.g., LOCSS) | Stabilize reactor/column    |
| 4--5        | Blank (empty tin capsule)         | Blank determination          |
| 6--8        | Standard A (e.g., USGS40)        | Normalization anchor 1       |
| 9--11       | Standard B (e.g., USGS41a)       | Normalization anchor 2       |
| 12--19      | Unknown samples                   |                              |
| 20--22      | Check standard (e.g., LOCSS)     | Drift/QC monitor             |
| 23--30      | Unknown samples                   |                              |
| 31--33      | Check standard                    | Drift/QC monitor             |
| ...         | Continue pattern                  |                              |
| Last 3      | Standard A replicates             | End-of-run normalization check |
| Last 3      | Standard B replicates             | End-of-run normalization check |

::: {.callout-note}
Each autosampler carousel holds 32 wells. Up to three carousels can be
stacked for 94 positions total. A full NC dual run of 94 samples takes
roughly 12 hours.
:::


## Running the Analysis

1. Complete all pre-analysis checks per
   [SOP-L208, Section 13](SOP-L208-irms-general-instrument.qmd#sec-pre-analysis):
   gas checks, backgrounds, on/off tests, and blank determination.
2. Build the sequence in Isodat (see SOP-L33, Section 15).
3. Load autosampler carousels. Verify that physical positions match the
   sequence.
4. Click **Start**. Save as:
   `YYYYMMDD_Project_SequenceID_Username.seq`
5. Monitor the first several samples to confirm chromatographic
   separation and peak shapes are normal.

### Approximate Run Times

| Analysis              | Time per Sample (approx.) |
|:----------------------|:-------------------------:|
| $\delta^{15}$N only   | ~4--5 min                 |
| $\delta^{13}$C only   | ~5--6 min                 |
| NC dual method        | ~7--8 min                 |

Times include ~1 minute for peak centering and magnet switching.


## End-of-Run Shutdown and Cleanup

1. Confirm the sequence completed successfully.
2. Follow the shutdown procedure in SOP-L208, Section 11.3.
3. Log gas usage and any issues in the lab logbook.
4. Remove samples and personal items. Clean up. Push in your chair.


## Post-Analysis Data Correction

### Overview

Corrections are applied in the following order:

1. **Blank correction:** Subtract the blank contribution from peak
   areas.
2. **Linearity correction:** Adjust $\delta$ values for signal-size
   dependence.
3. **Scale normalization:** Calibrate to the international scale using
   bracketing standards.
4. **Drift correction:** Correct for temporal trends during the run.
5. **QC evaluation:** Verify standard reproducibility and accuracy.

::: {.callout-important}
Apply corrections in this order. Skipping or reordering steps can
introduce systematic error.
:::

### Export Raw Data

Export completed results from Isodat as CSV. Include: sample identifier,
peak areas (Area 28, Area 44), raw $\delta$ values vs. reference gas,
and signal amplitudes (Ampl 28, Ampl 44). Save as:

`YYYYMMDD_Project_SeqID_User_raw.csv`

### Import and Classify

```{r}
#| label: import-classify
#| echo: true
#| eval: false

library(tidyverse)

raw <- read_csv("data/raw/YYYYMMDD_Project_SeqID_User_raw.csv")

# Classify rows
raw <- raw %>%
  mutate(
    type = case_when(
      str_detect(Identifier1, regex("blank|BLK", ignore_case = TRUE)) ~ "blank",
      str_detect(Identifier1, "LOCSS|HOCSS|Urea|USGS")               ~ "standard",
      TRUE                                                             ~ "sample"
    )
  )

count(raw, type)
```

### Blank Correction

Empty tin capsules contribute a small amount of N~2~ and CO~2~. Subtract
the mean blank peak area from all samples.

```{r}
#| label: blank-correction
#| echo: true
#| eval: false

blank_means <- raw %>%
  filter(type == "blank") %>%
  summarize(
    Area28_blank = mean(Area28, na.rm = TRUE),
    Area44_blank = mean(Area44, na.rm = TRUE)
  )

cat(sprintf("Mean blank Area 28: %.1f mVs\n", blank_means$Area28_blank))
cat(sprintf("Mean blank Area 44: %.1f mVs\n", blank_means$Area44_blank))

raw <- raw %>%
  mutate(
    Area28_corr = Area28 - blank_means$Area28_blank,
    Area44_corr = Area44 - blank_means$Area44_blank
  )
```

::: {.callout-note}
If blank areas exceed 5% of the smallest sample peak area, investigate
contamination (leaky capsules, dirty carousel) before proceeding.
:::

### Linearity Correction

Use the slope from the pre-analysis on/off tests to adjust $\delta$
values for signal-size dependence.

```{r}
#| label: linearity-correction
#| echo: true
#| eval: false

# Enter on/off test data from your pre-analysis records
onoff <- tibble(
  amplitude = c(1000, 2000, 4000, 6000, 8000, 10000),   # mV
  d15N_raw  = c(-0.15, -0.10, -0.02, 0.03, 0.07, 0.12)  # per mil vs ref gas
)

lin_fit <- lm(d15N_raw ~ amplitude, data = onoff)
lin_slope <- coef(lin_fit)["amplitude"]

cat(sprintf("Linearity slope (d15N): %.6f per mil / mV\n", lin_slope))

# Reference amplitude = mean amplitude of your normalization standards
ref_ampl <- mean(raw$Ampl28[raw$type == "standard"], na.rm = TRUE)

raw <- raw %>%
  mutate(d15N_lin = d15N_raw - lin_slope * (Ampl28 - ref_ampl))

# Repeat for d13C using Ampl44 and the corresponding on/off data
```

### Scale Normalization (Two-Point)

Map measured $\delta$ values onto the international scale using two
reference materials with certified values.

```{r}
#| label: two-point-normalization
#| echo: true
#| eval: false

std_cert <- tibble(
  Identifier1 = c("USGS40", "USGS41a"),
  d13C_true   = c(-26.39, 36.55),
  d15N_true   = c(-4.52, 47.6)
)

std_meas <- raw %>%
  filter(Identifier1 %in% std_cert$Identifier1) %>%
  group_by(Identifier1) %>%
  summarize(
    d13C_meas = mean(d13C_lin, na.rm = TRUE),
    d15N_meas = mean(d15N_lin, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  left_join(std_cert, by = "Identifier1")

# Calibration regressions: true = slope * measured + intercept
fit_C <- lm(d13C_true ~ d13C_meas, data = std_meas)
fit_N <- lm(d15N_true ~ d15N_meas, data = std_meas)

cat(sprintf("d13C calibration: slope = %.4f, intercept = %.4f\n",
            coef(fit_C)[2], coef(fit_C)[1]))
cat(sprintf("d15N calibration: slope = %.4f, intercept = %.4f\n",
            coef(fit_N)[2], coef(fit_N)[1]))

raw <- raw %>%
  mutate(
    d13C_cal = coef(fit_C)[1] + coef(fit_C)[2] * d13C_lin,
    d15N_cal = coef(fit_N)[1] + coef(fit_N)[2] * d15N_lin
  )
```

::: {.callout-tip}
### Multi-Point Normalization
If three or more reference materials are available (e.g., USGS40,
USGS41a, USGS61), include all of them. This improves robustness and
provides a check on calibration linearity through the residuals.
:::

### Drift Correction

If check standards (e.g., LOCSS) show a systematic shift over the
course of the run, apply a time-based correction.

```{r}
#| label: drift-correction
#| echo: true
#| eval: false

checks <- raw %>%
  filter(Identifier1 == "LOCSS") %>%
  arrange(Row) %>%
  mutate(
    d13C_residual = d13C_cal - (-26.66),
    d15N_residual = d15N_cal - 7.30
  )

drift_C <- lm(d13C_residual ~ Row, data = checks)
drift_N <- lm(d15N_residual ~ Row, data = checks)

# Apply correction only if drift is statistically significant
if (summary(drift_C)$coefficients["Row", "Pr(>|t|)"] < 0.05) {
  cat(sprintf("Significant d13C drift: %.5f per mil / position\n",
              coef(drift_C)["Row"]))
  raw <- raw %>%
    mutate(d13C_final = d13C_cal -
             predict(drift_C, newdata = tibble(Row = Row)))
} else {
  cat("No significant d13C drift.\n")
  raw <- raw %>% mutate(d13C_final = d13C_cal)
}

if (summary(drift_N)$coefficients["Row", "Pr(>|t|)"] < 0.05) {
  cat(sprintf("Significant d15N drift: %.5f per mil / position\n",
              coef(drift_N)["Row"]))
  raw <- raw %>%
    mutate(d15N_final = d15N_cal -
             predict(drift_N, newdata = tibble(Row = Row)))
} else {
  cat("No significant d15N drift.\n")
  raw <- raw %>% mutate(d15N_final = d15N_cal)
}
```


## Elemental Concentrations (%C, %N)

Elemental concentrations are determined from blank-corrected peak areas
using a calibration curve built from standards of known %C and %N.

```{r}
#| label: elemental-concentrations
#| echo: true
#| eval: false

# Build area-to-mass calibration from standards with known %C and %N
std_elemental <- tibble(
  Identifier1 = c("LOCSS", "HOCSS", "Urea"),
  pctC_true   = c(2.28, 42.1, 20.0),   # replace with actual certified values
  pctN_true   = c(0.20, 4.6, 46.6),    # replace with actual certified values
  mass_mg     = c(10.0, 2.5, 1.5)      # replace with actual weighed masses
)

# Merge with measured areas
std_areas <- raw %>%
  filter(Identifier1 %in% std_elemental$Identifier1) %>%
  group_by(Identifier1) %>%
  summarize(
    Area44_mean = mean(Area44_corr, na.rm = TRUE),
    Area28_mean = mean(Area28_corr, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  left_join(std_elemental, by = "Identifier1") %>%
  mutate(
    C_mass_ug = pctC_true / 100 * mass_mg * 1000,
    N_mass_ug = pctN_true / 100 * mass_mg * 1000
  )

# Fit: area = slope * mass_ug
fit_pctC <- lm(Area44_mean ~ C_mass_ug, data = std_areas)
fit_pctN <- lm(Area28_mean ~ N_mass_ug, data = std_areas)

# Apply to samples
raw <- raw %>%
  mutate(
    C_mass_ug = (Area44_corr - coef(fit_pctC)[1]) / coef(fit_pctC)[2],
    N_mass_ug = (Area28_corr - coef(fit_pctN)[1]) / coef(fit_pctN)[2],
    pctC = C_mass_ug / (sample_mass_mg * 1000) * 100,
    pctN = N_mass_ug / (sample_mass_mg * 1000) * 100
  )
```

::: {.callout-note}
The `sample_mass_mg` column must come from your sample log. Merge it
with the Isodat export before running this step.
:::


## QC/QA Criteria

### Blanks

Blank peak areas must be negligible (<5% of the smallest sample peak).

### Standard Reproducibility

| Parameter        | Maximum 1$\sigma$ (per mil) |
|:-----------------|:---------------------------:|
| $\delta^{13}$C   | 0.10                        |
| $\delta^{15}$N   | 0.15                        |

### Calibration Accuracy

After normalization, the mean calibrated value of each reference
material should fall within 0.2 per mil of its certified value.

### Drift

If the difference between first and last check standards exceeds
0.3 per mil, apply a drift correction. If drift exceeds 0.5 per mil
after correction, re-analyze the affected portion.

### QC Summary Report

```{r}
#| label: qc-summary
#| echo: true
#| eval: false

qc <- raw %>%
  filter(type == "standard") %>%
  group_by(Identifier1) %>%
  summarize(
    n         = n(),
    d13C_mean = mean(d13C_final, na.rm = TRUE),
    d13C_sd   = sd(d13C_final, na.rm = TRUE),
    d15N_mean = mean(d15N_final, na.rm = TRUE),
    d15N_sd   = sd(d15N_final, na.rm = TRUE),
    .groups   = "drop"
  )

print(qc)

# Write final results
results <- raw %>%
  filter(type == "sample") %>%
  select(Identifier1, d13C_final, d15N_final, pctC, pctN)

write_csv(results,
  "data/processed/YYYYMMDD_Project_SeqID_calibrated.csv")
write_csv(qc,
  "data/processed/YYYYMMDD_Project_SeqID_qc_summary.csv")
```


## Reporting

Report final values as:

- $\delta^{13}$C vs. VPDB (Vienna Pee Dee Belemnite)
- $\delta^{15}$N vs. Air-N~2~

Include in any data table or publication:

- Reference materials used and their certified values
- Within-run precision (1$\sigma$) of each reference material
- Which corrections were applied (blank, linearity, drift)
- The calibration equation (slope and intercept)


## Troubleshooting

See [SOP-L208, Section 18](SOP-L208-irms-general-instrument.qmd) for the
full troubleshooting table. Issues specific to soil CN analysis:

| Symptom                              | Possible Cause / Action                       |
|:-------------------------------------|:----------------------------------------------|
| Noisy or double N~2~ peak            | Capsule not fully sealed; air leak             |
| CO~2~ peak tailing into next sample  | Column needs bake-out or replacement           |
| %C or %N values systematically low   | Sample mass error; re-check sample log         |
| High blank at m/z 28                 | Dirty carousel; residual sample in drop tube   |
| Standards acceptable, samples erratic | Sample heterogeneity; grind finer or increase mass |


## References

Coplen, T. B. (2011). Guidelines and recommended terms for expression
of stable-isotope-ratio and gas-ratio measurement results. *Rapid
Communications in Mass Spectrometry*, 25(17), 2538--2560.

Paul, D., Skrzypek, G., & Forizs, I. (2007). Normalization of measured
stable isotopic compositions to isotope reference scales -- a review.
*Rapid Communications in Mass Spectrometry*, 21(18), 3006--3014.

Werner, R. A., & Brand, W. A. (2001). Referencing strategies and
techniques in stable isotope ratio analysis. *Rapid Communications in
Mass Spectrometry*, 15(7), 501--519.


## Related SOPs

| SOP   | Title                                       | Relationship                      |
|:------|:--------------------------------------------|:----------------------------------|
| [L002](SOP-L002-laboratory-safety.qmd)       | Laboratory Safety       | Required prerequisite             |
| [L802](SOP-L802-quality-control.qmd)         | Quality Control and Quality Assurance | General QC framework   |
| [L208](SOP-L208-irms-general-instrument.qmd) | IRMS -- General Instrument Use | Instrument-level procedures |
| [L415](SOP-L415-carbonate-gasbench-irms.qmd) | Carbonate Isotopes by GasBench-IRMS | Companion application SOP |


## Training Record

Before performing soil CN analysis independently, personnel must:

1. Read this SOP and [SOP-L208](SOP-L208-irms-general-instrument.qmd) in
   full
2. Be a certified IRMS User (SOP-L208, Section 6)
3. Demonstrate capsule preparation under supervision
4. Perform a supervised analytical run with acceptable QC results
5. Complete at least one supervised data reduction session
6. Be signed off by the lab manager or PI
