---
# =============================================================================
# SOP-L803: Data Entry and Database Management
# =============================================================================
title: "Data Entry and Database Management"
sop-number: "SOP-L803"
author:
  - name: "Marc Los Huertos"
    affiliation: "EA Program, Pomona College"
date: last-modified
date-format: "YYYY-MM-DD"
revision: "0.1"

revision-history:
  - version: "0.1"
    date: "2026-05-21"
    author: "Marc Los Huertos"
    changes: "Initial release"
---

## 1. Purpose

This SOP describes standardized procedures for entering, organizing,
validating, and managing research data in the EA Program. It covers
the full data lifecycle from initial recording through archival storage,
ensuring that all datasets are consistent, traceable, and reproducible
[@broman2018; @wickham2014].


## 2. Scope

- **Applies to:** All tabular data generated in the EA Program
  laboratory and field campaigns, including instrument outputs, field
  observations, and manually recorded measurements
- **Does not apply to:** Spatial/GIS data (see
  [SOP-L804](SOP-L804-gis-spatial-data.qmd)), bioinformatics sequence
  data (see [SOP-L705](SOP-L705-bioinformatics-pipeline.qmd)), or
  instrument-specific export procedures (see
  [SOP-L807](SOP-L807-instrument-data-export.qmd))
- **Location:** EA Program laboratory (SGM 133) and RStudio Server


## 3. Responsibilities

| Role              | Responsibility                              |
|:------------------|:--------------------------------------------|
| Lab PI            | Approve data management plan; ensure compliance with funder requirements |
| Lab Manager       | Maintain database templates; perform quarterly backups |
| Analyst / Student | Enter data following this SOP; run validation checks; document metadata |
| Data Reviewer     | Verify data entries against lab notebooks before analysis |


## 4. Safety and Hazards

Data entry is a computational activity with no physical hazards.
Ergonomic best practices (regular breaks, proper screen height) are
encouraged for extended data entry sessions.


## 5. Equipment and Materials

### Software

| Item                        | Access                | Notes                        |
|:----------------------------|:----------------------|:-----------------------------|
| RStudio Server              | rstudio.pomona.edu    | Primary data processing tool |
| R (version 4.x)             | Pre-installed on server | Required for validation scripts |
| Git / GitHub                | github.com/marclos    | Version control for all data files |
| Microsoft Excel / Google Sheets | Campus license   | Acceptable for initial entry only |
| Text editor (VS Code, nano) | Various               | For editing CSV files directly |

### File Storage

| Location                    | Purpose               | Backup Frequency             |
|:----------------------------|:----------------------|:-----------------------------|
| GitHub repository           | Active project data   | Every commit                 |
| College network drive       | Archival storage      | Semester end                 |
| External hard drive (lab)   | Physical backup       | Semester end                 |


## 6. Procedure

### 6.1 File Naming Conventions

All data files must follow this naming pattern:

```
[project]_[datatype]_[YYYY-MM-DD].[ext]
```

For example: `claremont_soilpH_2026-05-21.csv`

Rules for file names:

1. Use lowercase letters only
2. Separate words with underscores (not spaces or hyphens)
3. Include the date of collection or generation in ISO 8601 format
   (YYYY-MM-DD)
4. Use `.csv` as the default file format for tabular data

::: {.callout-tip}
### Tip
Never use spaces, special characters, or accented letters in file names.
These cause problems across operating systems and with R's file-reading
functions.
:::

### 6.2 Directory Structure

Organize project data using a consistent directory tree:

```
project_name/
├── data/
│   ├── raw/          # Unmodified instrument or field data
│   ├── processed/    # Cleaned, validated datasets
│   └── metadata/     # Data dictionaries, codebooks
├── analysis/
│   ├── scripts/      # R scripts and Quarto documents
│   └── output/       # Figures, tables, reports
└── README.md         # Project overview and data description
```

Raw data files in `data/raw/` must never be modified. All cleaning and
transformation steps should produce new files in `data/processed/` with
scripts that document every change [@wilson2017].

### 6.3 Data Entry from Lab Notebooks

1. Open the lab notebook to the relevant page
2. Create a new CSV file (or open the project template) in `data/raw/`
3. Enter column headers in the first row following the variable naming
   conventions below
4. Enter each observation as a separate row
5. Record the notebook page number, date, and analyst initials in the
   metadata file
6. Save and commit to Git with a descriptive message

### 6.4 Variable Naming Conventions

Column headers (variable names) must be:

1. Lowercase, with underscores separating words (e.g., `soil_ph`,
   `no3_mg_l`)
2. Free of spaces, special characters, and units in the name itself
3. Documented in an accompanying data dictionary (see Section 6.7)

Use consistent naming across related datasets. For example, always use
`site_id` (not `Site`, `site`, `SiteID`, or `location`) when referring
to the same sampling locations.

### 6.5 Data Types and Formatting

| Data Type       | Convention                  | Example               |
|:----------------|:----------------------------|:----------------------|
| Dates           | ISO 8601 (YYYY-MM-DD)       | `2026-05-21`          |
| Times           | 24-hour (HH:MM)             | `14:30`               |
| Missing values  | Leave cell empty or use `NA`| `NA`                  |
| Numeric         | No commas; use period for decimal | `1234.56`        |
| Categorical     | Consistent spelling; lowercase | `upstream`          |
| Below detection | Flag column; numeric = MDL/2 | See QC SOP           |

::: {.callout-warning}
### Common Mistakes

Never use `0` to represent a missing value. A zero is a valid
measurement. Use `NA` or leave the cell blank. Never merge cells in a
spreadsheet, and never use color coding as the sole carrier of
information [@broman2018].
:::

### 6.6 Data Validation in R

After entry, run the validation script to check for common errors.
A minimal validation template:

```{r}
#| label: validate-data
#| echo: true
#| eval: false

library(tidyverse)

# Read the raw data
dat <- read_csv("data/raw/project_variable_2026-05-21.csv")

# --- Structure checks ---
glimpse(dat)                         # Column types correct?
stopifnot(nrow(dat) > 0)            # File is not empty

# --- Missing data summary ---
dat %>%
  summarize(across(everything(), ~ sum(is.na(.)))) %>%
  pivot_longer(everything(), names_to = "variable",
               values_to = "n_missing") %>%
  filter(n_missing > 0)

# --- Range checks ---
dat %>%
  summarize(
    ph_min = min(soil_ph, na.rm = TRUE),
    ph_max = max(soil_ph, na.rm = TRUE)
  )
# Are values within physically plausible ranges?

# --- Duplicate check ---
dat %>%
  group_by(site_id, date) %>%
  filter(n() > 1)
# Flag unexpected duplicates for review

# --- Categorical consistency ---
dat %>% count(site_id, sort = TRUE)
# Look for typos (e.g., "upstrream" vs "upstream")
```

### 6.7 Data Dictionaries (Metadata)

Every dataset must have a companion data dictionary saved as
`data/metadata/[project]_datadictionary.csv` with the following
columns:

| Column         | Description                                   |
|:---------------|:----------------------------------------------|
| `variable`     | Column name in the data file                  |
| `description`  | Plain-language description                    |
| `units`        | Units of measurement (or "categorical", "text") |
| `type`         | Data type: numeric, integer, character, date, logical |
| `range`        | Expected range or valid categories            |
| `source_sop`   | SOP number for the method that generated the data |

### 6.8 Version Control with Git

All data files and scripts must be tracked in a Git repository
hosted on GitHub. Follow the procedures in
[SOP-L801](SOP-L801-rstudio-github.qmd) for committing and pushing
changes. Key rules:

1. Commit raw data files when they are first created (and never modify
   them after the initial commit)
2. Commit processed data files after each cleaning or transformation
   step
3. Write informative commit messages (e.g., "Add soil pH data from
   Apr 15 sampling" rather than "update")
4. Use branches for experimental analyses; merge to `main` when
   finalized

### 6.9 Database Management for Multi-Year Projects

For projects spanning multiple semesters or involving multiple
analysts, maintain a relational data structure:

1. Create a **sites table** (`data/metadata/sites.csv`) with location
   identifiers, coordinates, and site descriptions
2. Create a **samples table** (`data/metadata/samples.csv`) linking
   each sample to a site, date, and collector
3. Link analytical results to samples using unique `sample_id` values

This structure prevents duplication, supports merging across
semesters, and enables relational joins in R [@wickham2014]:

```{r}
#| label: relational-join
#| echo: true
#| eval: false

sites <- read_csv("data/metadata/sites.csv")
samples <- read_csv("data/metadata/samples.csv")
results <- read_csv("data/processed/chemistry_results.csv")

# Join all tables
full_data <- results %>%
  left_join(samples, by = "sample_id") %>%
  left_join(sites, by = "site_id")
```


## 7. Quality Control and Acceptance Criteria

- All data entries must be verified against the original lab notebook
  by a second person (or by the analyst after a 24-hour waiting period)
- Validation scripts must pass with zero unexpected errors before data
  are used in analysis
- The data dictionary must be complete for every variable in the
  dataset
- Refer to [SOP-L802](SOP-L802-quality-control.qmd) for analytical
  QC requirements that apply to the underlying measurements


## 8. Waste Disposal

Not applicable (computational procedure).


## 9. References

::: {#refs}
:::


## 10. Training Record

Training is documented in the EA Program training log. Before working
independently with project data, personnel must:

1. Read this SOP in full
2. Complete the RStudio and GitHub orientation
   ([SOP-L801](SOP-L801-rstudio-github.qmd))
3. Successfully enter, validate, and commit a practice dataset under
   supervision
4. Be signed off by the PI or lab manager


## Appendix A: CSV Template

A minimal CSV template for a typical water chemistry dataset:

```
sample_id,site_id,date,time,analyst,ph,conductivity_us_cm,temp_c,notes
W001,site_a,2026-05-21,09:30,MLH,7.2,345,18.5,
W002,site_b,2026-05-21,10:15,MLH,6.8,412,19.1,slight turbidity
```

## Appendix B: Recommended R Packages for Data Management

| Package        | Purpose                                       |
|:---------------|:----------------------------------------------|
| `tidyverse`    | Data import, cleaning, and transformation     |
| `janitor`      | Consistent column names, tabulation           |
| `validate`     | Rule-based data validation                    |
| `pointblank`   | Automated data quality reporting              |
| `here`         | Portable file paths across systems            |
| `skimr`        | Quick summary statistics                      |
