---
# =============================================================================
# SOP-L806: Statistical Analysis Templates (R)
# =============================================================================
title: "Statistical Analysis Templates (R)"
sop-number: "SOP-L806"
author:
  - name: "Marc Los Huertos"
    affiliation: "EA Program, Pomona College"
date: last-modified
date-format: "YYYY-MM-DD"
revision: "0.1"

revision-history:
  - version: "0.1"
    date: "2026-05-21"
    author: "Marc Los Huertos"
    changes: "Initial release"
---

## Purpose

This SOP provides standardized R analysis templates for common
statistical procedures used in EA Program research. It establishes
reproducible workflows for exploratory data analysis, hypothesis
testing, regression modeling, and reporting, ensuring that all
analyses are transparent, well-documented, and consistent across
projects [@zuur2010; @wickham2023].


## Scope

- **Applies to:** All quantitative statistical analyses performed
  for EA Program research, coursework, and publications
- **Does not apply to:** Bioinformatics sequence analysis (see
  [SOP-L705](SOP-L705-bioinformatics-pipeline.qmd)); GIS-specific
  spatial statistics (see [SOP-L804](SOP-L804-gis-spatial-data.qmd));
  instrument calibration curve fitting (see method-specific SOPs)
- **Location:** RStudio Server (rstudio.pomona.edu)


## Responsibilities

| Role              | Responsibility                              |
|:------------------|:--------------------------------------------|
| Lab PI            | Approve analytical approach; review statistical methods in manuscripts |
| Analyst / Student | Select appropriate methods; follow templates; document all decisions |
| Data Reviewer     | Verify that assumptions are checked and reported |
| Statistical Advisor | Consult on complex designs (mixed models, multivariate methods) |

<!-- Definitions snippet for SOP-L806: Statistical Analysis Templates R -->
<!-- Insert immediately before ## Safety and Hazards -->

## Definitions

- **ANOVA (Analysis of Variance):** A statistical test that partitions total variance in a response variable into components attributable to one or more categorical factors; tests whether group means differ significantly; assumes normality and homogeneity of variance.
- **Confidence interval:** A range of values computed from sample data that, under repeated sampling, contains the true population parameter with a stated probability (e.g., 95%); wider intervals reflect greater uncertainty.
- **Effect size:** A standardized measure of the magnitude of a difference or relationship, independent of sample size; examples include Cohen's d (mean differences), $r^2$ (explained variance), and Hedge's g; should always accompany p-values.
- **Mixed model:** A statistical model containing both fixed effects (factors whose levels are deliberately chosen) and random effects (factors whose levels represent a random sample from a larger population); appropriate when data have hierarchical or repeated-measures structure.
- **p-value:** The probability of observing a test statistic as extreme as the one obtained (or more extreme) if the null hypothesis is true; does not measure the probability that the null hypothesis is true; must be interpreted with effect size.
- **Power:** The probability of correctly rejecting the null hypothesis when it is false; increases with sample size, effect size, and significance level; should be considered in study design to ensure adequate sample sizes.
- **Residual:** The difference between an observed value and the value predicted by a statistical model; examination of residuals is the primary diagnostic for checking model assumptions.


## Safety and Hazards

Statistical analysis is a computational activity with no physical
hazards.


## Equipment and Materials

### Software

| Item                        | Access                | Notes                        |
|:----------------------------|:----------------------|:-----------------------------|
| R (version 4.x)            | RStudio Server        | Primary analysis platform    |
| RStudio IDE                 | rstudio.pomona.edu    | Integrated development environment |
| Quarto                      | Pre-installed         | Report generation            |
| Git / GitHub                | github.com/marclos    | Version control              |

### Core R Packages

| Package          | Purpose                                     |
|:-----------------|:--------------------------------------------|
| `tidyverse`      | Data wrangling and visualization            |
| `broom`          | Tidy model output                           |
| `car`            | Diagnostic tests (Levene, VIF)              |
| `lme4`           | Mixed-effects models                        |
| `emmeans`        | Estimated marginal means and contrasts      |
| `vegan`          | Community ecology and ordination            |
| `MuMIn`          | Model selection and multi-model inference   |
| `performance`    | Model diagnostics and assumption checks     |
| `report`         | Automated statistical reporting             |
| `ggplot2`        | Publication-quality figures                 |
| `knitr` / `kableExtra` | Formatted tables                      |


## Procedure

### Analysis Workflow Overview

Every statistical analysis should follow this sequence:

1. State the research question and hypotheses
2. Import and validate data (see [SOP-L803](SOP-L803-data-entry-database.qmd))
3. Conduct exploratory data analysis (EDA)
4. Check assumptions for the intended statistical test
5. Fit the model
6. Evaluate model diagnostics
7. Interpret results and generate figures/tables
8. Write up methods and results in a Quarto document
9. Commit the complete analysis to Git

### Quarto Document Header

Begin every analysis with a standard Quarto header:

```yaml
---
title: "Analysis: [Project Name]"
author: "[Your Name]"
date: today
format:
  html:
    toc: true
    code-fold: true
    embed-resources: true
bibliography: references.bib
---
```

### Exploratory Data Analysis Template

```{r}
#| label: eda-template
#| echo: true
#| eval: false

library(tidyverse)

# --- 1. Load and inspect data ---
dat <- read_csv("data/processed/project_data.csv")
glimpse(dat)

# --- 2. Summary statistics ---
dat %>%
  group_by(treatment) %>%
  summarize(
    n = n(),
    mean = mean(response, na.rm = TRUE),
    sd = sd(response, na.rm = TRUE),
    median = median(response, na.rm = TRUE),
    min = min(response, na.rm = TRUE),
    max = max(response, na.rm = TRUE)
  )

# --- 3. Distributional plots ---
ggplot(dat, aes(x = response)) +
  geom_histogram(bins = 20, fill = "steelblue", color = "white") +
  facet_wrap(~ treatment) +
  labs(x = "Response Variable (units)", y = "Count") +
  theme_minimal()

ggplot(dat, aes(x = treatment, y = response)) +
  geom_boxplot(fill = "steelblue", alpha = 0.5) +
  geom_jitter(width = 0.1, alpha = 0.4) +
  labs(x = "Treatment", y = "Response Variable (units)") +
  theme_minimal()

# --- 4. Missing data assessment ---
dat %>%
  summarize(across(everything(), ~ sum(is.na(.)))) %>%
  pivot_longer(everything(), names_to = "variable",
               values_to = "n_missing") %>%
  filter(n_missing > 0)
```

### Two-Sample Comparisons

```{r}
#| label: two-sample
#| echo: true
#| eval: false

# --- Check normality within groups ---
dat %>%
  group_by(treatment) %>%
  summarize(
    shapiro_p = shapiro.test(response)$p.value
  )

# --- Check equal variances ---
car::leveneTest(response ~ treatment, data = dat)

# --- Parametric: two-sample t-test ---
t_result <- t.test(response ~ treatment, data = dat, var.equal = FALSE)
t_result

# --- Nonparametric alternative: Wilcoxon rank-sum ---
wilcox.test(response ~ treatment, data = dat)

# --- Effect size ---
library(effectsize)
cohens_d(response ~ treatment, data = dat)
```

### One-Way ANOVA Template

```{r}
#| label: anova-template
#| echo: true
#| eval: false

# --- Fit ANOVA ---
aov_model <- aov(response ~ treatment, data = dat)

# --- Check assumptions ---
# Normality of residuals
shapiro.test(residuals(aov_model))
qqnorm(residuals(aov_model)); qqline(residuals(aov_model))

# Homogeneity of variance
car::leveneTest(response ~ treatment, data = dat)

# Residual vs. fitted plot
plot(aov_model, which = 1)

# --- ANOVA table ---
summary(aov_model)

# --- Post-hoc pairwise comparisons ---
library(emmeans)
emm <- emmeans(aov_model, ~ treatment)
pairs(emm, adjust = "tukey")

# --- Nonparametric alternative: Kruskal-Wallis ---
kruskal.test(response ~ treatment, data = dat)
```

::: {.callout-tip}
### Tip
Always check assumptions before interpreting p-values. If residuals
are non-normal or variances are unequal, consider a transformation
(log, square root) or switch to a nonparametric test. Document the
reasoning in your analysis script [@zuur2010].
:::

### Linear Regression Template

```{r}
#| label: regression-template
#| echo: true
#| eval: false

# --- Fit linear model ---
lm_model <- lm(response ~ predictor1 + predictor2, data = dat)

# --- Model summary ---
summary(lm_model)

# --- Tidy output ---
library(broom)
tidy(lm_model, conf.int = TRUE)
glance(lm_model)

# --- Diagnostics ---
par(mfrow = c(2, 2))
plot(lm_model)
par(mfrow = c(1, 1))

# Check multicollinearity
car::vif(lm_model)

# --- Model performance ---
library(performance)
check_model(lm_model)

# --- Visualization ---
ggplot(dat, aes(x = predictor1, y = response)) +
  geom_point(alpha = 0.6) +
  geom_smooth(method = "lm", se = TRUE, color = "steelblue") +
  labs(x = "Predictor (units)", y = "Response (units)",
       caption = sprintf("R-squared = %.3f, p = %.4f",
                         summary(lm_model)$r.squared,
                         glance(lm_model)$p.value)) +
  theme_minimal()
```

### Mixed-Effects Model Template

For data with nested or repeated-measures structure:

```{r}
#| label: mixed-model
#| echo: true
#| eval: false

library(lme4)
library(lmerTest)  # For p-values via Satterthwaite

# --- Fit mixed model ---
# Fixed effect: treatment
# Random effect: site (multiple samples per site)
mixed_model <- lmer(response ~ treatment + (1 | site_id), data = dat)

# --- Summary with p-values ---
summary(mixed_model)

# --- Marginal and conditional R-squared ---
MuMIn::r.squaredGLMM(mixed_model)

# --- Post-hoc comparisons ---
library(emmeans)
emm <- emmeans(mixed_model, ~ treatment)
pairs(emm, adjust = "tukey")

# --- Residual diagnostics ---
plot(mixed_model)
qqnorm(resid(mixed_model)); qqline(resid(mixed_model))
```

### Multivariate / Community Analysis Template

```{r}
#| label: multivariate
#| echo: true
#| eval: false

library(vegan)

# --- Prepare community matrix ---
# Rows = sites, columns = species/variables
comm_matrix <- dat %>%
  select(site_id, species, abundance) %>%
  pivot_wider(names_from = species, values_from = abundance,
              values_fill = 0) %>%
  column_to_rownames("site_id")

# --- Diversity indices ---
diversity_h <- diversity(comm_matrix, index = "shannon")
richness <- specnumber(comm_matrix)

# --- Ordination (NMDS) ---
nmds <- metaMDS(comm_matrix, distance = "bray", k = 2,
                trymax = 100)
stressplot(nmds)

# Plot
ordiplot(nmds, type = "n")
points(nmds, col = as.factor(env_data$treatment), pch = 16)
legend("topright", legend = levels(as.factor(env_data$treatment)),
       col = 1:nlevels(as.factor(env_data$treatment)), pch = 16)

# --- PERMANOVA ---
adonis2(comm_matrix ~ treatment, data = env_data,
        method = "bray", permutations = 999)
```

### Figure Formatting Standards

All figures for reports and publications must follow these conventions:

1. Use `theme_minimal()` or a consistent custom theme
2. Label axes with variable names and units
3. Use colorblind-friendly palettes (e.g., `viridis`, `RColorBrewer`
   "Set2")
4. Include figure captions using the `#| fig-cap:` chunk option
5. Save publication figures at 300 DPI as both PDF and PNG:

```{r}
#| label: save-figure
#| echo: true
#| eval: false

ggsave("analysis/output/fig_treatment_response.pdf",
       width = 6, height = 4, units = "in")
ggsave("analysis/output/fig_treatment_response.png",
       width = 6, height = 4, units = "in", dpi = 300)
```

### Reporting Statistical Results

When writing results in a Quarto document, follow APA conventions:

- Report test statistics, degrees of freedom, and p-values:
  "Treatment had a significant effect on soil pH (F(2, 27) = 4.53,
  p = 0.020)"
- Report effect sizes alongside p-values
- Include confidence intervals for parameter estimates
- Round to a consistent number of decimal places (typically 2 for
  test statistics, 3 for p-values)
- Use `broom::tidy()` and `knitr::kable()` for formatted tables

```{r}
#| label: results-table
#| echo: true
#| eval: false

library(broom)
library(knitr)

tidy(aov_model) %>%
  kable(digits = 3, caption = "ANOVA results for soil pH by treatment")
```


## Quality Control and Acceptance Criteria

- Every analysis must include explicit assumption checks (normality,
  homoscedasticity, independence) documented in the script
- All models must include diagnostic plots
- Results must be reproducible: running the script from start to
  finish on the same data must produce identical output
- Analyses must be committed to Git with clear commit messages
- The PI or a designated reviewer must approve the analytical approach
  before results are included in publications


## Waste Disposal

Not applicable (computational procedure).


## References

::: {#refs}
:::


## Training Record

Training is documented in the EA Program training log. Before
conducting statistical analyses independently, personnel must:

1. Read this SOP in full
2. Complete the RStudio and GitHub orientation
   ([SOP-L801](SOP-L801-rstudio-github.qmd))
3. Complete the introductory R statistics module (see
   [RTricks](https://marclos.github.io/RTricks/))
4. Demonstrate a complete analysis workflow (EDA through reporting)
   on a practice dataset under supervision
5. Be signed off by the PI


## Appendix A: Decision Guide for Choosing a Statistical Test

| Research Question                | Data Type          | Groups | Recommended Test         |
|:---------------------------------|:-------------------|:-------|:-------------------------|
| Compare 2 group means            | Continuous, normal | 2      | Two-sample t-test        |
| Compare 2 group medians          | Non-normal         | 2      | Wilcoxon rank-sum        |
| Compare 3+ group means           | Continuous, normal | 3+     | One-way ANOVA            |
| Compare 3+ group medians         | Non-normal         | 3+     | Kruskal-Wallis           |
| Relationship between 2 variables | Both continuous    | N/A    | Linear regression / correlation |
| Multiple predictors              | Continuous response| N/A    | Multiple regression      |
| Nested / repeated measures       | Continuous         | Any    | Mixed-effects model      |
| Community composition            | Abundance matrix   | Any    | NMDS + PERMANOVA         |
| Categorical association          | Counts             | Any    | Chi-squared / Fisher's exact |

## Appendix B: Common Data Transformations

| Transformation      | R Code                   | When to Use                     |
|:---------------------|:-------------------------|:--------------------------------|
| Log (natural)       | `log(x)`                 | Right-skewed data; multiplicative relationships |
| Log (base 10)       | `log10(x)`               | Concentration data spanning orders of magnitude |
| Square root         | `sqrt(x)`                | Count data; moderate skew       |
| Arcsine square root | `asin(sqrt(x))`          | Proportions (0 to 1)           |
| Box-Cox             | `MASS::boxcox(model)`    | Let the data determine the best transformation |

::: {.callout-warning}
### Caution
Always report whether data were transformed and provide results on both
the transformed and original scales when possible. Transformations
change the hypothesis being tested (e.g., log-transformed ANOVA
compares geometric means, not arithmetic means).
:::
