---
# =============================================================================
# SOP-L702: ddPCR Applications Guide -- Absolute Quantification of
#           Soil Nitrogen-Cycle Functional Genes
# =============================================================================
title: "ddPCR Applications Guide: Absolute Quantification of Soil Nitrogen-Cycle Functional Genes"
sop-number: "SOP-L702"
author:
  - name: "Marc Los Huertos"
    affiliation: "EA Program, Pomona College"
date: last-modified
date-format: "YYYY-MM-DD"
revision: "0.2"
bibliography: references.bib
link-citations: true

revision-history:
  - version: "0.1"
    date: "2025-05-14"
    author: "Marc Los Huertos"
    changes: "Initial draft (procedural: data export and analysis)"
  - version: "0.2"
    date: last-modified
    author: "Marc Los Huertos"
    changes: >
      Reframed from a step-by-step data-analysis SOP into an applications
      guide. Instrument operation moved to SOP-L214. Added scientific framing
      for the soil nitrogen-cycle application, gene-target reference table,
      and a bibliography file. Corrected internal references to SOP-L700.
---

## 1. Purpose

This document is an applications guide, not a step-by-step bench protocol. It
explains *why* and *how* droplet digital PCR (ddPCR) on the Bio-Rad QX200
system is used in this lab to make absolute measurements of nitrogen-cycle
functional gene abundance in soil, and how to turn the instrument's output
into ecologically meaningful, reportable results.

It connects three other procedures: instrument operation
([SOP-L214 QX200 Instrument Familiarization](SOP-L214-qx200-instrument-familiarization.qmd)),
assay and primer design
([SOP-L701](SOP-L701-ddpcr-nitrogen-cycle-primers.qmd)), and the reproducible
analysis scripts written in R [@rcoreteam2024]. The technical workflow of the
QX200 platform follows Bio-Rad's published guidance [@biorad_appsguide2018],
and quantification rests on the absolute-counting principle introduced with the
droplet platform [@hindson2011].

::: {.callout-note}
For the original "absolute quantification" application as Bio-Rad documents it,
see Chapter 3 of the Droplet Digital PCR Applications Guide
[@biorad_appsguide2018]. This SOP adapts that general application to the
specific case of soil functional-gene quantification.
:::


## 2. Scope

- **Applies to:** Interpretation and analysis of QX200 ddPCR data
  (QuantaSoft or QuantaSoft Analysis Pro output) for nitrogen-cycle functional
  genes (*amoA*, *nirK*, *nirS*, *nosZ*, *nifH*, *nrfA*) from environmental
  soil samples, quantified in absolute units (copies per gram dry soil or
  copies per nanogram DNA).
- **Does not apply to:** Operating the instrument (see SOP-L214); assay or
  primer design and validation (see SOP-L701); amplicon sequencing,
  metagenomics, or qPCR data from other platforms.
- **Location:** Instrument computer (QuantaSoft) for export and threshold
  review; RStudio Server for analysis
  (see [SOP-L801 RStudio and GitHub](SOP-L801-rstudio-github.qmd)).


## 3. Responsibilities

| Role              | Responsibility                                |
|:------------------|:----------------------------------------------|
| Lab PI            | Approve analysis pipeline and interpretation  |
| Trained Personnel | Export data, run QC, perform analysis         |
| New Students      | Complete training before independent analysis |


## 4. Why ddPCR for This Application

Quantitative PCR (qPCR) infers target abundance from amplification kinetics
against a standard curve, so its accuracy depends on amplification efficiency
and on the quality of the curve. ddPCR instead partitions each reaction into
roughly 20,000 nanoliter droplets, amplifies the target within each droplet to
endpoint, and counts droplets as positive or negative. The fraction of
positive droplets is fit to a Poisson distribution to recover the starting
concentration in copies per microliter, with no standard curve required
[@hindson2011; @biorad_appsguide2018].

For soil nitrogen-cycle work, this matters for several reasons:

- **Absolute, comparable counts.** Copy numbers are returned directly in
  copies per microliter, which can be normalized to soil mass or DNA and
  compared across plates, runs, and studies without a shared standard curve.
- **Tolerance of inhibitors.** Soil DNA extracts carry humic acids and other
  PCR inhibitors. Endpoint partitioning is less sensitive to partial
  inhibition than the kinetic measurement that qPCR depends on, although
  inhibition should still be checked (Section 6).
- **Precision for small differences.** The large number of partitions lets
  ddPCR resolve modest fold-changes in gene abundance between treatments
  [@hindson2011].
- **Low-abundance targets.** Genes such as *nosZ* or *nifH* are often present
  at low copy numbers relative to bulk DNA; partitioning improves the
  signal-to-noise ratio for rare targets.

These properties are the reason functional-gene abundance is increasingly
reported in absolute units when relating microbial communities to nitrogen
cycling processes [@levybooth2014].


## 5. The Soil Nitrogen Cycle and Functional Gene Markers

Soil nitrogen transformations are carried out by distinct microbial guilds,
each defined by the enzymes (and therefore the genes) it carries. Quantifying
these marker genes provides a direct, culture-independent estimate of the
genetic potential for each process [@kuypers2018; @levybooth2014]. The targets
used in this lab, the processes they mark, and the primary primer references
are summarized below.

| Gene        | Process / guild                              | What the abundance indicates                                  | Primary reference   |
|:------------|:---------------------------------------------|:--------------------------------------------------------------|:--------------------|
| *amoA* (AOB)| Ammonia oxidation, bacterial (nitrification) | Bacterial ammonia-oxidizer potential; first, rate-limiting step of nitrification | @rotthauwe1997 |
| *amoA* (AOA)| Ammonia oxidation, archaeal (nitrification)  | Archaeal ammonia-oxidizer potential; often dominant in low-N, acidic soils | @francis2005 |
| *nirK*      | Nitrite reduction (denitrification)          | Cu-type nitrite-reductase potential; one of two parallel denitrification pathways | @henry2004; @throback2004 |
| *nirS*      | Nitrite reduction (denitrification)          | cd1-type nitrite-reductase potential; the other nitrite-reductase pathway | @throback2004 |
| *nosZ*      | Nitrous oxide reduction (denitrification)    | N~2~O-reductase potential; capacity to reduce N~2~O to N~2~    | @henry2006; @throback2004 |
| *nifH*      | Nitrogen fixation                            | Diazotroph potential; biological N~2~ fixation                | @levybooth2014      |
| *nrfA*      | Dissimilatory nitrate reduction to ammonium  | DNRA potential; retains N in the system as NH~4~^+^ rather than losing it as gas | @levybooth2014 |

::: {.callout-note}
The exact primer sets, probe chemistry (hydrolysis probe vs. EvaGreen), and
validated annealing conditions for each assay are defined in SOP-L701. The
references above identify the canonical marker-gene primers; cite the specific
assay actually used from SOP-L701 in any manuscript. Gene-specific *nifH* and
*nrfA* primer references should be added to SOP-L701 and this table updated to
match.
:::

Read together, these genes describe the competing fates of nitrogen in soil:
inputs by fixation (*nifH*), oxidation of ammonium (*amoA*), and the branch
point between gaseous loss through denitrification (*nirK*, *nirS*, *nosZ*) and
retention through DNRA (*nrfA*). Interpreting changes in one gene in isolation
is rarely informative; the ratios and combinations in Section 7 are usually
more meaningful.


## 6. From Instrument Output to Quality-Controlled Data

### 6.1 Reviewing and Exporting Data

Instrument operation and the QuantaSoft interface are covered in SOP-L214. For
analysis, the entry point is the exported well-level data.

1. Open the completed experiment file (`.qlp`) in QuantaSoft and confirm the
   positive/negative threshold for each well (see SOP-L214). Re-apply
   thresholds manually in QuantaSoft Analysis Pro if needed, and record any
   change in the lab notebook.
2. Export with **Options > Export CSV** to produce a well-level results file.
   Key columns include `Well`, `Sample`, `Target`, `Concentration`
   (Poisson-corrected copies per microliter of the final reaction
   [@hindson2011]), `Positives`, `Negatives`, `AcceptedDroplets`,
   `MeanAmplitudeOfPositives`, and `MeanAmplitudeOfNegatives`.
3. Name the file `YYYY-MM-DD_gene_project_plate.csv`
   (e.g., `2025-05-14_amoA_wetland_plate1.csv`).
4. Archive both the `.qlp` and the exported `.csv` to the lab shared drive
   under `ddPCR_raw_data/` before any analysis begins.

::: {.callout-warning}
### Data Integrity

- Never overwrite raw exported data files. Work only with copies.
- Back up raw `.qlp` files immediately after each run.
- All analysis scripts must be version-controlled in the lab GitHub
  repository.
:::

### 6.2 Quality Control and Acceptance Criteria

Apply these checks (ideally scripted in R, see Appendix A) before any analysis.
They follow the data-quality expectations set out for digital PCR reporting
[@dmiqe2020; @biorad_appsguide2018].

1. **Droplet count filter.** Exclude wells with fewer than 10,000 accepted
   droplets; below this, the Poisson estimate is unreliable
   [@biorad_qx200_2024].
2. **No-template controls (NTC).** NTC wells must show 0 or fewer than 3
   positive droplets. More than 3 positives indicates contamination; rerun the
   affected target or plate.
3. **Extraction blanks.** Blanks carried from DNA extraction (SOP-L700) must
   fall below the NTC threshold. Elevated counts point to contamination in the
   extraction workflow. Note that direct soil extraction is itself a known
   source of quantitative bias [@frostegard1999], so blanks and recovery checks
   matter for absolute numbers.
4. **Positive controls.** Positive-control gBlock wells should recover within
   20% of the expected concentration for the dilution prepared.
5. **Technical replicate agreement.** Where replicates were run, a coefficient
   of variation above 25% suggests pipetting error or sample heterogeneity.

::: {.callout-tip}
Build a QC report template in Quarto that automatically flags wells failing any
criterion, so batch QC is consistent and nothing is missed.
:::


## 7. Turning Counts into Ecological Quantities

### 7.1 Normalizing Copies per Microliter

The QuantaSoft `Concentration` is in copies per microliter of the ddPCR
reaction. To express results per gram of dry soil:

$$
\text{Copies per g dry soil} =
\frac{C_{\text{ddPCR}} \times V_{\text{reaction}} \times DF \times V_{\text{elution}}}{V_{\text{template}} \times m_{\text{soil}} \times (1 - \theta_g)}
$$

where $C_{\text{ddPCR}}$ is the QuantaSoft concentration (copies/uL),
$V_{\text{reaction}}$ the reaction volume (typically 20 uL), $DF$ the template
dilution factor, $V_{\text{elution}}$ the DNA elution volume (uL; see
SOP-L700), $V_{\text{template}}$ the template volume added to the reaction
(uL), $m_{\text{soil}}$ the mass of soil extracted (g), and $\theta_g$ the
gravimetric soil moisture as a decimal
(see [SOP-L400](SOP-L400-soil-moisture.qmd)).

To express results per nanogram of extracted DNA instead:

$$
\text{Copies per ng DNA} =
\frac{C_{\text{ddPCR}} \times V_{\text{reaction}}}{V_{\text{template}} \times C_{\text{DNA}}}
$$

where $C_{\text{DNA}}$ is the Qubit DNA concentration (ng/uL) from SOP-L700.

::: {.callout-note}
Report copies per gram dry soil when the question is about the size of a guild
in the soil (abundance per unit habitat). Report copies per nanogram DNA when
the question is about the relative composition of the community (proportion of
the recovered DNA). The two can diverge, and both are often reported.
:::

```{r}
#| label: copy-number-calc
#| echo: true

# Example calculation: copies per g dry soil
c_ddpcr <- 52.3          # copies/uL (from QuantaSoft)
v_reaction <- 20         # uL
dilution_factor <- 10    # 1:10 dilution of template
v_elution <- 100         # uL (from SOP-L700)
v_template <- 2          # uL added to ddPCR reaction
soil_mass_g <- 0.25      # g soil extracted
moisture_decimal <- 0.35 # gravimetric moisture as decimal

copies_per_g <- (c_ddpcr * v_reaction * dilution_factor * v_elution) /
                (v_template * soil_mass_g * (1 - moisture_decimal))

cat(sprintf("Gene copies per g dry soil: %.0f\n", copies_per_g))
cat(sprintf("Gene copies per g dry soil (log10): %.2f\n", log10(copies_per_g)))
```

### 7.2 Organizing the Analysis-Ready Dataset

1. Merge QC-filtered results with sample metadata (site, treatment, depth,
   date, soil moisture, DNA concentration) on a shared sample ID.
2. Keep the data tidy (long): one row per sample-by-gene combination
   [@wickham2014].
3. Save the merged dataset as `.csv` in the project GitHub repository under
   `data/processed/`.
4. *TBD: fix the standard column naming convention for the lab's ddPCR
   datasets (e.g., `sample_id`, `gene`, `copies_per_g_dry`,
   `copies_per_ng_dna`, `site`, `treatment`, `depth_cm`, `date`).*

### 7.3 Interpreting Gene Abundance for Nitrogen-Cycle Questions

Functional-gene abundance is genetic *potential*, not measured process rate;
interpret it as such, and relate it to environmental and rate data where
possible [@levybooth2014]. Useful framings include:

**Single-gene abundance.** Compare a guild's size across treatments, sites, or
depths (Section 8). For example, a rise in *amoA* (AOB) under fertilization is
consistent with greater bacterial ammonia-oxidizer potential.

**Gene ratios.** Ratios place one process relative to another and are often
more interpretable than single genes [@hallin2009]:

- *nirK*/*nirS*: relative contribution of the two nitrite-reductase pathways
  to denitrification.
- *nosZ*/(*nirK* + *nirS*): completeness of denitrification; higher values
  suggest greater capacity to reduce N~2~O to N~2~ rather than release it.
- *amoA* (AOA)/*amoA* (AOB): relative dominance of archaeal vs. bacterial
  ammonia oxidation, which tends to shift with pH and N availability
  [@francis2005].
- *nrfA*/(*nirK* + *nirS*): relative importance of nitrogen-retaining DNRA vs.
  nitrogen-losing denitrification.

*TBD: decide which ratios are central to the lab's current questions and set
standard figure formats for them.*

**Correlation with environmental variables.** Relate abundances to soil pH,
moisture, total C and N, nitrate, and ammonium using Pearson or Spearman
correlations. Where multivariate structure is of interest, consider PCA or
redundancy analysis once sample size supports it.

**Relating to process rates.** Where N~2~O flux, nitrification potential, or
similar rates are measured, gene abundance can be related to them, but the
analytical framework (and its limits) should be agreed with the PI first.


## 8. Statistical Analysis

::: {.callout-note}
The appropriate approach depends on design (paired vs. independent sites,
factorial treatments, time series). Consult the PI before finalizing.
:::

1. **Exploration.** Plot log~10~-transformed copy numbers by treatment, site,
   or depth (boxplots or dot plots). Assess normality of the log-transformed
   data with Q-Q plots and Shapiro-Wilk tests; investigate outliers as either
   biological signal or analytical artifact.
2. **Two-group comparisons** (e.g., amended vs. control): Welch's t-test on
   log~10~ copies per g dry soil, or the Wilcoxon rank-sum test if normality
   does not hold.
3. **Multi-group comparisons:** one-way ANOVA on log~10~ data with Tukey HSD,
   or Kruskal-Wallis with Dunn's test for non-parametric data.
4. **Factorial / hierarchical designs** (e.g., treatment by depth, sites as
   blocks): two-way ANOVA or linear mixed-effects models with site as a random
   effect [@pinheiro2000; @bolker2009].
5. **Values below detection.** Decide how to handle non-detects (e.g.,
   substitution, censored regression, or Tobit models) rather than dropping
   them silently [@helsel2011].


## 9. Reporting

- Report abundance as copies per g dry soil (mean +/- standard error) and/or
  copies per ng DNA.
- Plot on a log~10~ scale; report back-transformed means and confidence
  intervals in tables where appropriate.
- Include QC metrics (droplet counts, NTC results, positive-control recovery)
  as supplementary material, and report the experiment to digital-PCR
  reporting standards [@dmiqe2020].
- Cite the specific assays used (from SOP-L701) and the instrument
  configuration (from SOP-L214).
- *TBD: build a standardized Quarto reporting template for ddPCR results.*


## 10. References

::: {.callout-note}
### External resources
- **Assays and reagents:** validated PrimePCR ddPCR assays, supermixes, oils,
  and consumables are ordered from the Bio-Rad digital assay catalog,
  <https://www.bio-rad.com/digital-assays>.
- **Published methods:** Bio-Rad's ddPCR publication database,
  <https://www.bio-rad.com/ddPCR/publications>, is useful for finding
  peer-reviewed environmental and microbial applications when designing or
  validating an assay for a new gene target.
:::

::: {#refs}
:::


## 11. Training Record

Before analyzing data independently, personnel must:

1. Read this guide, [SOP-L214 QX200 Instrument Familiarization](SOP-L214-qx200-instrument-familiarization.qmd),
   [SOP-L701 Primer Design](SOP-L701-ddpcr-nitrogen-cycle-primers.qmd), and
   [SOP-L700 DNA Extraction](SOP-L700-dna-extraction.qmd).
2. Complete [SOP-L801 RStudio and GitHub](SOP-L801-rstudio-github.qmd) training.
3. Demonstrate competence exporting and QC-filtering a dataset.
4. Run the analysis pipeline on a practice dataset under supervision.
5. Be signed off by the PI or lab manager.


## Appendix A: Example QC Filtering Script {.appendix}

```{r}
#| label: qc-filter-example
#| echo: true
#| eval: false

library(tidyverse)

# Read exported QuantaSoft CSV
raw <- read_csv("data/raw/2025-05-14_amoA_wetland_plate1.csv")

# QC filtering
qc_filtered <- raw %>%
  filter(
    AcceptedDroplets >= 10000        # Minimum droplet threshold
  ) %>%
  mutate(
    qc_flag = case_when(
      Sample == "NTC" & Positives > 3 ~ "NTC_contamination",
      Sample == "Extraction_Blank" & Positives > 3 ~ "Extraction_contamination",
      TRUE ~ "Pass"
    )
  )

# Flag and report QC failures
qc_failures <- qc_filtered %>%
  filter(qc_flag != "Pass")

if (nrow(qc_failures) > 0) {
  warning("QC failures detected:")
  print(qc_failures %>% select(Well, Sample, Target, Positives, qc_flag))
}

# Retain only passing sample wells
analysis_ready <- qc_filtered %>%
  filter(qc_flag == "Pass",
         !Sample %in% c("NTC", "Extraction_Blank", "Positive_Control"))
```


## Appendix B: Visualization Template {.appendix}

```{r}
#| label: gene-abundance-plot
#| echo: true
#| eval: false

library(tidyverse)

# Assumes analysis_ready has columns: sample_id, gene, copies_per_g_dry, treatment
analysis_ready %>%
  ggplot(aes(x = treatment, y = log10(copies_per_g_dry), fill = treatment)) +
  geom_boxplot(alpha = 0.7, outlier.shape = NA) +
  geom_jitter(width = 0.2, alpha = 0.5, size = 1.5) +
  facet_wrap(~ gene, scales = "free_y") +
  labs(
    x = "Treatment",
    y = expression(log[10]~"(gene copies g"^{-1}~"dry soil)"),
    title = "Nitrogen Cycle Gene Abundance by Treatment"
  ) +
  theme_minimal() +
  theme(legend.position = "none")
```


## Appendix C: Sending Samples to External Laboratories {.appendix}

*TBD: if ddPCR is outsourced to an external service provider (university core
facility or commercial lab), document the preferred provider and contact,
sample submission requirements (volume, concentration, format), the data
format returned, turnaround time, cost per sample and funding account, and how
to reconcile external data with in-house QC standards.*
