---
# =============================================================================
# SOP-L807: Instrument Data Export and File Management
# =============================================================================
title: "Instrument Data Export and File Management"
sop-number: "SOP-L807"
author:
  - name: "Marc Los Huertos"
    affiliation: "EA Program, Pomona College"
date: last-modified
date-format: "YYYY-MM-DD"
revision: "0.1"

revision-history:
  - version: "0.1"
    date: "2026-05-21"
    author: "Marc Los Huertos"
    changes: "Initial release"
---

## 1. Purpose

This SOP standardizes the procedures for exporting raw data from
laboratory instruments, transferring files to the data management
system, and organizing instrument outputs for downstream analysis.
It ensures that raw instrument data are preserved without modification,
properly documented, and accessible for reproducible analysis
[@gandrud2015].


## 2. Scope

- **Applies to:** All analytical instruments in the EA Program
  laboratory that produce digital data files, including
  spectrophotometers, ICP-OES, TOC analyzers, ion chromatographs,
  balances with data output, pH/conductivity meters with logging, and
  field sensors that store data internally
- **Does not apply to:** Manual readings recorded directly in lab
  notebooks (see [SOP-L010](SOP-L010-lab-notebook-recording.qmd));
  drone imagery (see [SOP-L805](SOP-L805-drone-photogrammetry.qmd));
  bioinformatics sequencing data (see
  [SOP-L705](SOP-L705-bioinformatics-pipeline.qmd))
- **Location:** EA Program laboratory (SGM 133)


## 3. Responsibilities

| Role              | Responsibility                              |
|:------------------|:--------------------------------------------|
| Lab PI            | Approve file management protocols; ensure backup compliance |
| Lab Manager       | Maintain instrument computers; manage shared storage; train users on export procedures |
| Analyst / Student | Export data following this SOP; verify file integrity; commit to version control |
| Data Reviewer     | Confirm that raw files match lab notebook entries and are properly archived |


## 4. Safety and Hazards

Data export is a computational activity with no physical hazards.
When interacting with instruments (connecting USB drives, accessing
instrument PCs), follow the safety protocols in the relevant
instrument SOP.


## 5. Equipment and Materials

### Instruments and Export Formats

| Instrument                    | Native Format         | Preferred Export       | Notes                        |
|:------------------------------|:----------------------|:-----------------------|:-----------------------------|
| Shimadzu UV-Vis (UV-1900i)    | `.spc` / `.csv`       | CSV                    | Export from UVProbe software |
| Thermo iCAP ICP-OES           | `.rst` / `.csv`       | CSV                    | Export from Qtegra software  |
| Shimadzu TOC-L                | `.toa` / `.csv`       | CSV                    | Export from TOC-Control      |
| Dionex ICS Ion Chromatograph  | `.cdf` / `.csv`       | CSV                    | Export from Chromeleon       |
| Mettler Toledo balance        | Direct serial output  | CSV via terminal log   | Use logging script           |
| YSI ProDSS multiparameter    | `.csv` / `.dat`       | CSV                    | Download via KorDSS software |
| Hach DR3900 spectrophotometer | Internal memory       | CSV via USB            | Export from instrument menu  |
| Bio-Rad QX200 ddPCR           | `.qlp` / `.csv`       | CSV                    | Export from QuantaSoft       |
| PANalytical XRF               | `.csv` / `.xls`       | CSV                    | Export from SuperQ           |

### Transfer Media

| Item                    | Purpose               | Notes                        |
|:------------------------|:----------------------|:-----------------------------|
| USB flash drive (lab)   | Sneakernet transfer   | Dedicated drive; virus-scanned |
| Network shared folder   | Direct instrument PC transfer | If instrument PC has network access |
| RStudio Server upload   | Upload to analysis environment | Via Files pane or `scp` |


## 6. Procedure

### 6.1 General Export Workflow

The same sequence applies to all instruments:

1. **Complete the analytical run** and verify results on the
   instrument software
2. **Export raw data** from the instrument software to CSV format
   (or the most open/portable format available)
3. **Name the file** following the convention:
   `[instrument]_[method]_[YYYY-MM-DD]_[batch].[ext]`
   (e.g., `icpoes_metals_2026-05-21_batch01.csv`)
4. **Transfer the file** to the project's `data/raw/` directory
   using USB or network transfer
5. **Verify the transfer** by opening the file and checking row
   counts, column headers, and a few spot-checked values against
   the instrument display
6. **Record the export** in the lab notebook: file name, instrument,
   date, number of samples, and any export settings
7. **Commit to Git** with a descriptive message

::: {.callout-warning}
### Critical Rule
Never modify raw instrument files after export. If corrections are
needed, create a new file in `data/processed/` and document the
changes in your processing script. The raw file is your permanent
record.
:::

### 6.2 Instrument-Specific Export Instructions

#### Shimadzu UV-Vis (UV-1900i / UVProbe)

1. In UVProbe, open the data file (`.spc`)
2. File > Save As > CSV
3. Select "Wavelength and Absorbance" columns
4. Save to the USB drive as
   `uvvis_[method]_[YYYY-MM-DD].csv`

#### Thermo iCAP ICP-OES (Qtegra)

1. In Qtegra, navigate to the completed sequence
2. Right-click the sequence > Export Results
3. Select CSV format; include sample IDs, element concentrations,
   and QC flags
4. Save to the USB drive as
   `icpoes_[method]_[YYYY-MM-DD]_[batch].csv`

#### Shimadzu TOC-L (TOC-Control)

1. In TOC-Control, open the sample table for the completed run
2. File > Export > CSV
3. Include sample name, injection number, area, concentration,
   and CV
4. Save as `toc_[method]_[YYYY-MM-DD].csv`

#### YSI ProDSS Field Sensor

1. Connect the ProDSS to a computer via USB
2. Open KorDSS software
3. Select the data log for the sampling event
4. Export as CSV, including date/time, site ID, and all parameters
5. Save as `ysi_[project]_[YYYY-MM-DD].csv`

#### Bio-Rad QX200 ddPCR (QuantaSoft)

1. In QuantaSoft, open the completed experiment
2. Options > Export CSV (Well Data or Cluster Data)
3. Include well ID, sample name, target, concentration,
   positive/negative droplet counts
4. Save as `ddpcr_[target]_[YYYY-MM-DD].csv`
5. Also export the amplitude data for gating verification

### 6.3 File Transfer to RStudio Server

After exporting to USB or local machine:

```{bash}
#| echo: true
#| eval: false

# Option 1: scp from local machine to RStudio Server
scp icpoes_metals_2026-05-21_batch01.csv \
  username@rstudio.pomona.edu:~/project/data/raw/

# Option 2: Use the RStudio Files pane
# Click Upload > Browse > select file > OK
```

### 6.4 Automated Import Scripts in R

Create a standardized import script for each instrument type. This
ensures consistent parsing across analysts and batches:

```{r}
#| label: import-icpoes
#| echo: true
#| eval: false

#' Import ICP-OES data from Qtegra CSV export
#'
#' @param file_path Path to the exported CSV file
#' @return A tibble with standardized column names
import_icpoes <- function(file_path) {
  raw <- read_csv(file_path, skip = 2)  # Skip Qtegra header rows

  raw %>%
    rename_with(tolower) %>%
    rename_with(~ gsub(" ", "_", .)) %>%
    mutate(
      date_analyzed = as.Date(date_analyzed, format = "%m/%d/%Y"),
      across(where(is.character) & !c(sample_id, sample_type),
             as.numeric)
    ) %>%
    # Flag values below detection
    mutate(across(
      where(is.numeric),
      ~ if_else(. < 0, NA_real_, .)
    ))
}

# Usage
icp_data <- import_icpoes("data/raw/icpoes_metals_2026-05-21_batch01.csv")
glimpse(icp_data)
```

```{r}
#| label: import-toc
#| echo: true
#| eval: false

#' Import TOC data from TOC-Control CSV export
#'
#' @param file_path Path to the exported CSV file
#' @return A tibble with standardized column names
import_toc <- function(file_path) {
  raw <- read_csv(file_path)

  raw %>%
    rename_with(tolower) %>%
    rename_with(~ gsub("[^a-z0-9]", "_", .)) %>%
    select(sample_name, injection, area, concentration_mg_l = conc,
           cv_percent = cv) %>%
    filter(!is.na(sample_name))
}
```

### 6.5 Data Integrity Checks

After importing, run basic integrity checks:

```{r}
#| label: integrity-checks
#| echo: true
#| eval: false

# --- Row count matches instrument ---
cat(sprintf("Rows in file: %d\n", nrow(icp_data)))
# Compare with number of samples in lab notebook

# --- No fully blank rows ---
stopifnot(!any(rowSums(is.na(icp_data)) == ncol(icp_data)))

# --- Sample IDs match the run sequence ---
expected_ids <- c("Blank", "Std1", "Std2", "Std3",
                  paste0("S", sprintf("%03d", 1:20)),
                  "CRM", "Dup")
setdiff(expected_ids, icp_data$sample_id)
# Should be empty if all expected samples are present

# --- Calibration standards in expected range ---
icp_data %>%
  filter(grepl("^Std", sample_id)) %>%
  select(sample_id, starts_with("fe_")) %>%
  print()
```

### 6.6 Archival and Backup

All raw instrument files must be archived in three locations:

| Location                    | When                  | Responsible            |
|:----------------------------|:----------------------|:-----------------------|
| GitHub repository (`data/raw/`) | Immediately after export | Analyst            |
| Lab instrument PC           | Retained on instrument | Lab Manager            |
| External backup drive       | End of each semester  | Lab Manager            |

::: {.callout-tip}
### Tip
Large binary files (e.g., raw `.spc` or `.qlp` files that exceed
GitHub's file size recommendations) should be stored on the external
backup drive and the network share, not in Git. Only the exported CSV
goes into the repository.
:::

### 6.7 Instrument Log Cross-Reference

Every data export event must be recorded in the instrument log
(see [SOP-L007](SOP-L007-equipment-inventory.qmd)). The log entry
should include:

1. Date and time of the analytical run
2. Analyst name
3. File name(s) exported
4. Number of samples (including QC samples)
5. Any instrument warnings or anomalies noted during the run
6. Cross-reference to the lab notebook page number


## 7. Quality Control and Acceptance Criteria

- Raw files must be unmodified after export (verify with file hash
  if needed: `md5sum filename.csv`)
- File names must follow the naming convention in Section 6.1
- Row counts and sample IDs must match the run sequence recorded in
  the lab notebook
- Calibration data in the export must match the calibration curve
  approved during the run
- Refer to [SOP-L802](SOP-L802-quality-control.qmd) for analytical
  QC criteria (blanks, duplicates, recoveries)


## 8. Waste Disposal

Not applicable (computational procedure).


## 9. References

::: {#refs}
:::


## 10. Training Record

Training is documented in the EA Program training log. Before
exporting instrument data independently, personnel must:

1. Read this SOP in full
2. Complete training on the specific instrument
   (see the instrument-specific SOP)
3. Demonstrate a complete export, transfer, verification, and
   commit workflow under supervision
4. Be signed off by the PI or lab manager


## Appendix A: File Hash Verification

To confirm that a file has not been modified after export, compute
and record an MD5 hash:

```{bash}
#| echo: true
#| eval: false

# On Linux / Mac / RStudio Server terminal
md5sum data/raw/icpoes_metals_2026-05-21_batch01.csv
# Record this hash in the lab notebook alongside the file name
```

```{r}
#| label: hash-in-r
#| echo: true
#| eval: false

# In R
tools::md5sum("data/raw/icpoes_metals_2026-05-21_batch01.csv")
```

## Appendix B: Instrument Data Format Quick Reference

| Instrument       | Header Rows to Skip | Delimiter | Encoding | Date Format       |
|:-----------------|:--------------------|:----------|:---------|:------------------|
| Shimadzu UV-Vis  | 1                   | Comma     | UTF-8    | MM/DD/YYYY        |
| Thermo iCAP      | 2                   | Comma     | UTF-8    | MM/DD/YYYY HH:MM  |
| Shimadzu TOC-L   | 0                   | Comma     | Shift-JIS (convert to UTF-8) | YYYY/MM/DD |
| Dionex ICS        | 3                   | Comma     | UTF-8    | MM/DD/YYYY HH:MM  |
| YSI ProDSS       | 0                   | Comma     | UTF-8    | MM/DD/YYYY HH:MM  |
| Bio-Rad QX200    | 0                   | Comma     | UTF-8    | YYYY-MM-DD        |
| PANalytical XRF  | Variable            | Tab or comma | UTF-8 | DD/MM/YYYY        |
