---
# =============================================================================
# SOP-L08: Quality Control and Quality Assurance
# =============================================================================
title: "Quality Control and Quality Assurance"
sop-number: "SOP-L802"
author:
  - name: "Marc Los Huertos"
    affiliation: "EA Program, Pomona College"
date: last-modified
date-format: "YYYY-MM-DD"
revision: "1.0"

revision-history:
  - version: "1.0"
    date: "2025-05-16"
    author: "Marc Los Huertos"
    changes: "Converted from LaTeX to Quarto; comprehensive revision"
---

## 1. Purpose

This SOP establishes the overarching quality assurance (QA) and quality
control (QC) framework for all analytical work in the EA Program
laboratory. It defines the minimum QC requirements that apply across
methods, describes how to calculate and interpret QC metrics, specifies
detection limit procedures, and establishes the data documentation and
reproducibility practices required for every analysis. Individual
analytical SOPs reference this document for general QC principles and
add method-specific requirements [@usepa2007; @apha2012].

Quality assurance is the system of management actions that ensures
the analytical process produces results of known and documented quality.
Quality control is the set of technical activities (blanks, standards,
duplicates, spikes) embedded in each analytical batch to measure and
control that quality in real time [@taylor1987].


## 2. Scope

- **Applies to:** All quantitative chemical, physical, and biological
  analyses performed in the EA Program laboratory, including those
  described in SOPs L002 through L999 and analyses of field-collected
  samples (SOPs F001 through F999)
- **Does not apply to:** Qualitative screening tests (e.g., visual soil
  texture by feel, [SOP-F401](../field_sops/SOP-F401-soil-texture-by-feel.qmd))
  unless those tests include a quantitative component
- **Location:** EA Program laboratory, SGM 133


## 3. Responsibilities

| Role              | Responsibility                              |
|:------------------|:--------------------------------------------|
| Lab PI (QA Manager) | Approve this SOP; review QC data quarterly; authorize corrective actions |
| Lab Manager       | Maintain instrument logs, reagent inventory, and QC records; train new analysts |
| Analyst           | Implement all QC requirements for each batch; flag QC failures; document corrective actions |
| Data Reviewer     | Verify QC compliance before data are reported or published |


## 4. Safety and Hazards

QA/QC activities use the same reagents and equipment as the parent
analytical method. Refer to the method-specific SOP for safety
information. When preparing standards from stock solutions, observe
all PPE and fume hood requirements specified in
[SOP-L003 Hazardous Materials Handling](SOP-L003-hazardous-materials-handling.qmd).


## 5. Definitions

**Accuracy:** The closeness of a measured value to the true or accepted
reference value. Assessed using certified reference materials (CRMs),
laboratory fortified blanks (LFBs), and matrix spikes.

**Precision:** The closeness of replicated measurements to each other.
Assessed using duplicates, triplicates, and expressed as relative
percent difference (RPD), relative standard deviation (RSD), or
coefficient of variation (CV).

**Bias:** A systematic error that causes measured values to deviate
consistently from the true value in one direction.

**Analytical batch:** A group of samples processed together using the
same calibration, reagents, and instrument settings during a single
analytical session.

**Method Detection Limit (MDL):** The minimum concentration that can be
measured and reported with 99% confidence that the value is greater
than zero [@usepa_mdl_2016].

**Limit of Quantitation (LOQ):** The lowest concentration at which a
measurement meets predefined accuracy and precision criteria, typically
defined as 10 times the standard deviation of replicate blank
measurements (approximately 3.3 times the MDL) [@armbruster2008].

**Calibration verification standard (CVS):** A mid-range standard
analyzed during a run to check that the calibration remains valid.

**Laboratory Fortified Blank (LFB):** A blank matrix spiked with a
known concentration of analyte, processed through the entire method.
Assesses accuracy independent of matrix effects.

**Matrix Spike (MS):** A sample aliquot spiked with a known
concentration of analyte, processed through the entire method. Assesses
the effect of the sample matrix on analyte recovery.

**Certified Reference Material (CRM):** A material with certified
analyte concentrations traceable to a national or international standard
(e.g., NIST SRM).


## 6. Procedure

### 6.1 Calibration

#### 6.1.1 Initial Calibration

1. Prepare a minimum of five calibration standards spanning the expected
   concentration range of the samples, plus a calibration blank
   (zero-concentration standard prepared in the same matrix as the
   standards)
2. Run the standards from lowest to highest concentration
3. Fit a calibration model (linear, weighted linear, or quadratic as
   appropriate for the method)
4. Acceptance criteria:
   - Linear calibration: $r^2 \geq 0.995$
   - Each standard must back-calculate to within +/- 10% of its true
     value (or +/- 15% at the lowest standard)

5. If calibration fails, identify and correct the problem (contaminated
   standard, instrument drift, incorrect dilution) and recalibrate
   before proceeding

#### 6.1.2 Continuing Calibration Verification

1. Analyze a CVS (mid-range standard) at the start of each batch,
   after every 10 samples, and at the end of the batch
2. Acceptance: CVS recovery within 90--110% of the known value
3. If the CVS fails:
   - Re-analyze the CVS. If it fails again, recalibrate.
   - All samples analyzed after the last passing CVS must be re-analyzed.

#### 6.1.3 Calibration Documentation

Record the following for every calibration:

- Date, analyst, instrument ID
- Standard concentrations and source (lot number, preparation date)
- Calibration equation and $r^2$
- Back-calculated concentrations for each standard
- Pass/fail determination

### 6.2 Blanks

Blanks identify contamination introduced at each stage of the
analytical process:

| Blank Type           | Composition                       | Frequency              | Acceptance Criterion |
|:---------------------|:----------------------------------|:-----------------------|:---------------------|
| Calibration blank    | Same matrix as standards, no analyte | Every calibration   | Reading less than MDL |
| Method blank         | Reagent water processed through the entire method | 1 per batch (or 1 per 20 samples) | Less than MDL |
| Field blank          | Clean matrix (e.g., DI water) taken to the field and back | 1 per field event | Less than MDL |
| Equipment blank      | Reagent water rinsed through sampling equipment | 1 per equipment set per event | Less than MDL |
| Bottle blank         | DI water in a sample container, preserved identically to samples | 1 per 10 bottles per type ([SOP-F300](../field_sops/SOP-F300-surface-water-sampling.qmd)) | Less than MDL |

If a blank exceeds the MDL:

1. Identify the contamination source (reagent, glassware, environment)
2. Correct and re-analyze the blank
3. If blank contamination cannot be resolved, report the blank value
   alongside the sample results and flag the affected data

### 6.3 Duplicates and Precision

1. Analyze at least one duplicate per batch or 10% of samples,
   whichever is greater
2. Calculate the relative percent difference (RPD):

$$
\text{RPD} = \frac{|A - B|}{(A + B)/2} \times 100
$$

where $A$ and $B$ are the duplicate results.

3. Acceptance criteria (default; individual SOPs may specify tighter
   limits):

| Concentration Range         | Maximum RPD |
|:----------------------------|:------------|
| Greater than 10 times the MDL | 20%       |
| 5 to 10 times the MDL       | 30%        |
| Less than 5 times the MDL   | Not applicable; report both values |

4. If RPD exceeds the criterion, investigate (sample heterogeneity,
   instrument drift, operator error). Re-analyze if possible. Document
   the cause and corrective action.

### 6.4 Accuracy: Laboratory Fortified Blanks and Matrix Spikes

#### Laboratory Fortified Blank (LFB)

1. Prepare an LFB at a mid-range concentration in the method matrix
   (e.g., DI water, extractant solution)
2. Process through the entire analytical method alongside samples
3. Calculate percent recovery:

$$
\text{Recovery (\%)} = \frac{C_{\text{measured}}}{C_{\text{expected}}} \times 100
$$

4. Acceptance: 80--120% (default; method-specific SOPs may specify
   different windows)

#### Matrix Spike (MS)

1. Select a representative sample and divide into two aliquots
2. Spike one aliquot with a known concentration of the target analyte
   (typically at 1--5 times the expected sample concentration or at the
   mid-range calibration level)
3. Process both aliquots through the entire method
4. Calculate percent recovery:

$$
\text{Spike Recovery (\%)} = \frac{C_{\text{spiked}} - C_{\text{unspiked}}}{C_{\text{added}}} \times 100
$$

5. Acceptance: 75--125% (default)
6. Low or high recovery indicates matrix interference. Document and
   consider dilution, matrix-matched calibration, or standard addition
   as corrective measures.

#### Certified Reference Materials (CRMs)

1. Analyze a CRM with each batch when available for the target analyte
2. Recovery should fall within the certified confidence interval, or
   within 80--120% of the certified value if no interval is provided
3. Maintain a control chart (Section 6.8) to track CRM performance
   over time

### 6.5 Detection Limits

#### 6.5.1 Method Detection Limit (MDL)

The MDL is determined per the EPA procedure [@usepa_mdl_2016]:

1. Prepare at least seven replicate samples spiked at 1--5 times the
   estimated detection limit in the method matrix
2. Process all replicates through the entire method
3. Calculate the standard deviation ($s$) of the replicate
   measurements
4. Compute:

$$
\text{MDL} = t_{(n-1, \alpha=0.01)} \times s
$$

where $t$ is the one-sided Student's $t$-value at the 99% confidence
level for $n - 1$ degrees of freedom (for $n = 7$, $t = 3.143$).

5. Verify the MDL by confirming that the spike concentration is within
   1--10 times the calculated MDL. If not, adjust the spike level and
   repeat.
6. Re-determine the MDL annually, whenever there is a significant
   change in method or instrument, or when ongoing blank data suggest
   the MDL has changed.

#### 6.5.2 Limit of Quantitation (LOQ)

$$
\text{LOQ} = 10 \times s
$$

where $s$ is the same standard deviation used for the MDL calculation.
The LOQ is the lowest concentration that can be reliably quantified
with acceptable precision and accuracy.

#### 6.5.3 Reporting Values Below the MDL

- Results below the MDL are reported as "less than MDL" (e.g.,
  "< 0.05 mg/L")
- Results between the MDL and LOQ are reported with a qualifier flag
  (e.g., "J" flag) indicating estimated value
- Results above the LOQ are reported without qualification
- For statistical analyses of datasets containing censored values
  (below MDL), use appropriate methods such as Kaplan-Meier estimation,
  regression on order statistics (ROS), or maximum likelihood estimation
  rather than simple substitution [@helsel2012]

### 6.6 Holding Times and Sample Integrity

Samples have finite stability. Analyses must be completed within the
method-specified holding time. Record the collection date/time and the
analysis date/time for every sample. If a holding time is exceeded,
flag the result with an "H" qualifier.

Common holding times for EA Program analyses:

| Analysis                  | Holding Time              | Preservation              | Reference SOP |
|:--------------------------|:--------------------------|:--------------------------|:--------------|
| pH                        | 15 minutes (field); 24 h (lab) | None; analyze ASAP   | [L300](SOP-L300-pH-conductivity-bench.qmd), [L402](SOP-L402-soil-pH-salinity.qmd) |
| Conductivity              | 28 days                   | Refrigerate 4 C           | [L300](SOP-L300-pH-conductivity-bench.qmd) |
| Nitrate/nitrite           | 48 hours                  | Refrigerate 4 C; H~2~SO~4~ to pH < 2 | [L403](SOP-L403-nitrate-ammonium-kcl.qmd) |
| Ammonium                  | 48 hours                  | Refrigerate 4 C; H~2~SO~4~ to pH < 2 | [L403](SOP-L403-nitrate-ammonium-kcl.qmd) |
| Total phosphorus          | 28 days                   | H~2~SO~4~ to pH < 2; 4 C | [L405](SOP-L405-extractable-phosphorus.qmd) |
| Chlorophyll *a* (filter)  | 28 days frozen            | Freeze at --20 C in dark  | [L310](SOP-L310-chlorophyll-extraction.qmd) |
| Metals (acid-digested)    | 6 months                  | HNO~3~ to pH < 2; 4 C    | [L410](SOP-L410-hotplate-acid-digestion.qmd), [L411](SOP-L411-mars-microwave-digestion.qmd) |
| Coliforms (Colilert)      | 6 hours to incubation     | Refrigerate 4 C           | [L305](SOP-L305-colilert-quanti-tray.qmd) |
| Soil KCl extracts         | 24 hours                  | Refrigerate 4 C           | [L403](SOP-L403-nitrate-ammonium-kcl.qmd) |
| Soil moisture             | Analyze immediately or freeze | Freeze if delayed > 24 h | [L400](SOP-L400-soil-moisture.qmd) |

### 6.7 Corrective Actions

When a QC criterion is not met:

1. Stop analysis. Do not proceed with samples until the problem is
   resolved.
2. Investigate the root cause (contamination, calibration drift,
   reagent degradation, instrument malfunction, operator error)
3. Implement a corrective action (re-prepare reagents, recalibrate,
   clean equipment, re-train analyst)
4. Re-analyze the affected QC sample. If it passes, resume analysis.
5. Re-analyze all samples analyzed after the last passing QC check.
6. Document the failure, root cause, corrective action, and re-analysis
   results in the QC log and lab notebook.

### 6.8 Control Charts

Maintain control charts for each analyte/method combination to track
long-term analytical performance:

1. Plot CRM, LFB, or CVS recovery for each batch on a time-series
   chart
2. Calculate and draw lines for:
   - **Center line:** Mean recovery over the last 20+ batches
   - **Warning limits:** +/- 2 standard deviations from the mean
   - **Action limits:** +/- 3 standard deviations from the mean
3. A result between warning and action limits is a warning; investigate
   but do not necessarily re-analyze
4. A result beyond the action limits requires corrective action and
   re-analysis
5. Systematic trends (7+ consecutive results on one side of the center
   line, or a steady upward/downward drift) require investigation even
   if individual points are within limits


## 7. Metadata Documentation and Reproducibility

### 7.1 Laboratory Notebook Requirements

Every analytical session must be documented in the lab notebook (paper
or electronic) with the following minimum information:

- Date, analyst name
- SOP number and revision used
- Instrument ID and calibration status
- Reagent lot numbers and preparation dates
- Sample IDs, matrix type, and collection dates
- Extraction/digestion batch identifiers
- Calibration data (standards, equation, $r^2$)
- QC results (blanks, duplicates, LFBs, spikes, CRMs) with pass/fail
- Raw data file names and storage location
- Any deviations from the SOP, with justification
- Corrective actions taken for QC failures

### 7.2 Electronic Data Management

1. Store all raw instrument data files on the lab shared drive in a
   folder structure organized by project, year, and instrument:

```
lab_data/
  project_name/
    YYYY/
      instrument_name/
        raw/          # Unmodified instrument output files
        processed/    # R-generated output (cleaned data, QC reports)
        scripts/      # R scripts used for processing
```

2. Never modify raw data files. All processing must be done
   programmatically in R with version-controlled scripts.
3. Commit all analysis scripts to the lab GitHub repository per
   [SOP-L801](SOP-L801-rstudio-github.qmd)

### 7.3 R-Based QC Tracking System

The following R package and script framework is the EA Program standard
for QC documentation and reproducibility. All analysts must use this
workflow or an equivalent that produces the same outputs.

```{r}
#| label: qc-tracking-setup
#| echo: true
#| eval: false

# =============================================================================
# EA Program QA/QC Tracking Framework
# Source this script at the start of every analytical R session
# =============================================================================

library(tidyverse)
library(lubridate)
library(knitr)

# --- Session metadata (fill in for each batch) ---
session_meta <- list(
  analyst = "Your Name",
  date = Sys.Date(),
  sop = "SOP-L06 v1.0",
  method_sop = "SOP-LXX vX.X",    # The specific analytical SOP

  instrument = "Instrument ID",
  project = "Project Name",
  notes = ""
)

cat("QC session initialized:\n")
cat(sprintf("  Analyst:    %s\n", session_meta$analyst))
cat(sprintf("  Date:       %s\n", session_meta$date))
cat(sprintf("  Method SOP: %s\n", session_meta$method_sop))
cat(sprintf("  Instrument: %s\n", session_meta$instrument))
```

```{r}
#| label: qc-functions
#| echo: true
#| eval: false

# =============================================================================
# QC calculation functions
# =============================================================================

#' Calculate Method Detection Limit (EPA procedure)
#' @param reps Numeric vector of replicate measurements at the spike level
#' @param alpha Significance level (default 0.01 for 99% confidence)
#' @return Named list with MDL, LOQ, n, sd, t-value
calc_mdl <- function(reps, alpha = 0.01) {
  n <- length(reps)
  s <- sd(reps)
  t_val <- qt(1 - alpha, df = n - 1)
  mdl <- t_val * s
  loq <- 10 * s

  list(
    n = n,
    mean = mean(reps),
    sd = s,
    t_value = round(t_val, 3),
    MDL = round(mdl, 4),
    LOQ = round(loq, 4),
    spike_to_mdl_ratio = round(mean(reps) / mdl, 1)
  )
}

#' Calculate Relative Percent Difference for duplicates
#' @param a First measurement
#' @param b Second measurement
#' @return RPD as a percentage
calc_rpd <- function(a, b) {
  abs(a - b) / ((a + b) / 2) * 100
}

#' Calculate percent recovery
#' @param measured Measured concentration
#' @param expected Expected (true) concentration
#' @return Recovery as a percentage
calc_recovery <- function(measured, expected) {
  (measured / expected) * 100
}

#' Calculate spike recovery
#' @param spiked Concentration of spiked sample
#' @param unspiked Concentration of unspiked sample
#' @param added Concentration of spike added
#' @return Spike recovery as a percentage
calc_spike_recovery <- function(spiked, unspiked, added) {
  ((spiked - unspiked) / added) * 100
}

#' Evaluate QC pass/fail against criteria
#' @param qc_data Tibble with columns: qc_type, value
#' @param criteria Named list of acceptance criteria
#' @return Tibble with pass/fail flags
evaluate_qc <- function(qc_data, criteria) {
  qc_data %>%
    mutate(
      criterion = map_chr(qc_type, ~ criteria[[.x]] %||% "Not specified"),
      pass = case_when(
        qc_type == "blank" ~ value < criteria$mdl,
        qc_type == "rpd" ~ value <= criteria$max_rpd,
        qc_type == "recovery" ~ value >= criteria$min_recovery &
                                 value <= criteria$max_recovery,
        qc_type == "spike" ~ value >= criteria$min_spike &
                              value <= criteria$max_spike,
        TRUE ~ NA
      ),
      status = if_else(pass, "PASS", "FAIL")
    )
}
```

```{r}
#| label: qc-batch-report
#| echo: true
#| eval: false

# =============================================================================
# Example: Complete QC batch report
# =============================================================================

# --- 1. Read raw data ---
raw <- read_csv("data/raw/2025-05-16_phosphorus_batch1.csv")

# --- 2. Define QC acceptance criteria for this method ---
criteria <- list(
  mdl = 0.02,             # mg/L (from MDL study)
  loq = 0.07,             # mg/L
  max_rpd = 20,           # percent
  min_recovery = 80,      # percent (LFB)
  max_recovery = 120,     # percent (LFB)
  min_spike = 75,         # percent (MS)
  max_spike = 125,        # percent (MS)
  r_squared_min = 0.995,  # calibration
  cvs_min = 90,           # percent
  cvs_max = 110           # percent
)

# --- 3. Check calibration ---
cal_data <- raw %>% filter(sample_type == "standard")
cal_model <- lm(response ~ concentration, data = cal_data)
r_sq <- summary(cal_model)$r.squared

cal_check <- tibble(
  check = "Calibration r-squared",
  value = round(r_sq, 5),
  criterion = sprintf(">= %.3f", criteria$r_squared_min),
  status = if_else(r_sq >= criteria$r_squared_min, "PASS", "FAIL")
)

# --- 4. Check blanks ---
blanks <- raw %>%
  filter(sample_type == "method_blank") %>%
  summarize(max_blank = max(concentration, na.rm = TRUE))

blank_check <- tibble(
  check = "Method blank",
  value = blanks$max_blank,
  criterion = sprintf("< %.3f mg/L (MDL)", criteria$mdl),
  status = if_else(blanks$max_blank < criteria$mdl, "PASS", "FAIL")
)

# --- 5. Check duplicates ---
dups <- raw %>%
  filter(sample_type == "duplicate") %>%
  group_by(sample_id) %>%
  summarize(
    a = first(concentration),
    b = last(concentration),
    rpd = calc_rpd(a, b),
    .groups = "drop"
  )

dup_check <- dups %>%
  summarize(
    check = "Duplicate RPD (max)",
    value = round(max(rpd), 1),
    criterion = sprintf("<= %d%%", criteria$max_rpd),
    status = if_else(max(rpd) <= criteria$max_rpd, "PASS", "FAIL")
  )

# --- 6. Check LFB recovery ---
lfb <- raw %>%
  filter(sample_type == "LFB") %>%
  mutate(recovery = calc_recovery(concentration, expected_concentration))

lfb_check <- tibble(
  check = "LFB recovery",
  value = round(mean(lfb$recovery), 1),
  criterion = sprintf("%d--%d%%", criteria$min_recovery, criteria$max_recovery),
  status = if_else(
    all(lfb$recovery >= criteria$min_recovery &
        lfb$recovery <= criteria$max_recovery),
    "PASS", "FAIL"
  )
)

# --- 7. Check CVS ---
cvs <- raw %>%
  filter(sample_type == "CVS") %>%
  mutate(recovery = calc_recovery(concentration, expected_concentration))

cvs_check <- tibble(
  check = "CVS recovery (all)",
  value = sprintf("%.1f--%.1f%%", min(cvs$recovery), max(cvs$recovery)),
  criterion = sprintf("%d--%d%%", criteria$cvs_min, criteria$cvs_max),
  status = if_else(
    all(cvs$recovery >= criteria$cvs_min &
        cvs$recovery <= criteria$cvs_max),
    "PASS", "FAIL"
  )
)

# --- 8. Compile and print QC summary ---
qc_summary <- bind_rows(cal_check, blank_check, dup_check,
                         lfb_check, cvs_check)

cat("\n========== QC BATCH REPORT ==========\n")
cat(sprintf("Date:       %s\n", session_meta$date))
cat(sprintf("Analyst:    %s\n", session_meta$analyst))
cat(sprintf("Method:     %s\n", session_meta$method_sop))
cat(sprintf("Instrument: %s\n", session_meta$instrument))
cat("=====================================\n\n")

print(kable(qc_summary, format = "pipe"))

# Flag overall batch status
if (any(qc_summary$status == "FAIL")) {
  warning("\n*** QC FAILURE DETECTED -- review and take corrective action ***\n")
} else {
  cat("\nAll QC checks PASSED.\n")
}

# --- 9. Save QC report ---
write_csv(qc_summary,
          sprintf("data/processed/qc_report_%s_%s.csv",
                  session_meta$method_sop,
                  session_meta$date))
```

```{r}
#| label: mdl-study
#| echo: true
#| eval: false

# =============================================================================
# MDL Determination Study
# Run once per method, or when method/instrument changes significantly
# =============================================================================

# Seven replicate measurements at the spike level (1-5x estimated DL)
mdl_reps <- c(0.032, 0.028, 0.035, 0.031, 0.029, 0.033, 0.030)

mdl_result <- calc_mdl(mdl_reps)

cat("MDL Study Results:\n")
cat(sprintf("  n replicates:        %d\n", mdl_result$n))
cat(sprintf("  Mean:                %.4f mg/L\n", mdl_result$mean))
cat(sprintf("  Std dev:             %.4f mg/L\n", mdl_result$sd))
cat(sprintf("  t-value (99%%):       %.3f\n", mdl_result$t_value))
cat(sprintf("  MDL:                 %.4f mg/L\n", mdl_result$MDL))
cat(sprintf("  LOQ:                 %.4f mg/L\n", mdl_result$LOQ))
cat(sprintf("  Spike/MDL ratio:     %.1f (should be 1-10)\n",
            mdl_result$spike_to_mdl_ratio))
```

```{r}
#| label: control-chart
#| echo: true
#| eval: false

# =============================================================================
# Control chart for long-term QC tracking
# =============================================================================

library(tidyverse)

# Read historical QC data (one row per batch)
qc_history <- read_csv("data/processed/qc_history_phosphorus.csv") %>%
  mutate(date = as_date(date))

# Calculate control limits
center <- mean(qc_history$crm_recovery)
sd_qc <- sd(qc_history$crm_recovery)
warning_upper <- center + 2 * sd_qc
warning_lower <- center - 2 * sd_qc
action_upper <- center + 3 * sd_qc
action_lower <- center - 3 * sd_qc

ggplot(qc_history, aes(x = date, y = crm_recovery)) +
  geom_point(size = 2) +
  geom_line(alpha = 0.3) +
  geom_hline(yintercept = center, color = "darkgreen", linewidth = 1) +
  geom_hline(yintercept = c(warning_upper, warning_lower),
             color = "orange", linetype = "dashed") +
  geom_hline(yintercept = c(action_upper, action_lower),
             color = "red", linetype = "dotted") +
  annotate("text", x = min(qc_history$date), y = center + 1,
           label = "Center", color = "darkgreen", hjust = 0) +
  annotate("text", x = min(qc_history$date), y = warning_upper + 1,
           label = "+/- 2 SD (Warning)", color = "orange", hjust = 0) +
  annotate("text", x = min(qc_history$date), y = action_upper + 1,
           label = "+/- 3 SD (Action)", color = "red", hjust = 0) +
  labs(
    x = "Date",
    y = "CRM Recovery (%)",
    title = "Phosphorus CRM Control Chart"
  ) +
  theme_minimal()
```

### 7.4 Quarto Report Templates

For each analytical method, create a Quarto (`.qmd`) report template
that:

1. Reads the raw data file
2. Runs all QC checks programmatically using the functions above
3. Produces a formatted QC summary table
4. Generates calibration plots, control charts, and diagnostic figures
5. Exports a clean, QC-verified dataset as a CSV
6. Renders to HTML or PDF for archiving

Store templates in the project GitHub repository under `analysis/templates/`. An analyst runs the template with their batch-specific data file as input, producing a self-contained, reproducible QC report for every batch [@gandrud2015; @xie2015].


## 8. QC Requirements by Method Category

The table below summarizes the minimum QC sample frequencies for
each method category. Individual SOPs may specify additional
requirements.

| QC Element         | Water Chemistry | Soil Extraction | Digestion (Metals) | Biological (Colilert) | Instrumental (TOC, ICP) |
|:-------------------|:---------------|:---------------|:-------------------|:---------------------|:----------------------|
| Calibration blank  | Each cal       | Each cal       | Each cal           | N/A                  | Each cal              |
| Method blank       | 1/batch        | 1/batch        | 1/batch            | 1/batch              | 1/batch               |
| Field blank        | 1/event        | N/A            | N/A                | N/A                  | N/A                   |
| Duplicate          | 10%            | 10%            | 10%                | 10%                  | 10%                   |
| LFB                | 1/batch        | 1/batch        | 1/batch            | N/A                  | 1/batch               |
| Matrix spike       | 5% (or 1/batch)| 5%             | 5%                 | N/A                  | 5%                    |
| CRM                | 1/batch (if avail.) | 1/batch   | 1/batch            | Positive control     | 1/batch               |
| CVS                | Every 10 samples | Every 10     | Every 10           | N/A                  | Every 10              |


## 9. Waste Disposal

QC samples generate the same waste streams as routine samples. Dispose
of QC waste (blanks, standards, spikes) according to the parent
method's waste disposal procedure.


## 10. References

- APHA, AWWA, WEF. (2012). *Standard Methods for Examination of Water
  and Wastewater*. 22nd ed. American Public Health Association (Eds.).
  Washington. 1360 pp.
- Armbruster, D.A. & Pry, T. (2008). Limit of blank, limit of
  detection and limit of quantitation. *Clinical Biochemistry Reviews*,
  29(Suppl 1), S49--S52.
- Gandrud, C. (2015). *Reproducible Research with R and RStudio*. 2nd
  ed. CRC Press, Boca Raton, FL.
- Helsel, D.R. (2012). *Statistics for Censored Environmental Data
  Using Minitab and R*. 2nd ed. John Wiley & Sons, Hoboken, NJ.
- Taylor, J.K. (1987). *Quality Assurance of Chemical Measurements*.
  Lewis Publishers, Chelsea, MI.
- U.S. Environmental Protection Agency. (2007). *Guidance for Quality
  Assurance Project Plans*. EPA/240/R-02/009. Office of Environmental
  Information, Washington, DC.
- U.S. Environmental Protection Agency. (2016). *Definition and
  procedure for the determination of the Method Detection Limit*.
  Revision 2. 40 CFR Part 136, Appendix B.
- Xie, Y. (2015). *Dynamic Documents with R and knitr*. 2nd ed. CRC
  Press, Boca Raton, FL.

::: {#refs}
:::


## 11. Training Record

Before performing analytical work independently, personnel must:

1. Read this SOP in full
2. Complete [SOP-L002 Laboratory Safety](SOP-L002-laboratory-safety.qmd)
   and [SOP-L801 RStudio and GitHub](SOP-L801-rstudio-github.qmd)
   training
3. Complete training on the specific analytical SOP(s) they will use
4. Demonstrate ability to prepare calibration standards, blanks, and
   QC samples
5. Run at least one complete batch with all required QC elements under
   supervision
6. Produce a QC batch report using the R framework (Section 7.3) with
   all checks passing
7. Be signed off by the PI or lab manager


## Appendix A: QC Decision Flowchart {.appendix}

For each QC check that fails during an analytical batch:

1. **Stop.** Do not analyze additional samples.
2. **Identify.** Which QC element failed? (blank, CVS, duplicate, LFB,
   spike, CRM)
3. **Investigate.** Check reagents, standards, instrument, and
   technique.
4. **Correct.** Fix the identified problem.
5. **Re-analyze.** Re-run the failed QC sample.
6. **Pass?** If yes, re-analyze all samples after the last passing QC
   and resume. If no, escalate to the lab manager.
7. **Document.** Record the failure, investigation, corrective action,
   and resolution in the QC log and lab notebook.


## Appendix B: Data Qualifier Flags {.appendix}

| Flag | Meaning                                              |
|:-----|:-----------------------------------------------------|
| U    | Analyte not detected; result reported at the MDL     |
| J    | Estimated value; detected between MDL and LOQ        |
| B    | Analyte detected in the method blank                 |
| H    | Holding time exceeded                                |
| R    | Result rejected due to QC failure                    |
| E    | Estimated value; result exceeds the calibration range |
| D    | Diluted sample; reported value corrected for dilution |
| MS   | Matrix spike recovery outside acceptance limits      |



