---
# =============================================================================
# SOP-L805: Drone Image Processing (Photogrammetry)
# =============================================================================
title: "Drone Image Processing (Photogrammetry)"
sop-number: "SOP-L805"
author:
  - name: "Marc Los Huertos"
    affiliation: "EA Program, Pomona College"
date: last-modified
date-format: "YYYY-MM-DD"
revision: "0.1"

revision-history:
  - version: "0.1"
    date: "2026-05-21"
    author: "Marc Los Huertos"
    changes: "Initial release"
---

## Purpose

This SOP describes the post-processing workflow for aerial imagery
collected by uncrewed aerial systems (UAS/drones). It covers image
organization, photogrammetric reconstruction (orthomosaics, digital
surface models, and point clouds), ground control point integration,
accuracy assessment, and data archiving [@anderson2019; @iglhaut2019].


## Scope

- **Applies to:** Photogrammetric processing of RGB, multispectral,
  and thermal drone imagery collected for EA Program research
- **Does not apply to:** Drone flight planning, airspace authorization,
  or FAA Part 107 compliance (covered in the drone operations field
  SOP); GIS analysis of processed outputs (see
  [SOP-L804](SOP-L804-gis-spatial-data.qmd))
- **Location:** EA Program laboratory (SGM 133) and processing
  workstations


## Responsibilities

| Role              | Responsibility                              |
|:------------------|:--------------------------------------------|
| Lab PI            | Approve processing workflows; ensure data storage compliance |
| Drone Pilot / Operator | Deliver organized imagery with flight logs; place and survey GCPs |
| Processing Analyst | Run photogrammetric software; assess quality; archive outputs |
| Data Reviewer     | Verify accuracy metrics before outputs are used in analysis |


## Safety and Hazards

::: {.callout-warning}
### Hazards

Photogrammetric processing is computationally intensive but poses no
physical hazards. However, extended processing sessions may generate
significant heat from workstation components. Ensure adequate
ventilation around processing workstations.
:::

For field-related hazards during drone operations and GCP placement,
refer to [SOP-F001 Field Safety](../field_sops/SOP-F001-field-safety.qmd).


## Equipment and Materials

### Software

| Item                        | Access                | Notes                        |
|:----------------------------|:----------------------|:-----------------------------|
| Agisoft Metashape Professional | Campus license     | Primary photogrammetry tool  |
| OpenDroneMap (ODM)          | Free / open source    | Alternative processing       |
| WebODM                      | Self-hosted           | Browser interface for ODM    |
| QGIS (3.x)                 | Free download         | Visualization and QC         |
| R with `terra` / `sf`      | RStudio Server        | Post-processing analysis     |
| CloudCompare                | Free download         | Point cloud inspection       |

### Hardware

| Item                        | Minimum Specs         | Notes                        |
|:----------------------------|:----------------------|:-----------------------------|
| Processing workstation      | 32 GB RAM, dedicated GPU, SSD | More RAM for large datasets |
| External storage            | 2 TB+ portable SSD   | Raw image archive            |

### Drone Platforms

| Platform                    | Sensor                | Resolution               |
|:----------------------------|:----------------------|:-------------------------|
| DJI Mavic 3 Enterprise      | RGB (Hasselblad)     | 20 MP; 0.7 cm/px at 100 m |
| DJI Matrice 350 RTK         | Zenmuse P1 / L2      | High-res RGB or LiDAR   |
| Custom multispectral rig    | MicaSense RedEdge-P  | 5-band multispectral     |


## Procedure

### Image Organization

Before processing, organize raw imagery following a consistent
directory structure:

```
project_name/
├── drone/
│   ├── raw_images/
│   │   ├── flight_2026-05-21_rgb/
│   │   │   ├── DJI_0001.JPG
│   │   │   ├── DJI_0002.JPG
│   │   │   └── ...
│   │   └── flight_2026-05-21_multi/
│   ├── flight_logs/
│   │   └── flight_2026-05-21.csv
│   ├── gcp/
│   │   └── gcp_coordinates.csv
│   ├── processed/
│   │   ├── orthomosaic/
│   │   ├── dsm/
│   │   ├── point_cloud/
│   │   └── report/
│   └── README.md
```

Check that all images have valid EXIF data (GPS coordinates, camera
parameters). Remove obviously corrupted or blurry images before
processing.

```{r}
#| label: exif-check
#| echo: true
#| eval: false

library(exifr)

# Read EXIF from all images in a flight folder
exif_data <- read_exif(
  list.files("drone/raw_images/flight_2026-05-21_rgb/",
             pattern = "\\.JPG$", full.names = TRUE),
  tags = c("FileName", "GPSLatitude", "GPSLongitude",
           "GPSAltitude", "FocalLength", "ImageWidth",
           "ImageHeight", "DateTimeOriginal")
)

# Check for missing GPS
missing_gps <- exif_data %>%
  filter(is.na(GPSLatitude) | is.na(GPSLongitude))
cat(sprintf("%d images missing GPS coordinates\n", nrow(missing_gps)))

# Summary
cat(sprintf("Total images: %d\n", nrow(exif_data)))
cat(sprintf("Date range: %s to %s\n",
            min(exif_data$DateTimeOriginal),
            max(exif_data$DateTimeOriginal)))
```

### Ground Control Points (GCPs)

GCPs are surveyed reference points used to improve the geometric
accuracy of photogrammetric outputs.

1. Place GCPs before the flight. Use high-contrast targets (black
   and white checkerboard, minimum 30 cm) visible from flight altitude
2. Survey GCP positions using a differential GPS or RTK receiver.
   Record coordinates in WGS 84 (EPSG:4326) and a local projected
   CRS (e.g., EPSG:26911)
3. Format GCP data as a CSV file:

```
gcp_id,longitude,latitude,elevation_m,crs
GCP01,-117.7123,34.0966,398.2,EPSG:4326
GCP02,-117.7118,34.0971,401.5,EPSG:4326
```

4. Use a minimum of 5 GCPs distributed across the study area.
   Reserve 2 or more as independent check points for accuracy
   assessment

### Photogrammetric Processing (Agisoft Metashape)

The following workflow applies to Agisoft Metashape Professional.
Adapt parameters for your specific project needs.

**Step 1: Create project and import photos**

1. Create a new Metashape project and save to `drone/processed/`
2. Add photos from the flight folder
3. Verify camera positions on the reference pane

**Step 2: Align photos**

1. Workflow > Align Photos
2. Settings: Accuracy = High; Key point limit = 40,000;
   Tie point limit = 10,000; Adaptive camera model fitting = enabled
3. Review the sparse point cloud and camera positions

**Step 3: Import and assign GCPs**

1. Import GCP coordinates from the CSV file
2. Assign each GCP to its corresponding target in at least 3 photos
3. Uncheck GCPs designated as check points
4. Run Optimize Cameras

**Step 4: Build dense point cloud**

1. Workflow > Build Point Cloud
2. Quality = High (or Medium for initial assessment);
   Depth filtering = Moderate
3. Inspect the dense cloud for artifacts; remove obvious noise
   manually or with CloudCompare

**Step 5: Build DEM and orthomosaic**

1. Workflow > Build DEM (Source: Point Cloud)
2. Workflow > Build Orthomosaic (Surface: DEM;
   Blending mode: Mosaic)

**Step 6: Export products**

1. Export DEM as GeoTIFF to `drone/processed/dsm/`
2. Export orthomosaic as GeoTIFF to `drone/processed/orthomosaic/`
3. Export point cloud as `.laz` to `drone/processed/point_cloud/`
4. Export processing report as PDF to `drone/processed/report/`

::: {.callout-tip}
### Tip
For large datasets (more than 1,000 images), process on the
lab workstation rather than a laptop. Expect processing times of
several hours for dense cloud generation at High quality.
:::

### OpenDroneMap (Alternative Workflow)

For projects using the open-source pipeline:

```{bash}
#| echo: true
#| eval: false

# Run ODM with standard settings
docker run -ti --rm \
  -v /path/to/project/drone:/datasets \
  opendronemap/odm \
  --project-path /datasets \
  --dsm \
  --orthophoto-resolution 2.0 \
  --dem-resolution 5.0 \
  --gcp /datasets/gcp/gcp_list.txt
```

### Accuracy Assessment

Evaluate positional accuracy using the reserved check points:

```{r}
#| label: accuracy-assessment
#| echo: true
#| eval: false

library(terra)
library(sf)

# Load the DSM
dsm <- rast("drone/processed/dsm/dsm_2026-05-21.tif")

# Load check points
check_pts <- read_csv("drone/gcp/check_points.csv") %>%
  st_as_sf(coords = c("longitude", "latitude"), crs = 4326) %>%
  st_transform(crs = crs(dsm))

# Extract DSM elevation at check points
check_pts$dsm_elev <- extract(dsm, check_pts)$value

# Calculate vertical RMSE
check_pts <- check_pts %>%
  mutate(error_m = dsm_elev - elevation_m)

rmse_z <- sqrt(mean(check_pts$error_m^2, na.rm = TRUE))
mae_z <- mean(abs(check_pts$error_m), na.rm = TRUE)

cat(sprintf("Vertical RMSE: %.3f m\n", rmse_z))
cat(sprintf("Vertical MAE:  %.3f m\n", mae_z))
cat(sprintf("n check points: %d\n", nrow(check_pts)))
```

Report accuracy metrics in all publications and project reports.
Typical targets for environmental applications: horizontal RMSE
less than 2 times the ground sampling distance (GSD); vertical RMSE
less than 3 times the GSD [@james2017].

### Multispectral Processing

For multispectral imagery (e.g., MicaSense RedEdge):

1. Include reflectance calibration panel images (taken before and
   after each flight)
2. Process each band alignment in Metashape or use MicaSense's
   Atlas or `micasense` Python library
3. Export calibrated reflectance orthomosaics (not raw DN values)
4. Calculate vegetation indices in R:

```{r}
#| label: ndvi-calc
#| echo: true
#| eval: false

library(terra)

# Load calibrated reflectance bands
red <- rast("drone/processed/orthomosaic/red_reflectance.tif")
nir <- rast("drone/processed/orthomosaic/nir_reflectance.tif")

# Calculate NDVI
ndvi <- (nir - red) / (nir + red)
names(ndvi) <- "NDVI"

writeRaster(ndvi, "drone/processed/orthomosaic/ndvi_2026-05-21.tif",
            overwrite = TRUE)
```


## Quality Control and Acceptance Criteria

- Photo alignment: at least 90% of images must align successfully
- Reprojection error: mean below 1.0 pixel after GCP optimization
- Check point RMSE: document and compare against project requirements
- Visual inspection: no visible seam lines, ghosting, or geometric
  distortion in the orthomosaic
- File integrity: verify that all exported GeoTIFFs have valid CRS
  headers and correct extent


## Waste Disposal

Not applicable (computational procedure).


## References

::: {#refs}
:::


## Training Record

Training is documented in the EA Program training log. Before
processing drone imagery independently, personnel must:

1. Read this SOP in full
2. Complete Agisoft Metashape (or ODM) training with a supervised
   practice dataset
3. Demonstrate a complete processing workflow from raw images to
   accuracy-assessed outputs
4. Be signed off by the PI or drone operations lead


## Appendix A: Processing Parameter Reference

| Parameter             | Low Quality    | Medium         | High (Recommended) |
|:----------------------|:---------------|:---------------|:-------------------|
| Photo alignment       | Low accuracy   | Medium         | High               |
| Dense cloud           | Low            | Medium         | High               |
| DEM resolution        | 10x GSD        | 5x GSD         | 2x GSD             |
| Orthomosaic           | 10x GSD        | 5x GSD         | Native GSD         |
| Approx. processing time (500 images) | 30 min | 2 hr | 6 hr+          |

## Appendix B: Troubleshooting

| Symptom                       | Probable Cause           | Corrective Action              |
|:------------------------------|:-------------------------|:-------------------------------|
| Low alignment rate (<80%)     | Insufficient overlap     | Re-fly with 80%+ front/side overlap |
| High reprojection error       | Incorrect GCP placement  | Re-identify GCP markers in photos |
| Artifacts in orthomosaic      | Moving objects (water, cars) | Mask affected areas; adjust blending |
| Out-of-memory error           | Dataset too large for RAM | Reduce quality; split into chunks |
| DSM shows "doming" distortion | No GCPs; insufficient geometry | Add GCPs; increase flight line diversity |
