\documentclass[12pt]{../SOP3_beta}
\usepackage[english]{babel}
\usepackage{blindtext}
\usepackage{lipsum}


\title{Surface Waters Sampling}
\date{02/02/2017}
\author{Marc Los Huertos}
\approved{Marc Los Huertos}
\ReviseDate{\today}
\SOPno{20 v0.1}

\usepackage{Sweave}
\begin{document}
\input{Surface_Water_Sampling_v0.1-concordance}

\maketitle

\section{Scope and Application}

\NP The scope of this SOP is train researchers to sample surface waters using the approved SWAMP protocols. 

\NP As a researcher, sampling surface waters that are consistent with state and federal regulators assures that data meet specified standards.

\section{Summary of Method}

\section{Definitions}

\section{Procedures}

\subsection{Preparation}

\subsection{Field Collections}

\NP 

\section{Interferences}

\section{Health and Safety}

\subsection{Safety and Personnnel Protective Equipment}


\section{Personnel \& Training Responsibilities}

Researchers using this SOP should be trained for the following SOPs:

\begin{itemize}
  \item SOP03 Field Work
  \item SOP01 Lab Safety
  \item SOP07
  \item SOP08
  \item SOP10
\end{itemize}

\section{Required Materials}

\subsection{Item 1 w/catalog number!}
\subsection{Item 2}

\section{Estimated Time}

\NP This will take XX minutes...

\section{Procedure}

\NP Prepare \dots

\NP

\section{References}

\NP APHA, AWWA. WEF. (2012) Standard Methods for examination of water and wastewater. 22nd American Public Health Association (Eds.). Washington. 1360 pp. (2014).

\NP \url{http://www.waterboards.ca.gov/water_issues/programs/swamp/docs/collect_bed_sediment_update.pdf}

\end{document}
